---
title: "GH2025_Manuscript"
author: "Tianyuan"
date: "2026-03-21"
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

```{r environment setup for FTIR spectra correction and normalisation, include = FALSE}
#Load necessary packages
{
  #Read OPUS files
  library(opusreader)
  
  #Smoothing
  library(prospectr)
  
  #Baseline correction
  library(baseline)
}

#Declare necessary functions
{
  #Declare a function to automatically split the name of spectral file, 
    #to get all necessary
    #folderNames - single character string, the original file name to be split
      #to extract the batch information
    #seperator - single character string, the symbol used in the original file 
      #name for separating different types of batch information
    #defaultSettings - vector with length of 10, indicating the default batch
      #information
  namesSplit <- function(folderNames, seperator, defaultSettings) {
    
    splitedNames <- defaultSettings
    
    #Split the names
    splitedNamesRaw <- strsplit( x = folderNames, split = seperator)[[1]]
    
    #Check the information given in the folder name, 
    #and change the corresponding sections in output if necessary
    if(splitedNamesRaw[2] != 'GH'){
      
      splitedNames[c(1:8, 10)] <- splitedNamesRaw[c(1:9)]
      
    } else if(splitedNamesRaw[4] == 'cul'){
      
      splitedNames[c(1,3:8, 10)] <- splitedNamesRaw[c(1:8)]
      
    }else {
      
      splitedNames[c(1,3:10)] <- splitedNamesRaw[c(1:9)]
      
    }
    
    if(splitedNamesRaw[5]  == 'ROOF' | splitedNamesRaw[6] == 'ROOF') {
      
      splitedNames[c(1,3:5, 8,10)]  <- splitedNamesRaw[c(1,5,3,4,6,7)]
      
      splitedNames[c(6,7,9)]  <- NA
      
    }

    return(splitedNames)
  }
  
  #Declare a function to calculate the number of all elements in a list given
    #tempList - list to be used for calculation
  numOfEle <- function(tempList) {
    
    #Vector storing the number
    tempVec <- 0
    
    for (loopI in 1:length(tempList)) {
      tempVec <- tempVec + length(tempList[[loopI]])
    }
    
    return(tempVec)
  }
  
  #Create a function to re-sample data according to given sequence from a vector
    #tempVec - numeric vector, the FTIR spectrum signals at original wavenumbers
    #inputSeq - numeric vector, the sequence of wavenumber for input FTIR 
      #spectrum
    #outputSeq - numeric vector, the sequence of wavenumber to be output
  resamData <- function(tempVec, inputSeq, outputSeq){
    tempOutVec <- approx(x = inputSeq,
                         y = tempVec,
                         xout = outputSeq)[[2]]
    
    return(tempOutVec)
  }
  
  #Function to normalise the spectral absorbance to maximum intensity
    #x - numeric vector, the FTIR spectrum to be normalised
  zstandMax <- function(x) {
    return((x - mean(x))/ sd(x))
  }
}
```

### FTIR spectra reading and format convertion. 
The original FTIR spectra collected from NMBU lab were in format of OPUS, which require a conversion to readable and editable format
```{r OPUS data reading and conversion, include = FALSE}
#Read all OPUS files
{
  #List all the folder for different measuring days
  daysVec <- list.files(path = file.path('..', 'Data_R', 'FTIR_Raw_Spectra/'))
  
  #Loop to list all .o files in each folder above
  #List for storing all names of spectra files
  namesList <- list()
  for (loopI in 1:length(daysVec)) {
   
    namesList[[length(namesList) + 1]] <- 
      list.files(path = file.path('..', 'Data_R', 'FTIR_Raw_Spectra',
                                  daysVec[loopI]), 
                 pattern = '\\.0$')
     
  }
  
  #Loop to read all spectra
  #List for storing all spectra
  rawSpecList <- list()
  
  #Matrix for storing all batch information
  rawSpecBatchInfo <- matrix(nrow = numOfEle(namesList),
                             ncol = 10,
                             dimnames = list(c(),
                                             c('Exp_Year',
                                               'Sample_Year',
                                               'Location',
                                               'Species',
                                               'Materials',
                                               'Group',
                                               'Treatment',
                                               'Sample_ID',
                                               'Sample_Dup',
                                               'FTIR_Dup')))
  
  
  #Two recorders on row index for matrix filling
  tempInd <- 0
  
  for (loopI in 1:length(daysVec)) {
    
    rawSpecList[[loopI]] <- list()
    
    for (loopJ in 1:length(namesList[[loopI]])) {
      rawSpecList[[loopI]][[loopJ]] <- 
        opus_read(file = 
                    file.path('..',
                              'Data_R',
                              'FTIR_Raw_Spectra',
                              daysVec[loopI], 
                              namesList[[loopI]][loopJ]
                              ),
                  simplify = TRUE)
      
      #Extract the sample information from the file names at the same time
      tempVec <- namesSplit(substr(namesList[[loopI]][loopJ],
                                   1,
                                   nchar(namesList[[loopI]][loopJ]) - 2), '_',
                            c(2025,2025,
                              'GH', 'Pmu', 
                              'cul', 'A', 
                              'UV', '001',
                              '1', '1'))
      
        
      tempInd <- tempInd + 1
      
      rawSpecBatchInfo[tempInd,] <- tempVec
    
      rm(tempVec)
    }
  }
  
  #Create a matrix to store all raw spectra
  rawSpecMat <- matrix(nrow = numOfEle(namesList),
                       ncol = length(rawSpecList[[1]][[1]][['spec']]),
                       dimnames = list(c(),
                                       c(rawSpecList[[1]][[1]][['wavenumbers']])
                                       )
                       )
  
  #Loop to fill the matrix
  #Create a vector recording the current index
  tempIndRawSpec <- 0
  
  for (loopI in 1:length(rawSpecList)) {
    
    for (loopJ in 1:length(rawSpecList[[loopI]])) {
      
      tempIndRawSpec <- tempIndRawSpec + 1
      
      rawSpecMat[tempIndRawSpec,] <- rawSpecList[[loopI]][[loopJ]][['spec']]
    }
  }
  
  #Re-sample the spectra to a fixed sequence of wavenumber
  #Specify the target wavenumber sequence
  tempWaveSeq <- seq(from = 3996, to = 600, by = -4)
  
  rawSpecResamMat <- t(apply(rawSpecMat, 1, 
                           FUN = resamData, 
                           inputSeq = as.numeric(colnames(rawSpecMat)),
                           outputSeq = tempWaveSeq))
  
  colnames(rawSpecResamMat) <- tempWaveSeq
  
  #Remove redundant variables
  rm(loopI,
     loopJ,
     tempInd,
     tempIndRawSpec)
}
```

### Correction and standardisation on raw FTIR spectra. 
The FTIR spectra used further for all analysing purposes in the manuscript were firstly corrected following the routes of: 1. Smoothing(savitzkyGolay, m = 0, p = 2, w = 11) and baseline correction (baseline.modpolyfit, degree = 3, tol = 0.00002, rep = 100)

```{r smoothing and baseline correction on raw FTIR spectra, include = FALSE}
#Smoothing and baseline correction
{
  #Apply savitzky-Golay filter
  smooSpec <- savitzkyGolay(apply(rawSpecResamMat, c(1,2), as.numeric),
                            m = 0, p = 2, w = 11)
  
  #Apply polynomial linear baseline model
  bcSpec <- baseline.modpolyfit(smooSpec,
                                degree = 3, tol = 0.00002, 
                                rep = 100)[[2]]
  
  #Rename the colomn of baseline corrected matrix, 
    #using the correct wavenumber sequence
  colnames(bcSpec) <- tempWaveSeq[c(-1:-5, 
                                    (-length(tempWaveSeq) +4): 
                                      -length(tempWaveSeq))]
}
```

2.  Z-score standardisation

```{r z-score standardisation, include = FALSE}
  #Z-score
  bcNorSpec <- t(apply(bcSpec, 1, zstandMax))
```

3.  Output the correction and standardised spectra

```{r output for corrected and standardised spectra, include = FALSE}
  #Z-score
  write.csv(cbind(rawSpecBatchInfo, bcNorSpec), 
            file = file.path('..', 'Output_R', 'bcNorSpec.csv'))
```

### FTIR spectra processing

```{r envionment setup for data visualisation, include = FALSE}
#Load necessary packages--------------------------------------------------------
{
  #Find peaks and extract peak statistics
  library(photobiology)
  
  #Paralell processing
  library(doSNOW)
  
  #Plot
  library(ggplot2)
  library(patchwork)
  
  #Load arial font into R
  library(showtext)
  font_add(family = "arial", regular = file.path('..', 
                                                 'Packages_R', 
                                                 'Fonts',
                                                 'arial.ttf'))
  showtext_auto()
  
  #PCA
  library(factoextra)
  library(FactoMineR)
  
  #Palette
  library(RColorBrewer)
  
  #Log-ratio transformation
  library(compositions)
  library(easyCODA)
  library(robCompositions)
  
  #PLS
  library(pls)
 
  #Smoothing
  library(prospectr)
  
  #Baseline correction
  library(baseline) 
  
  #Data manipulation
  library(reshape2)
  
  #String edit
  library(stringr)
}

#Declare necessary functions----------------------------------------------------
{
  #Declare a function to calculate the derivatives of a given spectrum
    #specVec - one single spectrum in a numeric vector
    #degree - degree of derivative spectrum
  derivaSpectrum <- function(specVec, degree) {
    
    tempvec <- specVec
    
    #Loop to calculate the derivative
    loopI <- 1
    while (loopI <= degree) {
      
      tempVecLast <- tempvec[-1]
      
      tempVecFirst <- tempvec[-length(tempvec)]
      
      tempvec <- (tempVecLast - tempVecFirst)
      
      loopI <- loopI + 1
    }
    
    return(tempvec)
  }
  
  #Benjamin A Bell github codes, for detecting and calculating peak area--------
  {
    # Load peak detection script
    source("https://raw.githubusercontent.com/benbell95/peak-detection/main/r/peak_detection.r")
    # Load peak area script
    source("https://raw.githubusercontent.com/benbell95/peak-detection/main/r/peak_area.r")
  }
  
  #Create a function to automatically identify all qualified peaks in given
    #spectrum, and extract the peak height and peak area under each peaks
    #tempSpec - one single spectra in numeric vector, with wavenumber information
      #saved in elements' names
    #tempMinPeakWid - one numeric value indicating the minimum peak width 
      #to be identified, should be bigger than the wavenumber interval in the 
      #spectrum input
    #tempMinPeakWidMax - one numeric value indicating the maximum peak width 
      #to be identified, should be bigger than tempMinPeakWid
    #plot - Boolean value, indicating whether a plot on current spectrum labelled
      #with identified peaks is output
    #tempTol - one numeric value defining the minimum peak singal to be 
      #identified
  autoPeakDecQuan <- function(
    tempSpec, 
    tempMinPeakWid, 
    tempMinPeakWidMax,
    plot = FALSE, 
    tempTol = 0
    ) {
    #Extract the spectra resolution
    tempResu <- floor(tempMinPeakWid / 
                        abs(as.numeric(names(tempSpec))[2] - as.numeric(names(tempSpec))[1])
    )
    
    tempResu2 <- floor(tempMinPeakWidMax / 
                        abs(as.numeric(names(tempSpec))[2] - as.numeric(names(tempSpec))[1])
    )
    
    #Get peak positions (central position) for all peaks identified in current
      #spectrum
    tempPeakPos <- as.numeric(names(tempSpec))[
      which(find_peaks(x = as.numeric(tempSpec),
                       local.threshold = tempTol,
                       span = tempResu,
                       strict = FALSE)
            )
      ]
    
    #Find the peak ranges, acting on the peak position identified before
    #Create a matrix storing the peak range information
    tempPeakRange <- matrix(nrow = length(tempPeakPos),
                            ncol = 5,
                            dimnames = list(c(),
                                            c('Peak_Pos',
                                              'Peak_L',
                                              'Peak_R',
                                              'Peak_Height',
                                              'Peak_Area')
                            )
    )
    
    #Fill in the matrix
    tempPeakRange[ ,1] <- tempPeakPos
    
    #Loop to go through all identified peaks - left
    for (loopI in 1:length(tempPeakPos)) {
      
      tempLeftSig <- tempSpec[which(as.numeric(names(tempSpec)) == 
                                      tempPeakPos[loopI]
      )
      ]
      
      #Assign an indicator to terminate the loop for range searching
      tempIndcator <- 1
      
      tempIndLeftIndcator <- 1
      
      while (tempIndcator == 1) {
        
        if ((which(as.numeric(names(tempSpec)) == 
                  tempPeakPos[loopI]) - tempIndLeftIndcator < 1) | (tempIndLeftIndcator > tempResu2)) {
          
          #Assign the terminate indicator to 0, for preparing to stop the loop
          tempIndcator <- 0
          
          break
          
        } else {
          
          if (tempLeftSig > tempSpec[which(as.numeric(names(tempSpec)) == 
                                           tempPeakPos[loopI]
          ) - tempIndLeftIndcator]) {
            
            tempLeftSig <- tempSpec[which(as.numeric(names(tempSpec)) == 
                                            tempPeakPos[loopI]
            ) - tempIndLeftIndcator]
            
            tempIndLeftIndcator <- tempIndLeftIndcator + 1
            
          }else{
            
            #Assign the terminate indicator to 0, for preparing to stop the loop
            tempIndcator <- 0
            
            #This indicate that the search might approach its local peak
              #boundary, then apply a boundary checking referring to the 
              #boundary tolerance index defined by the spectra resolution
            for (loopJ in 1:ceiling(tempResu / 4)) {
              
              if ((which(as.numeric(names(tempSpec)) == 
                         tempPeakPos[loopI]
              ) - tempIndLeftIndcator - loopJ) < 1) {
                
                tempRightSig <- tempSpec[which(as.numeric(names(tempSpec)) == 
                                                 tempPeakPos[loopI]
                ) + tempIndLeftIndcator - loopJ + 1]
                
                tempIndcator <- 0
                
                tempIndLeftIndcator <- tempIndLeftIndcator + loopJ
                
                break
                
              }else {
                
                #Search for the local minimum for left section of the current peak
                if (tempLeftSig > tempSpec[which(as.numeric(names(tempSpec)) == 
                                                 tempPeakPos[loopI]
                ) - tempIndLeftIndcator - loopJ]) {
                  
                  tempLeftSig <- tempSpec[which(as.numeric(names(tempSpec)) == 
                                                  tempPeakPos[loopI]
                  ) - tempIndLeftIndcator - loopJ]
                  
                  #Reassign the terminate indicator to 1, to continue the loop
                  tempIndcator <- 1
                  
                  tempIndLeftIndcator <- tempIndLeftIndcator + loopJ
                  
                  #Stop the loop for boundary detection
                  break
                }
              }
            }
          }
        }
        #Record the peak range inot matrix
        tempPeakRange[loopI, 2] <- names(tempSpec)[which(as.numeric(names(tempSpec)) == 
                                                           tempPeakPos[loopI]) - 
                                                     tempIndLeftIndcator + 1
        ]
      }
    }
    
    #Loop to go through all identified peaks - right
    for (loopI in 1:length(tempPeakPos)) {
      
      tempRightSig <- tempSpec[which(as.numeric(names(tempSpec)) == 
                                       tempPeakPos[loopI]
      )
      ]
      
      #Assign an indicator to terminate the loop for range searching
      tempIndcator <- 1
      
      tempIndRightIndcator <- 1
      
      while (tempIndcator == 1) {
        
        if (which(as.numeric(names(tempSpec)) == 
                  tempPeakPos[loopI]) + tempIndRightIndcator > ncol(tempSpec) | (tempIndRightIndcator > tempResu2)) {
          
          #Assign the terminate indicator to 0, for preparing to stop the loop
          tempIndcator <- 0
          
          break
          
        } else {
          
          if (tempRightSig > tempSpec[which(as.numeric(names(tempSpec)) == 
                                            tempPeakPos[loopI]
          ) + tempIndRightIndcator]) {
            
            tempRightSig <- tempSpec[which(as.numeric(names(tempSpec)) == 
                                             tempPeakPos[loopI]
            ) + tempIndRightIndcator]
            
            tempIndRightIndcator <- tempIndRightIndcator + 1
            
          }else{
            
            #Assign the terminate indicator to 0, for preparing to stop the loop
            tempIndcator <- 0
            
            #This indicate that the search might approach its local peak
              #boundary, then apply a boundary checking referring to the 
              #boundary tolerance index defined by the spectra resolution
            for (loopJ in 1:ceiling(tempResu / 4)) {
              
              if ((which(as.numeric(names(tempSpec)) == 
                         tempPeakPos[loopI]
              ) + tempIndRightIndcator + loopJ) > ncol(tempSpec)) {
                
                tempRightSig <- tempSpec[which(as.numeric(names(tempSpec)) == 
                                                 tempPeakPos[loopI]
                ) + tempIndRightIndcator + loopJ - 1]
                
                tempIndcator <- 0
                
                tempIndRightIndcator <- tempIndRightIndcator + loopJ
                
                break
                
              }else {
                
                #Search for the local minimum for left section of the current peak
                if (tempRightSig > tempSpec[which(as.numeric(names(tempSpec)) == 
                                                  tempPeakPos[loopI]
                ) + tempIndRightIndcator + loopJ]) {
                  
                  tempRightSig <- tempSpec[which(as.numeric(names(tempSpec)) == 
                                                   tempPeakPos[loopI]
                  ) + tempIndRightIndcator + loopJ]
                  
                  #Reassign the terminate indicator to 1, to continue the loop
                  tempIndcator <- 1
                  
                  tempIndRightIndcator <- tempIndRightIndcator + loopJ
                  
                  #Stop the loop for boundary detection
                  break
                }
                
              }
            }
          }
        }
      }
      
      #Record the peak range inot matrix
      tempPeakRange[loopI, 3] <- names(tempSpec)[which(as.numeric(names(tempSpec)) == 
                                                         tempPeakPos[loopI]) + 
                                                   tempIndRightIndcator - 1
      ]
      
      tempPeakRange[loopI, 4] <- as.numeric(tempSpec)[which(as.numeric(names(tempSpec)) ==
                                                              tempPeakPos[loopI])
      ]
      
      tempPeakRange[loopI, 5] <- unlist(spd_area_bands(t(rev(rbind(tempSpec, rep(0, length(tempSpec))))),
                                                       bands = tempPeakRange[loopI, c(3,2)],
                                                       refine = FALSE)$area)[1]
    }
    
    
    #Plot the current spectrum with information extracted, if corresponding
    #variable is set as true
    if (plot == TRUE) {
      #View an example of specta read
      plot(x = as.numeric(names(tempSpec)),
           y = tempSpec,
           type = 'l',
           xlim = rev(range(as.numeric(names(tempSpec)
           )
           )
           )
      )
      
      #Add reference lines around peaks identified on the sample spectra
      abline(v=tempPeakPos, col="blue")
    }
    
    #Output the matrix
    return(tempPeakRange)
  }
  
  #Declare a function to combine all matrix in a list into one matrix
    #tempList - a list to be combined, each element in this list should be a 
      #matrix, all matrix in different elements should have the same column 
      #title, this list must have length bigger than 1
  combineList <- function(tempList) {
    
    #Create a temporary matrix, and put the fisrt list element into the matrix
    tempMat <- tempList[[1]]
    
    #Loop to combind all other list elements into the matrix above
    for (loopI in 2:length(tempList)) {
      tempMat <- rbind(tempMat, tempList[[loopI]])
    }
    
    #Output the combined matrix
    return(tempMat)
  }
  
  #Declare a function to combine some of rows and columns in all matrix in a 
    #list into one matrix
    #tempList - a list to be combined, each element in this list should be a 
      #matrix, all matrix in different elements should have the same column 
      #title, this list must have length bigger than 1
    #indexRow - a numeric vector indicating the indexes of all rows in each
      #matrix which will be combined, should have length bigger than 1
    #indexCol - optional, a numeric vector indicating the indexes of all columns
      #in each matrix which will be combined. Will be set as all column indexes
      #in each matrix if not specified
  combineListEle <- function(tempList, indexRow, indexCol) {
    
    #Check whether variable indexCol is input. Set it to all column indexes in  
      #single matrix if not specified
    if (exists('indexcol')) {
      indexcol <- indexCol
    }else{
      indexcol <- 1:ncol(tempList[[1]])
    }
    
    #Create a temporary matrix and copy the first matrix element in the list 
      #into it
    tempMat <- tempList[[1]][indexRow, indexCol]
    
    #Loop to combine all specified rows and columns in each matrix element in 
      #the list together
    for (loopI in 2:length(tempList)) {
      tempMat <- rbind(tempMat, tempList[[loopI]][indexRow, indexCol])
    }
    
    #Output the combined matrix
    return(tempMat)
  }
  
  #Create a function to find peaks with  similar positions among different peak
    #information matrix, and modified all original peak information matrix into
    #another one that deletes the singular peak positions (the peaks that is 
    #not identified in all of other spectra)
    #peakInfoList - the original list of peak information, each element in this
      #list is a matrix containing all peak statistics identified from function
      #'autoPeakDecQuan', should have length bigger than 1
    #tempPeakPos - numeric vector, indicating the peak position to be remained,
      #should have length bigger than 1
    #temppeakTol -  numeric vector, indicating the tolerance (drift 
      #distance between the theoretical peak position specified in variable
      #'tempPeakPos', peaks farer than this distance will not be considered as
      #the same peak and will be deleted at last. Should have the same length as
      #variable 'tempPeakPos'
  autoPeakResam <- function(peakInfoList, tempPeakPos, temppeakTol) {
    
    #Combine the input peak information list into one matrix
    tempMat <- combineList(peakInfoList)
    
    #Create a matrix to resample from the peak information matrix
    tempPeakInfoMat <- matrix(nrow = length(tempPeakPos),
                              ncol = 2 + ncol(tempMat),
                              dimnames = list(c(),
                                              c('Pre_Peak_Pos',
                                                'Peak_Pos_Tol',
                                                colnames(tempMat)
                                              )
                              )
    )
    
    #Fill in the matrix
    tempPeakInfoMat[, 1] <- tempPeakPos
    tempPeakInfoMat[, 2] <- temppeakTol
    
    #Create a list to store the resampled peak information
    tempList <- vector('list', length = length(peakInfoList))
    
    #Loop around all spectra
    for (loopI in 1:length(tempList)) {
      
      tempList[[loopI]] <- tempPeakInfoMat
      
      #Loop around possible peak positions to find the nearest peak identified in the
      #Spectra
      for (loopJ in 1:nrow(tempPeakInfoMat)) {
        
        #Check if there is at least one peak matched with any of peaks listed in
          #'tempPeakPos'
        if (length(which(
          (abs(
            as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
          ) == 
          min(
            abs(
              as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
            ))) & (abs(
              as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
            ) <= temppeakTol[loopJ]
            ))) >= 1){
          
          #Check if there is more than one peak matched with the same peak 
            #listed in 'tempPeakPos'
          if (length(which(
            (abs(
              as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
            ) == 
            min(
              abs(
                as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
              ))) & (abs(
                as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
              ) <= temppeakTol[loopJ]
              ))) > 1) {
            
            #if there is more than one peaks matching the same peak in 
              #'tempPeakPos'
            tempList[[loopI]][loopJ, 3:ncol(tempList[[loopI]])] <-
              apply(peakInfoSpec[[loopI]][which(
                (
                  abs(
                    as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
                  ) == 
                    min(
                      abs(
                        as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
                      )
                    )
                ) & (
                  abs(
                    as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
                  ) <= temppeakTol[loopJ]
                )
              ), ], 2, mean)
          }else {
            
            #If there is only one peak matching a peak in 'tempPeakPos'
            tempList[[loopI]][loopJ, 3:ncol(tempList[[loopI]])] <-
              peakInfoSpec[[loopI]][which(
                (
                  abs(
                    as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
                  ) == 
                    min(
                      abs(
                        as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
                      )
                    )
                ) & (
                  abs(
                    as.numeric(peakInfoSpec[[loopI]][, 1]) - tempPeakPos[loopJ]
                  ) <= temppeakTol[loopJ]
                )
              ), ]
          }
        }
      }
      
      #Adjust the value format into numeric
      tempList[[loopI]] <- apply(tempList[[loopI]], c(1,2), as.numeric)
    }
    
    #Output the modified list that deletes peaks that do not show in all spectra
    return(tempList)
  }
  
  #Check NAs in a vector, return true if the correconding element is NA
    #tempVec - numeric vector to be treated for deleting NAs, should have length
      #bigger than 1
  cleanNA <- function(tempVec) {
    
    for (loopI in 1:length(tempVec)) {
      
      if (is.na(tempVec[loopI])) {
        
        return(FALSE)
        
        break
        
      }else if(loopI == length(tempVec)) {
        
        return(TRUE)
        
      }
    }
  }
  
  #Detect the symbol of given value, and output 1 or -1 accordingly
    #tempVec - single numeric value to be treated
  symbolDec <- function(tempVec){
    
    tempVec2 <- tempVec
    
    for (loopI in 1:length(tempVec)) {
      if (tempVec[loopI] <= 0) {
        
        tempVec2[loopI] <- -1
        
      }else{
        
        tempVec2[loopI] <- 1
        
      }
    }
    
    return(tempVec2)
    
  }
  
  #Function for producing a text string with regression function and coefficients
  lm_eqn2 <- function(compI){
    m <- lm(modelHeight$model$calibrated ~ modelHeight$fitted.values[,,compI]);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                     list(a = format(unname(coef(m)[1]), digits = 2),
                          b = format(unname(coef(m)[2]), digits = 2),
                          r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
  }
  
  #Create a function to output significance markers
    #tempSig - single numeric value
    #gradient - numeric vector, significance criteria judging the significant 
      #markers for different levels, should be at length of 3, and in decreasing
      #order
  sigMarkter <- function(tempSig, gradient) {
    
    #Significant level 1, marked by *
    if ((as.numeric(tempSig) < gradient[1]) & 
        (as.numeric(tempSig) > gradient[2])) {
      
      return('*')
      
    #Significance level 2, marked by **  
    } else if ((as.numeric(tempSig) < gradient[2]) & 
               (as.numeric(tempSig) > gradient[3])) {
      
      return('**')
    
    #Significance level 3, marked by ***
    } else if (as.numeric(tempSig) < gradient[3]) {
      
      return('***')
      
    } else {
      
    #Not significant, marked by nothing
      return('')
      
    }
    
  }
} 
```

Read the standardised spectra and identify all peaks in each spectrum. I
use some codes (integraated in the function of autoPeakDecQuan)
following these routes: 1. Identify all local maximum in a spectrum,
which could potentially (but not exactly) be a peak (can be named
'candidate peak'), referring to the function of 'find_peaks' in
'photobiology' package 2. For each spectrum, and acting on the
(candidate) peaks' positions , the codes search for boundary for each
candidate peak point by point, starting from the exactly signal local
maximum (position of the candidate peak), and moving to both left and
right direction relative to the local maximum. I define the 'boundary'
of a peak as 'the point where the decreasing or increasing trend between
adjacent two points disappears consecutively for given times (this is
set as 3 points here) 3. Basing on the peak boundaries found in the last
step, the codes calculates the peak statistics (peak height, peak area,
peak width), and will exclude peaks with width wider or narrower than
the expected value (both can be set though corresponding variables of
the function). For the qualified peaks passing the width examination, a
matrix containing thier statistics will be finally output.

```{r FTIR spectra, include = FALSE}
#Read FTIR data and peak signal extration---------------------------------------
{
  #Read raw spectra (smoothed and baseline corrected, z-score standardisation)
  {
    #Read standardised FTIR spectra
    bcSpec <- read.csv(file.path('..', 'Output_R', 'bcNorSpec.csv'),
                       row.names = 1,
                       header = TRUE)
    
    #Remove spectra of roof samples
    bcSpec <- bcSpec[which(bcSpec[, 3] != 'ROOF'),]
    
    #Rename the column names for the smoothed spectra
    colnames(bcSpec)[11:ncol(bcSpec)] <- seq(3976, 620, by = -4)
    
    #Convert spectra into numeric values
    bcSpec[, 11:ncol(bcSpec)] <- apply(bcSpec[, 11:ncol(bcSpec)], 
                                       c(1,2),
                                       as.numeric)
  }
  
  #Extract peak information (position and range) from raw spectra
  {
    #Create a list to store all peak information for all spectra
    peakInfoSpec <- vector('list', nrow(bcSpec))
    
    #Setup cores for parallel processing
    cl <- makeCluster(8, type="SOCK")
    
    #Regist multiple-core setup
    registerDoSNOW(cl)
    
    #Apply peak identification function on all spectra, and output the peak
      #information matrix into a list
    peakInfoSpec <- foreach(loopI = 1:length(peakInfoSpec),
                            .packages = c('photobiology')) %dopar% 
      {
        autoPeakDecQuan(bcSpec[loopI, -1:-10], 20, 100, FALSE
        )
      }
    
    #Quite the setup of multiple-core processing
    stopCluster(cl)
    
    #Clean up some of variables
    rm(cl)
    
    #Assign possible peak positions together with tolerance range for peak drifts
    peakPos <- c(3324, 2924, 2856, 1740, 
                 1636, 1556, 1516, 1440, 1376,
                 1324, 1268, 1036, 996, 832)
    peakTol <- c(200, rep(50, 13))
    
    #Rearrange the peak information into an uniform sequence
    peakInfoSpecUni <- autoPeakResam(peakInfoSpec, peakPos, peakTol)
  }
  
  #Rearrange all peak information into one matrix
  {
    #Create a matirx for all peak information - Height
    pcaMatHeight <- matrix(nrow = nrow(bcSpec),
                           ncol = 10 + length(peakPos),
                           dimnames = list(c(),
                                           c(colnames(bcSpec)[1:10], 
                                             peakPos)))
    
    #Create a matirx for all peak information - Area
    pcaMatArea <- matrix(nrow = nrow(bcSpec),
                         ncol = 10 + length(peakPos),
                         dimnames = list(c(),
                                         c(colnames(bcSpec)[1:10], 
                                           peakPos)))
    
    #Fill the batch information into matrix for peak height and area
    pcaMatHeight[, 1:10] <- as.matrix(bcSpec[, 1:10])
    pcaMatArea[, 1:10] <- as.matrix(bcSpec[, 1:10])
    
    #Loop to fill in the matrix with individual spectrum on corresponding row
    for (loopI in 1:nrow(bcSpec)) {
      
      pcaMatHeight[loopI, c(11:ncol(pcaMatHeight))] <- peakInfoSpecUni[[loopI]][, 6]
      
      pcaMatArea[loopI, 11:ncol(pcaMatArea)] <- peakInfoSpecUni[[loopI]][, 7]
      
    }
  }
}
```

### GCMS signal processing

```{r GCMS signals, and matching them with FTIR spectra, include = FALSE}
#Read GC-MS signals and combine them with FTIR peak signals---------------------
{
  #Read GC-MS data
  {
    #GC-MS data on cuttings and pot cultivars
    GCDataRawCut <- as.matrix(read.csv(
      file.path('..', 'Data_R', '04_11_2025_gc_cutting_data_2025.csv'),
      header = TRUE,
      row.names = 1))
    GCDataRawCul <- as.matrix(read.csv(
      file.path('..', 'Data_R', '01_07_2025_gc_cultivar_data_2025.csv'),
      header = TRUE,
      row.names = 1))
    
    #Combine these two raw datasets, with batch information
    GCData <- rbind(GCDataRawCut[, c(8, 10:14)],
                    cbind(GCDataRawCul[, c(8, 10:12)], 
                          sampling_number = NA, 
                          GCDataRawCul[, 13])
    )
    
    #Exclude the last batch (u), as an unexpected crash happened before this
      #batch, making reseult from this batch abnormal
    GCData <- GCData[-which(GCData[, 6] == 'u'),]
    
    #Loop to calculate the mean values for each sample
    uniqueBatchInfo <- unique(GCData[, 2:5])
    
    #Create a matrix to fill in the mean GC-MS signals
    GCDataMean <- matrix(nrow = nrow(uniqueBatchInfo),
                         ncol = ncol(uniqueBatchInfo) + 2,
                         dimnames = list(c(),
                                         c('calibrated',
                                           'sd',
                                           colnames(uniqueBatchInfo)
                                         )
                         )
    )
    
    #Add batch number to the GCMS data matrix
    GCDataMean[, 3:ncol(GCDataMean)] <- uniqueBatchInfo
    
    #Loop to calculate the mean signals of GCMS duplicates
    for (loopI in 1:nrow(GCDataMean)) {
      
      #Check wheter current sample is cutting or pot cultivar 
      if (is.na(GCDataMean[loopI, 6])) {
        
        #Mean and sd for pot cultivar
        GCDataMean[loopI, 1] <- mean(as.numeric(
          GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                         (GCData[, 3] == GCDataMean[loopI, 4]) &
                         (GCData[, 4] == GCDataMean[loopI, 5]) 
        )
        , 1])
        )
        GCDataMean[loopI, 2] <- sd(as.numeric(
          GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                         (GCData[, 3] == GCDataMean[loopI, 4]) &
                         (GCData[, 4] == GCDataMean[loopI, 5])
        )
        , 1])
        )
        
      }else {
        
        #Mean and sd for cutting
        GCDataMean[loopI, 1] <- mean(as.numeric(
          GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                         (GCData[, 3] == GCDataMean[loopI, 4]) &
                         (GCData[, 4] == GCDataMean[loopI, 5]) &
                         (GCData[, 5] == GCDataMean[loopI, 6])
        )
        , 1])
        )
        GCDataMean[loopI, 2] <- sd(as.numeric(
          GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                         (GCData[, 3] == GCDataMean[loopI, 4]) &
                         (GCData[, 4] == GCDataMean[loopI, 5]) &
                         (GCData[, 5] == GCDataMean[loopI, 6])
        )
        , 1])
        )
        
      }
    }
    
    #Transfer the matrix back to data frame, and transfer values to number if
      #necessary
    GCDataMean <- as.data.frame(GCDataMean)
    GCDataMean[, 1] <- as.numeric(GCDataMean[, 1])
  }
  
  #Combine GC-MS data with FTIR data
  {
    #Create a vector to stroe the index for batch information matching
    tempInd <- vector('numeric', length = nrow(bcSpec))
    
    #Loop to match the batch information between GC-MS and FTIR datasets
    for (loopI in 1:nrow(bcSpec)) {
      
      #Check whether current sample is pot cultivar
      if (bcSpec[loopI, 5] == 'cul') {
        
        #Check whether current FTIR spectrum has any GCMS signal
        if (any((GCDataMean[, 3] == bcSpec[loopI, 6]) &
                (GCDataMean[, 4] == bcSpec[loopI, 7]) &
                (GCDataMean[, 5] == bcSpec[loopI, 8]) &
                (is.na(GCDataMean[, 6]))
        )
        ) {
          
          #Record the FTIR matrix row index for current GCMS signal
          tempInd[loopI] <- which((GCDataMean[, 3] == bcSpec[loopI, 6]) &
                                    (GCDataMean[, 4] == bcSpec[loopI, 7]) &
                                    (GCDataMean[, 5] == bcSpec[loopI, 8]) &
                                    (is.na(GCDataMean[, 6])))
          
        }else {
          
          #Record nothing if there is no match between current FTIR specrum and
            #any GCMS signal
          tempInd[loopI] <- NA
          
        }
        
      #Check whether current sample is cutting
      } else if (bcSpec[loopI, 5] == 'cut') {
        
        #Check whether current FTIR spectrum has any GCMS signal
        if (any((GCDataMean[, 3] == bcSpec[loopI, 6]) &
                (GCDataMean[, 4] == bcSpec[loopI, 7]) &
                (GCDataMean[, 5] == bcSpec[loopI, 8]) &
                (GCDataMean[, 6] == bcSpec[loopI, 9])
        )
        ) {
          
          #Record the FTIR matrix row index for current GCMS signal
          tempInd[loopI] <- which((GCDataMean[, 3] == bcSpec[loopI, 6]) &
                                    (GCDataMean[, 4] == bcSpec[loopI, 7]) &
                                    (GCDataMean[, 5] == bcSpec[loopI, 8]) &
                                    (GCDataMean[, 6] == bcSpec[loopI, 9]))
          
        }else {
          
          #Record nothing if there is no match between current FTIR spectrum and
            #any GCMS signal
          tempInd[loopI] <- NA
          
        }
        
      } else {
        
        #Record nothing if there is no match between current FTIR spectrum and
          #any GCMS signal
        tempInd[loopI] <- NA
        
      }
    }
    
    #Attach GCMS signals to  peak statistics (height or area)
    pcaMatHeight <- cbind(pcaMatHeight,
                          GCDataMean[tempInd, c(1,2)])
    pcaMatArea <- cbind(pcaMatArea,
                        GCDataMean[tempInd, c(1,2)])
    
    rm(loopI,
       tempInd)
  }
}
```

### PLS analysis 
PLS analysis is based on identifed and checked peaks, where 14 peaks are input as variables in PLS model. Some peaks were not used for PLS analysis as they are not identifed in every spectrum.

```{r PLS analysis, include = FALSE}
#PLS analysis-------------------------------------------------------------------
{
  #Height
  {
    #Clean up the matrix to exclude NA values in GC-MS data
    pcaMatHeight <- pcaMatHeight[which(!is.na(pcaMatHeight[, ncol(pcaMatHeight) - 1])), ]
    pcaMatHeight <- pcaMatHeight[which(apply(pcaMatHeight[, -1:-10], 1, cleanNA)),]

    #PLS model
    modelHeight <- plsr(calibrated ~ ., 
                        data = as.data.frame(apply(pcaMatHeight[,c(-1:-10, -ncol(pcaMatHeight))],
                                                   c(1,2),
                                                   as.numeric)), 
                        scale=FALSE, validation="CV")
    
    #Investigate important results
    validationplot(modelHeight)
    validationplot(modelHeight, val.type="R2")
    
    #Setup the number of components to be analysed
    numCompsHeight <- selectNcomp(modelHeight, method = "randomization", plot = TRUE)
    
    #Plot PLS predictions with GC-MS
    plsPlotHeight <- ggplot() + 
      geom_point(aes(x = modelHeight$fitted.values[,,numCompsHeight],
                     y = modelHeight$model$calibrated)) + 
      geom_smooth(aes(x = modelHeight$fitted.values[,,numCompsHeight],
                      y = modelHeight$model$calibrated),
                  method = "lm") +
      geom_text(
        aes(x = 0.35, y = 0.7,
            label = lm_eqn2(numCompsHeight)), parse = TRUE) +
      labs(x = 'PLS fitted values (peak height)',
           y = 'GC-MS (ng/150gr)') + 
      theme_classic() + 
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #Plot correlation coefficient under the selected number of components
    modelHeightCof <- ggplot() +
      geom_bar(aes(x = peakPos, y = modelHeight$coefficients[, , numCompsHeight]),
               stat = 'identity') +
      geom_hline(yintercept = 0) +
      geom_text(
        aes(x = peakPos, y = modelHeight$coefficients[, ,numCompsHeight] + 
              min(abs(modelHeight$coefficients[, ,numCompsHeight])) *0.4 * symbolDec(modelHeight$coefficients[, ,numCompsHeight]),
            label = peakPos), parse = TRUE, 
        check_overlap = T) +
      scale_x_reverse(limit = c(3500, 500),
                      breaks = seq(3500, 500, by = -500)) +
      labs(x = 'Wavenumber', y = paste('Variable coefficients ',
                                       numCompsHeight, ' components', sep = '', collapse = NULL)) +
      theme_classic() + 
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #Plot variable loadings under the first components
    modelHeightLoa <- ggplot() +
      geom_bar(aes(x = peakPos, y = modelHeight$loadings[, 1]),
               stat = 'identity') +
      geom_hline(yintercept = 0) +
      geom_text(
        aes(x = peakPos, y = modelHeight$loadings[, 1] + 
              min(abs(modelHeight$loadings[, 1])) *0.4 * symbolDec(modelHeight$loadings[, 1]),
            label = peakPos), parse = TRUE, 
        check_overlap = T) +
      scale_x_reverse(limit = c(3500, 500),
                      breaks = seq(3500, 500, by = -500)) +
      labs(x = 'Wavenumber', y = paste('Variable loadings ',
                                       1, sep = '', collapse = NULL)) +
      theme_classic() + 
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #Calculate variance explained by different components
    varCompMat <- matrix(nrow = numCompsHeight,
                         ncol = 2,
                         dimnames = list(c(),
                                         c('Accu',
                                           'Partial')))
    varCompMat[, 1] <- modelHeight[[12]][1:numCompsHeight]
    for (loopI in 1:nrow(varCompMat)) {
      varCompMat[loopI, 2] <- sum(modelHeight[[12]][1:loopI])
    }
    
    #Plot variance explained by different components
    modelHeightVar <- ggplot() +
      geom_bar(aes(x = 1:numCompsHeight, y = varCompMat[, 2] / sum(modelHeight[[12]])),
               fill = 'red',
               stat = 'identity') +
      geom_bar(aes(x = 1:numCompsHeight, y = varCompMat[, 2] / sum(modelHeight[[12]] )- 
                     varCompMat[, 1] / sum(modelHeight[[12]])),
               fill = 'grey',
               stat = 'identity') +
      labs(x = 'Component', y = 'Variance explained') +
      scale_x_continuous(breaks = seq(1, numCompsHeight, by = 1)) +
      theme_classic() + 
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #Plot PLS residuals against GCMS signals under the selected number of
      #components
    plsPlotHeightResi <- ggplot() + 
      geom_point(aes(x = modelHeight$fitted.values[,,numCompsHeight],
                     y = modelHeight$residuals[,,numCompsHeight])) + 
      geom_hline(yintercept = 0) +
      labs(x = 'PLS fitted values (peak height)',
           y = 'Residuals') + 
      theme_classic() + 
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #Remove unused variables
    rm(loopI)
  }
  
  #Output PLS results
  {
    #Combine plot and output PLS model
    pdf(file = file.path('..', 'Figures_R', 'PLS_Nor.pdf'),
        width = 8,
        height = 12)
    
    (plsPlotHeight | modelHeightVar) /
      (modelHeightCof | modelHeightLoa) /
      (plsPlotHeightResi | plot_spacer() )
    
    dev.off()
    
    #Combine the PLS predicted values to the FTIR peak statistics
    tempMat <- cbind(pcaMatHeight, modelHeight$fitted.values[,, numCompsHeight])
    
    #Output PLS predicted values and FTIR statistics together
    write.csv(tempMat, file = file.path('..', 'Output_R', 'PLS_Nor.csv'))
    
    #Remove unused variables
    rm(tempMat)
  }
}
```

### Figures on PLS analysis The printed PDF version is available in
'Figures_R' folder, the dataset used for PLS model (and the PLS fitted
values) is available in 'Output_R' folder

```{r PLS figure, echo = FALSE}
 (plsPlotHeight | modelHeightVar) /
      (modelHeightCof | modelHeightLoa) /
      (plsPlotHeightResi | plot_spacer() )
```

### Malformation analysis The plots on malformation rates, GCMS signals
between UV/Control groups, and FTIR PLS fitted values between UV/Control
groups are available in 'Figures_R' folder

```{r malformation, include = FALSE}
#Read malformation data and combine them with FTIR/GCMS data--------------------
{
  #Read malformation data
  {
    #Read the cutting branch data, and replace all na with 0
    cuttingsRaw <- read.csv(file = 
                              file.path('..', 
                                        'Data_R', 
                                        'Malformations_Cuttings.csv'),
                            row.names = 1)
    
    #Replace all NAs with 0
    cuttingsRaw[is.na(cuttingsRaw)] <- 0
    
    #Extract the batch information
    batchInfoTemp <- strsplit(x = cuttingsRaw[,1], split = '_')
    cuttingsBatchInfo <- cbind(rep('Cuttings', length(batchInfoTemp)),
                               rep('2025', length(batchInfoTemp)),
                               matrix(unlist(batchInfoTemp), ncol = 4, byrow = TRUE))
    colnames(cuttingsBatchInfo) <- c('Sample', 'Year', 'Frame', 'Treatment',
                                     'Plant', 'Duplicate')
    
    #Delete unused variables
    rm(batchInfoTemp)
    
    #Read the cultivar branch data, and replace all na with 0
    cultivarsRaw <- read.csv(file = 
                               file.path('..', 
                                         'Data_R', 
                                         'Malformations_Cultivars.csv'),
                             row.names = 1)
    
    #Replace all NAs with 0
    cultivarsRaw[is.na(cultivarsRaw)] <- 0
    
    #Extract the batch information
    batchInfoTemp <- strsplit(x = cultivarsRaw[,1], split = '_')
    cultivarsBatchInfo <- cbind(rep('Cultivars', length(batchInfoTemp)),
                                matrix(unlist(batchInfoTemp), ncol = 4, byrow = TRUE))
    colnames(cultivarsBatchInfo) <- c('Sample', 'Year', 'Frame', 'Treatment',
                                      'Plant')
    
    #Delete unused variables
    rm(batchInfoTemp)
    
    #Combine counting results in duplicated samples, by summing them up
    tempVec <- unique(cultivarsRaw[, 1])
    
    #Matrix for all malformation data
    tempMat <- matrix(nrow = length(tempVec),
                      ncol = ncol(cultivarsRaw),
                      dimnames = list(c(),
                                      colnames(cultivarsRaw)))
    
    #Matrix for all batch information on malformation data
    tempMat2 <- matrix(nrow = length(tempVec),
                       ncol = ncol(cultivarsBatchInfo),
                       dimnames = list(c(),
                                       colnames(cultivarsBatchInfo)))
    
    #Fill sample label into matrix
    tempMat[, 1] <- tempVec
    
    #Loop to add duplicates together
    for (loopI in 1:nrow(tempMat)) {
      
      #Check  if there are duplicates in current sample
      if (length(which(cultivarsRaw[, 1] == tempMat[loopI, 1])) > 1) {
        
        #Sum up all counting results among duplicates
        tempMat[loopI, 2:ncol(tempMat)] <- apply(cultivarsRaw[
          which(cultivarsRaw[, 1] == tempMat[loopI, 1]), 
          2:ncol(tempMat)
        ], 2, sum)
        
      }else{
        
        #Transfer couting results into matrix
        tempMat[loopI, 2:ncol(tempMat)] <- cultivarsRaw[
          which(cultivarsRaw[, 1] == tempMat[loopI, 1]), 2:ncol(tempMat)
        ]
        
      }
      
      #Record the corresponding batch information
      tempMat2[loopI,] <- cultivarsBatchInfo[which(cultivarsRaw[, 1] == tempMat[loopI, 1])[1],]
      
    }
    
    #Combine batch information with counting results
    cultivarsCom <- cbind(tempMat2, tempMat[, -1])
    
    rm(loopI,
       tempMat,
       tempMat2,
       tempVec)
  }
  
  #Combine malformation data with FTIR and GC-MS data
  {
    #Add batch information to FTIR data
    pcaMatFTIR <- cbind(pcaMatHeight[, c(1:10)],
                        modelHeight$fitted.values[,,numCompsHeight],
                        pcaMatHeight[, c(25:26)])
    
    colnames(pcaMatFTIR)[-1:-10] <- c('PLS', 'GCMS', 'GCMS_SD')
    
    #Vector for recording row index in FTIR spectra matrix
    tempInd <- vector(length = nrow(pcaMatFTIR))
    
    #Combine batch information of cuttings and pot cultivars together
    tempMat <- rbind(cultivarsCom[,1:6],
                     cuttingsBatchInfo)
    
    #Add malformation data to FTIR data
    for (loopI in 1:nrow(pcaMatFTIR)) {
      
      #Check and skip the roof samples
      if (pcaMatFTIR[loopI, 3] == 'ROOF') {
        
        tempInd[loopI] <- NA
        
      }else{
        
        #For all pot cultivars
        if (pcaMatFTIR[loopI, 5] == 'cul') {
          
          if (any((tempMat[, 3] == bcSpec[loopI, 6]) &
                  (tempMat[, 4] == bcSpec[loopI, 7]) &
                  (tempMat[, 5] == bcSpec[loopI, 8]) &
                  (!is.na(as.numeric(tempMat[, 6])))
                  )
          ) {
            
            #Record the row index for FTIR spectra matrix
            tempInd[loopI] <- which(
              (tempMat[, 3] == pcaMatFTIR[loopI, 6]) &
                (tempMat[, 4] == pcaMatFTIR[loopI, 7]) &
                (tempMat[, 5] == 
                   as.character(as.numeric(pcaMatFTIR[loopI, 8]))) &
                (!is.na(as.numeric(tempMat[, 6])))
              )
            
          }else {
            
            tempInd[loopI] <- NA
            
          }
        
        #For all cuttings
        } else if (pcaMatFTIR[loopI, 5] == 'cut') {
          
          #Check if current FTIR sample has malformation counting results
          if (any((tempMat[, 3] == pcaMatFTIR[loopI, 6]) &
                  (tempMat[, 4] == pcaMatFTIR[loopI, 7]) &
                  (tempMat[, 5] == as.character(as.numeric(pcaMatFTIR[loopI, 8]))) &
                  (tempMat[, 6] == pcaMatFTIR[loopI, 9])
          )
          ) {
            
            #Record the row index for FTIR spectra matrix
            tempInd[loopI] <- which(
              (tempMat[, 3] == pcaMatFTIR[loopI, 6]) &
                (tempMat[, 4] == pcaMatFTIR[loopI, 7]) &
                (tempMat[, 5] == 
                   as.character(as.numeric(pcaMatFTIR[loopI, 8]))) &
                (tempMat[, 6] == pcaMatFTIR[loopI, 9]))
            
          }else {
            
            tempInd[loopI] <- NA
            
          }
          
        } else {
          
          tempInd[loopI] <- NA
          
        }
      }
    }
    
    #Combine all counting results together from both cuttings and pot cultivars
    tempMat <- rbind(cultivarsCom[, -1:-5],
                     as.matrix(cuttingsRaw[, -1]))
    
    #Add malformation counting results to FTIR and GCMS matrix
    pcaMatFTIRGCMal <- cbind(pcaMatFTIR, tempMat[tempInd, c(1, 14)])
    
    #Adjest value type to numeric
    pcaMatFTIRGCMal[, 11:15] <- apply(pcaMatFTIRGCMal[, 11:15], c(1,2), as.numeric)
    
    #Calculate the malformation ratio in each sample, and add them to the whole
      #matrix
    pcaMatFTIRGCMalFull <- cbind(pcaMatFTIRGCMal, 
                                 Mal_Ratio = 
                                   (pcaMatFTIRGCMal[,ncol(pcaMatFTIRGCMal)] - pcaMatFTIRGCMal[,ncol(pcaMatFTIRGCMal) - 1]) / 
                                   pcaMatFTIRGCMal[,ncol(pcaMatFTIRGCMal)])
    
    #Remove samples that do not have malformation counting results
    pcaMatFTIRGCMalFull <- pcaMatFTIRGCMalFull[which(!is.na(pcaMatFTIRGCMalFull[, 6])),]
    
    #Calculate the mean and sd of FTIR duplicates
    #Get unique batch information
    tempMat <- unique(pcaMatFTIRGCMalFull[, 1:8])
    
    #Matrix for recording mean and sd of FTIR duplicates
    tempMat2 <- matrix(nrow = nrow(tempMat),
                       ncol = ncol(pcaMatFTIRGCMalFull) - 1,
                       dimnames = list(c(),
                                       c(colnames(pcaMatFTIRGCMalFull)[-9:-10], 
                                         paste(colnames(pcaMatFTIRGCMalFull)[11], '_SD')
                                       )
                       )
    )
    
    tempMat2[,1:8] <- as.matrix(tempMat)
    
    #Calculate the mean FTIR signals
    for (loopI in 1:nrow(tempMat2)) {
      
      #Row indexes for all duplicates of current sample
      tempVec <- which((pcaMatFTIRGCMalFull[, 5] == tempMat2[loopI, 5]) &
                         (pcaMatFTIRGCMalFull[, 6] == tempMat2[loopI, 6]) &
                         (pcaMatFTIRGCMalFull[, 7] == tempMat2[loopI, 7]) &
                         (as.numeric(pcaMatFTIRGCMalFull[, 8]) == as.numeric(tempMat2[loopI, 8]))
      )
      
      #Check whetehr there are duplicates
      if (length(tempVec) > 1) {
        
        #Calculate mean and sd of duplicates
        tempMat2[loopI, 9:14] <- apply(pcaMatFTIRGCMalFull[tempVec, 11:16],
                                       2,
                                       mean)
        tempMat2[loopI, 15] <- sd(pcaMatFTIRGCMalFull[tempVec, 11])
      }else {
        
        #Copy the signals to corresponding position
        tempMat2[loopI, 9:14] <- as.numeric(pcaMatFTIRGCMalFull[tempVec, 11:16])
        #Sd is assigned to 0 for no-duplicate sample
        tempMat2[loopI, 15] <- 0
        
      }
    }
    
    #Transfer the combined matrix to named matrix
    pcaMatFTIRGCMalFullMean <- tempMat2
    pcaMatFTIRGCMalFullMean <- as.data.frame(pcaMatFTIRGCMalFullMean)
    pcaMatFTIRGCMalFullMean[, 9:15] <- apply(pcaMatFTIRGCMalFullMean[, 9:15], 
                                             c(1,2),
                                             as.numeric)
    
    rm(tempInd,
       tempMat,
       tempMat2,
       loopI
    )
  }
}

#Plot on malformation, GCMS and FTIR data---------------------------------------
{
  #Box plot on FTIR and GCMS data
  {
    #FTIR
    #PLS
    boxFTIRPLS <- ggplot(pcaMatFTIRGCMalFull, 
                                aes(y =  modelHeight$fitted.values[
                                  ,,numCompsHeight],
                                    x = Treatment)) +
      geom_boxplot(outliers = FALSE) +
      geom_jitter(aes(color = Materials), shape=16, position=position_jitter(0.2)) +
      labs(y = 'PLS predicted values') +
      scale_color_manual(values = c('cul' = 'red', 'cut' = 'blue'),
                         labels = c('Cultivars', 'Cuttings')) +
      theme_classic() +
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
            legend.position = c(0.9, 0.9)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #Sub-divided by cultivars and cuttings
    #PLS
    boxSubFTIRPLS <- ggplot(pcaMatFTIRGCMalFull, 
                                   aes(y = modelHeight$fitted.values[
                                     ,,numCompsHeight],
                                       x = Treatment,
                                       fill = Materials)) +
      geom_boxplot(outliers = FALSE) +
      geom_jitter(aes(fill = Materials), shape=21, position=position_jitterdodge(0.2),
                  color = 'black',
                  stroke = 1,
                  size = 2) +
      labs(y = 'PLS predicted values') +
      scale_fill_manual(values = c('cul' = 'red', 'cut' = 'blue'),
                        labels = c('Cultivars', 'Cuttings')) +
      theme_classic() +
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
            legend.position = c(0.9, 0.9)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #Extract the group information between cuttings and pot cultivars
    GCDataSub <- is.na(GCDataMean[, 6])
    GCDataSub <- replace(GCDataSub, GCDataSub == FALSE,
                         c('Cuttings'))
    GCDataSub <- replace(GCDataSub, GCDataSub == TRUE,
                         c('Cultivars'))
    
    #GC-MS
    boxGCMS <- ggplot(GCDataMean, 
                      aes(y = calibrated,
                          x = Treatment)) +
      geom_boxplot(outliers = FALSE) +
      geom_jitter(aes(color = GCDataSub), shape=16, position=position_jitter(0.2)) +
      scale_color_manual(values = c('Cuttings' = 'blue', 'Cultivars' = 'red')
      ) +
      labs(y = 'UV GC-MS (ng/100gr)', color = 'Materials') +
      theme_classic() +
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
            legend.position = c(0.9, 0.9)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #GC-MS subgroup
    boxSubGCMS <- ggplot(GCDataMean, 
                         aes(y = calibrated,
                             x = Treatment,
                             fill = GCDataSub)) +
      geom_boxplot(outliers = FALSE) +
      geom_jitter(aes(fill = GCDataSub) , shape=21, position=position_jitterdodge(0.2),
                  color = 'black',
                  stroke = 1,
                  size = 2) +
      scale_color_manual(values = c('Cuttings' = 'blue', 'Cultivars' = 'red')
      ) +
      scale_fill_manual(values = c('Cuttings' = 'blue', 'Cultivars' = 'red'),
                        name = 'Materials') +
      labs(y = 'UV GC-MS (ng/100gr)', color = 'Materials') +
      theme_classic() +
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
            legend.position = c(0.9, 0.9)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
  }
  
  #Scatter plot referring to malformation rate
  {
    #Create a matrix recording all signifcance results bettwen diffent subgroups
    cuttingsPlantsMat <- matrix(nrow = length(unique(
      pcaMatFTIRGCMalFullMean[which(pcaMatFTIRGCMalFullMean[, 5] == 'cut'),
                              8
      ]
    )
    ),
    ncol = 17,
    dimnames = list(c(),
                    c('Plant', 'UV_Mal', 
                      'CON_Mal',
                      'GCMS',
                      'PLS',
                      'GCMS_UV',
                      'GCMS_CON',
                      'PLS_UV',
                      'PLS_CON',
                      'UV_Mal_sd',
                      'CON_Mal_sd',
                      'GCMS_sd',
                      'PLS_sd',
                      'GCMS_UV_sd',
                      'GCMS_CON_sd',
                      'PLS_UV_sd',
                      'PLS_CON_sd')))
    
    #Fill batch information into the matrix
    cuttingsPlantsMat[, 1] <- sort(as.numeric(unique(
      pcaMatFTIRGCMalFullMean[which(pcaMatFTIRGCMalFullMean[, 5] == 'cut'),
                              8
      ])), decreasing = F)
    
    #Row indexes for all cuttings in the general matrix
    tempMat <-  pcaMatFTIRGCMalFullMean[
      which(pcaMatFTIRGCMalFullMean[, 5] == 'cut'),]

    #Loop to fill the matrix
    for (loopI in cuttingsPlantsMat[, 1]) {
      
      #Mean of malformation rates in all UV groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 2] <- mean(tempMat[
        which((tempMat[, 7] == 'UV' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 14],
                                          na.rm = TRUE)
      
      #Mean of malformation rates in all Control groups, and sub grouped by 
        #plants
      cuttingsPlantsMat[loopI, 3] <- mean(tempMat[
        which((tempMat[, 7] == 'CON' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 14],
                                          na.rm = TRUE)
      
      #Mean of GCMS signals sub grouped by plants
      cuttingsPlantsMat[loopI, 4] <- mean(tempMat[
        which((as.numeric(tempMat[, 8]) == loopI)), 10],
                                          na.rm = TRUE)
      
      #Mean of PLS predicted values sub grouped by plants
      cuttingsPlantsMat[loopI, 5] <- mean(tempMat[
        which((as.numeric(tempMat[, 8]) == loopI)), 9],
                                          na.rm = TRUE)
      
      #Mean of GCMS signals in all UV groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 6] <- mean(tempMat[
        which(((tempMat[, 7] == 'UV' ) & 
                 as.numeric(tempMat[, 8]) == loopI)), 10],
                                          na.rm = TRUE)
      
      #Mean of GCMS signals in all Control groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 7] <- mean(tempMat[
        which((tempMat[, 7] == 'CON' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 10],
                                           na.rm = TRUE)
      
      #Mean of PLS predicted values in all UV groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 8] <- mean(tempMat[
        which((tempMat[, 7] == 'UV' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 9],
                                           na.rm = TRUE)
      
      #Mean of PLS predicted values in all Control groups, and sub grouped by 
        #plants
      cuttingsPlantsMat[loopI, 9] <- mean(tempMat[
        which((tempMat[, 7] == 'CON' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 9],
                                           na.rm = TRUE)

      #Sd of malformation rates in all UV groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 10] <- sd(tempMat[
        which((tempMat[, 7] == 'UV' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 14],
                                         na.rm = TRUE)
      
      #Sd of malformation rates in all Control groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 11] <- sd(tempMat[
        which((tempMat[, 7] == 'CON' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 14],
                                         na.rm = TRUE)
      
      #Sd of GCMS signals sub grouped by plants
      cuttingsPlantsMat[loopI, 12] <- sd(tempMat[
        which((as.numeric(tempMat[, 8]) == loopI)), 10],
                                         na.rm = TRUE)
      
      #Sd of PLS predicted values sub grouped by plants
      cuttingsPlantsMat[loopI, 13] <- sd(tempMat[
        which((as.numeric(tempMat[, 8]) == loopI)), 9],
                                         na.rm = TRUE)
      
      #Sd of GCMS signals in all UV groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 14] <- sd(tempMat[
        which(((tempMat[, 7] == 'UV' ) & 
                 as.numeric(tempMat[, 8]) == loopI)), 10],
                                         na.rm = TRUE)
      
      #Sd of GCMS signals in all Control groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 15] <- sd(tempMat[
        which((tempMat[, 7] == 'CON' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 10],
                                         na.rm = TRUE)
      
      #Sd of PLS predicted values in all UV groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 16] <- sd(tempMat[
        which((tempMat[, 7] == 'UV' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 9],
                                         na.rm = TRUE)
      
      #Sd of PLS predicted values in all Control groups, and sub grouped by plants
      cuttingsPlantsMat[loopI, 17] <- sd(tempMat[
        which((tempMat[, 7] == 'CON' ) & 
                (as.numeric(tempMat[, 8]) == loopI)), 9],
                                         na.rm = TRUE)
    }
    
    #Transfer matrix into dataframe
    cuttingsPlantsMat <- as.data.frame(cuttingsPlantsMat)
    cuttingsPlantsMat[, 1] <- as.character(cuttingsPlantsMat[, 1])
    
    #Calculate malformation rates in pot cultivars
    tempVec <- (cultivarsRaw[, 15] - cultivarsRaw[, 2])/ cultivarsRaw[, 2]
    cultivarMean <- vector(length = 4)
    cultivarMean[1] <- mean(tempVec[which(cultivarsBatchInfo[, 4] == 'UV')])
    cultivarMean[2] <- sd(tempVec[which(cultivarsBatchInfo[, 4] == 'UV')])
    cultivarMean[3] <- mean(tempVec[which(cultivarsBatchInfo[, 4] == 'CON')])
    cultivarMean[4] <- sd(tempVec[which(cultivarsBatchInfo[, 4] == 'CON')])
    
    #Point plot to compare ratios between con and uv treatment, on different plants
    cuttingsPlantsPlot <-
      ggplot() +
      geom_point(data = cuttingsPlantsMat,
                 aes(x = CON_Mal, y = UV_Mal, col = Plant)) +
      geom_point(aes(x = cultivarMean[3] , y = cultivarMean[1], 
                     color = 'Cultivars')) +
      geom_errorbar(data = cuttingsPlantsMat,
                    aes(x = CON_Mal, y = UV_Mal, col = Plant,
                        ymin = UV_Mal - UV_Mal_sd,
                        ymax = UV_Mal + UV_Mal_sd),
                    linewidth = 1,
                    width = 0.01) +
      geom_errorbar(data = cuttingsPlantsMat,
                    aes(x = CON_Mal, y = UV_Mal, col = Plant,
                        xmin = CON_Mal - CON_Mal_sd,
                        xmax = CON_Mal + CON_Mal_sd),
                    linewidth = 1,
                    width = 0.01,
                    orientation = 'y') +
      geom_errorbar(aes(x = cultivarMean[3] , y = cultivarMean[1], color = 'Cultivars',
                        ymin = cultivarMean[1] - cultivarMean[2],
                        ymax = cultivarMean[1] + cultivarMean[2]),
                    linewidth = 1,
                    width = 0.005) +
      geom_errorbar(aes(x = cultivarMean[3] , y = cultivarMean[1], color = 'Cultivars',
                        xmin = cultivarMean[3] - cultivarMean[4],
                        xmax = cultivarMean[3] + cultivarMean[4]),
                    linewidth = 1,
                    width = 0.005,
                    orientation = 'y') +
      geom_abline (slope=1, linetype = "dashed", color="grey") +
      scale_x_continuous(limits = c(-0.05, 0.15), name = 'CON malformation percentage',
                         breaks = seq(0, 0.15, by = 0.05)) +
      scale_y_continuous(limits = c(-0.05, 0.25), name = 'UV malformation percentage',
                         breaks = seq(0, 0.25, by = 0.05)) +
      scale_color_manual(name = 'Plant',
                         values = c('1' = brewer.pal(6, 'Set2')[1],
                                    '2' = brewer.pal(6, 'Set2')[2],
                                    '3' = brewer.pal(6, 'Set2')[3],
                                    '4' = brewer.pal(6, 'Set2')[4],
                                    '5' = brewer.pal(6, 'Set2')[5],
                                    '6' = brewer.pal(6, 'Set2')[6],
                                    'Cultivars' = 'black')) +
      theme_classic() +
      theme(
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
        axis.title.x = element_text(size = 20, family = 'arial'),
        axis.text.x = element_text(size = 16, family = 'arial'),
        
        axis.text.y = element_text(size = 20, family = 'arial'),
        axis.title.y = element_text(size = 16, family = 'arial'),
        legend.position = "inside",
        legend.position.inside = c(0.8, 0.2),
        legend.text = element_text(size = 14, family = 'arial')
      )
  }
  
  #Bar plot - plant divided
  {
    #GCMS
    cuttingsPlantsMatReOr1 <- cbind(
      melt(cuttingsPlantsMat[,
                             which(
                               (colnames(cuttingsPlantsMat) == 'Plant') |
                                 (colnames(cuttingsPlantsMat) == 'GCMS_UV') |
                                 (colnames(cuttingsPlantsMat) == 'GCMS_CON'))], 
           id = 'Plant',
           value.name = 'Signal',
           variable.name = 'GCMS'),
      melt(cuttingsPlantsMat[,
                             which((colnames(cuttingsPlantsMat) == 'Plant') |
                                     (colnames(cuttingsPlantsMat) == 
                                        'GCMS_UV_sd') | 
                                     (colnames(cuttingsPlantsMat) == 
                                        'GCMS_CON_sd'))], 
                                         id = 'Plant',
                                         value.name = 'SD',
                                         variable.name = 'GCMS_SD')[, -1]
    )
    
    #Bar plot on GC-MS
    barPlotPlantGCMS <- ggplot() +
      geom_boxplot(data = GCDataMean[c(1:65, 67:108),],
                   aes(x = as.character(TreeNumber),
                       y = calibrated,
                       fill = factor(Treatment, levels = c('UV', 'CON'))),
                   # stat = 'identity',
                   # position="dodge",
                   outliers = FALSE,
                   alpha = 0.6) +
      geom_boxplot(data = GCDataMean[109:133,],
                   aes(x = 'Cultivars',
                       y = calibrated,
                       fill = factor(Treatment, levels = c('UV', 'CON'))),
                   # stat = 'identity',
                   # position="dodge",
                   outliers = FALSE,
                   alpha = 0.6) +
      geom_jitter(data = GCDataMean[109:133,],
                  aes(x = 'Cultivars',
                      y = calibrated,
                      fill = factor(Treatment, levels = c('UV', 'CON'))), shape=21, 
                  position = position_jitterdodge(0.6),
                  color = 'black',
                  stroke = 1,
                  size = 4,
                  alpha = 0.6) +
      geom_jitter(data = GCDataMean[c(1:65, 67:108),],
                  aes(x = as.character(TreeNumber),
                      y = calibrated,
                      fill = factor(Treatment, levels = c('UV', 'CON'))), shape=21, 
                  position = position_jitterdodge(0.6),
                  color = 'black',
                  stroke = 1,
                  size = 4,
                  alpha = 0.6) +
      scale_y_continuous(name = 'UV GC-MS (ng/100gr)',
                         # breaks = seq(0, 0.5, by = 0.1)
      ) +
      scale_fill_manual(values = c('GCMS_UV' = 'red',
                                   'GCMS_CON' = 'blue',
                                   'UV' = 'red',
                                   'CON' = 'blue'),
                        labels = c('UV', 'CON', 'UV', 'CON'),
                        name = 'Treatment') +
      labs(x = 'Tree') +
      theme_classic() +
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
            legend.position = c(0.8, 0.9),
            legend.key = element_rect(fill = NA)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
    
    #FTIR
    cuttingsPlantsMatReOr4 <- cbind(
      melt(cuttingsPlantsMat[,
                             which((colnames(cuttingsPlantsMat) == 'Plant') |
                                     (colnames(cuttingsPlantsMat) == 
                                        'PLS_UV') |
                                     (colnames(cuttingsPlantsMat) == 
                                        'PLS_CON'))], 
           id = 'Plant',
           value.name = 'Signal',
           ariable.name = 'PLS'),
      melt(cuttingsPlantsMat[,
                             which((colnames(cuttingsPlantsMat) == 'Plant') |
                                     (colnames(cuttingsPlantsMat) == 
                                        'PLS_UV_sd') | 
                                     (colnames(cuttingsPlantsMat) == 
                                        'PLS_CON_sd'))],
           id = 'Plant',
           value.name = 'SD',
           variable.name = 'PLS_SD')[, -1]
    )
    
    barPlotPlantPLS <- ggplot() +
      geom_boxplot(data = pcaMatFTIRGCMalFull[which(pcaMatFTIRGCMalFull[, 5] == 'cut'),],
                   aes(x = factor(as.numeric(Sample_ID), levels = c(1:6)),
                       y = PLS,
                       fill = factor(Treatment, levels = c('UV', 'CON'))),
                   # stat = 'identity',
                   # position="dodge",
                   outliers = FALSE,
                   alpha = 0.6) +
      geom_boxplot(data = pcaMatFTIRGCMalFull[which(pcaMatFTIRGCMalFull[, 5] == 'cul'),],
                   aes(x = 'Cultivars',
                       y = PLS,
                       fill = factor(Treatment, levels = c('UV', 'CON'))),
                   # stat = 'identity',
                   # position="dodge",
                   outliers = FALSE,
                   alpha = 0.6) +
      geom_jitter(data = pcaMatFTIRGCMalFull[which(pcaMatFTIRGCMalFull[, 5] == 'cut'),],
                  aes(x = factor(as.numeric(Sample_ID), levels = c(1:6)),
                      y = PLS,
                      fill = factor(Treatment, levels = c('UV', 'CON'))), shape=21, 
                  position = position_jitterdodge(0.6),
                  color = 'black',
                  stroke = 1,
                  size = 4,
                  alpha = 0.6) +
      geom_jitter(data = pcaMatFTIRGCMalFull[which(pcaMatFTIRGCMalFull[, 5] == 'cul'),],
                  aes(x = 'Cultivars',
                      y = PLS,
                      fill = factor(Treatment, levels = c('UV', 'CON'))), shape=21, 
                  position = position_jitterdodge(0.6),
                  color = 'black',
                  stroke = 1,
                  size = 4,
                  alpha = 0.6) +
      scale_y_continuous(name = 'PLS fitted value',
                         # breaks = seq(0, 3, by = 1)
      ) +
      scale_x_discrete(name = 'Plant') +
      scale_fill_manual(values = c('PLS_UV' = 'red',
                                   'PLS_CON' = 'blue',
                                   'UV' = 'red',
                                   'CON' = 'blue'),
                        labels = c('UV', 'CON', 'UV', 'CON'),
                        name = 'Treatment') +
      labs(x = 'Tree') +
      theme_classic() +
      theme(plot.title = element_blank(), 
            axis.title.y = element_text(size = 14, family = 'arial'),
            axis.text.y = element_text(size = 12, family = 'arial'),
            axis.title.x = element_text(size = 14, family = 'arial'),
            axis.text.x = element_text(size = 12, family = 'arial'),
            panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
            legend.position = c(0.8, 0.9),
            legend.background = element_rect(fill = NA)
            # axis.line.y = element_blank(),
            # axis.text.x = element_blank(), 
            # axis.line.x = element_blank(),
            # axis.title.x = element_blank(),
            # axis.ticks.x = element_blank()
      )
  }
  
  #Output figures
  {
    #FTIR boxplot
    pdf(file = file.path('..', 'Figures_R', 'FTIR.pdf'),
        width = 12, # The width of the plot in inches
        height = 12)
    
    (boxFTIRPLS | boxSubFTIRPLS) /
      (barPlotPlantPLS | barPlotPlantPLS)
    
    dev.off()
    
    #GCMS boxplot
    pdf(file = file.path('..', 'Figures_R', 'GCMS.pdf'),
        width = 12, # The width of the plot in inches
        height = 12)
    
    (boxGCMS | boxSubGCMS) /
      (barPlotPlantGCMS | barPlotPlantGCMS)
    
    dev.off()
    
    #Malformation
    pdf(file = file.path('..', 'Figures_R', 'Malformation.pdf'),
        width = 12, # The width of the plot in inches
        height = 12)
    
    cuttingsPlantsPlot
    
    dev.off()
    
    #GCMS VS FTIR
    pdf(file =  file.path('..', 'Figures_R', 'GCMS_FTIR.pdf'),
        width = 12, # The width of the plot in inches
        height = 6)
    
    barPlotPlantPLS | barPlotPlantGCMS
    
    dev.off()
    
  }
}
```

### Figures on Malformation rate

```{r malformation rate plot, echo = TRUE}
cuttingsPlantsPlot
```

### Figures on GCMS and FTIR signals

```{r GCMS & FTIR plot, echo = FALSE}
 (boxGCMS | boxSubGCMS) /
      (barPlotPlantGCMS | barPlotPlantGCMS)

(boxFTIRPLS | boxSubFTIRPLS) /
      (barPlotPlantPLS | barPlotPlantPLS)
```

### ANOVA analysis This analysis is to test the significance of GCMS/FTIR
signals between different subgroups (UV/CONTROL, plant 1-6,
cuttings/cultivars). Student-t test is used. The table recording all
significane test results are available in 'Output_R' folder.

```{r statistic analysis, include = FALSE}
#Statistic analysis on peak signals and malformation rates----------------------
{
  #GCMS
  {
    #GCMS whole
    tempVec1 <- which(GCDataMean[, 4] == 'UV')
    tempVec2 <- which(GCDataMean[, 4] == 'CON')
    
    stuTMatGCMS <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                          ncol = 2,
                          dimnames = list(c(),
                                          c('UV', 'CON')))
    
    stuTMatGCMS[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMS[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMS <- t.test(stuTMatGCMS[, 1], stuTMatGCMS[, 2],
                          var.equal = FALSE, alternative = 'greater')
    
    #GCMS whole subgroup UV
    tempVec1 <- which((GCDataMean[, 4] == 'UV') &
                        (is.na(GCDataMean[, 6])))
    tempVec2 <- which(GCDataMean[, 4] == 'UV'&
                        (!is.na(GCDataMean[, 6])))
    
    stuTMatGCMSSubUV <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                               ncol = 2,
                               dimnames = list(c(),
                                               c('Cul', 'Cut')))
    
    stuTMatGCMSSubUV[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSSubUV[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSSubUV <- t.test(stuTMatGCMSSubUV[, 1], stuTMatGCMSSubUV[, 2],
                               var.equal = FALSE, alternative = 'greater')
    
    #GCMS whole subgroup CON
    tempVec1 <- which((GCDataMean[, 4] == 'CON') &
                        (is.na(GCDataMean[, 6])))
    tempVec2 <- which(GCDataMean[, 4] == 'CON'&
                        (!is.na(GCDataMean[, 6])))
    
    stuTMatGCMSSubCON <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                ncol = 2,
                                dimnames = list(c(),
                                                c('Cul', 'Cut')))
    
    stuTMatGCMSSubCON[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSSubCON[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSSubCON <- t.test(stuTMatGCMSSubCON[, 1], stuTMatGCMSSubCON[, 2],
                                var.equal = FALSE, alternative = 'greater')
    
    #GCMS cul and cut subgroup
    tempVec1 <- which((GCDataMean[, 4] == 'UV')&
                        is.na(GCDataMean[, 6]))
    tempVec2 <- which((GCDataMean[, 4] == 'CON')&
                        is.na(GCDataMean[, 6]))
    
    stuTMatGCMSCul <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                             ncol = 2,
                             dimnames = list(c(),
                                             c('UV', 'CON')))
    
    stuTMatGCMSCul[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSCul[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSCul <- t.test(stuTMatGCMSCul[, 1], stuTMatGCMSCul[, 2],
                             var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((GCDataMean[, 4] == 'UV')&
                        !is.na(GCDataMean[, 6]))
    tempVec2 <- which((GCDataMean[, 4] == 'CON')&
                        !is.na(GCDataMean[, 6]))
    
    stuTMatGCMSCut <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                             ncol = 2,
                             dimnames = list(c(),
                                             c('UV', 'CON')))
    
    stuTMatGCMSCut[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSCut[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSCut <- t.test(stuTMatGCMSCut[, 1], stuTMatGCMSCut[, 2],
                             var.equal = FALSE, alternative = 'greater')
    
    #GCMS Plant subgroup
    tempVec1 <- which((GCDataMean[, 4] == 'UV')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '1'))
    tempVec2 <- which((GCDataMean[, 4] == 'CON')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '1'))
    
    stuTMatGCMSPlant1 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                ncol = 2,
                                dimnames = list(c(),
                                                c('UV', 'CON')))
    
    stuTMatGCMSPlant1[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSPlant1[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSPlant1 <- t.test(stuTMatGCMSPlant1[, 1], stuTMatGCMSPlant1[, 2],
                                var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((GCDataMean[, 4] == 'UV')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '2'))
    tempVec2 <- which((GCDataMean[, 4] == 'CON')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '2'))
    
    stuTMatGCMSPlant2 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                ncol = 2,
                                dimnames = list(c(),
                                                c('UV', 'CON')))
    
    stuTMatGCMSPlant2[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSPlant2[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSPlant2 <- t.test(stuTMatGCMSPlant2[, 1], stuTMatGCMSPlant2[, 2],
                                var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((GCDataMean[, 4] == 'UV')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '3'))
    tempVec2 <- which((GCDataMean[, 4] == 'CON')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '3'))
    
    stuTMatGCMSPlant3 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                ncol = 2,
                                dimnames = list(c(),
                                                c('UV', 'CON')))
    
    stuTMatGCMSPlant3[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSPlant3[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSPlant3 <- t.test(stuTMatGCMSPlant3[, 1], stuTMatGCMSPlant3[, 2],
                                var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((GCDataMean[, 4] == 'UV')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '4'))
    tempVec2 <- which((GCDataMean[, 4] == 'CON')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '4'))
    
    stuTMatGCMSPlant4 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                ncol = 2,
                                dimnames = list(c(),
                                                c('UV', 'CON')))
    
    stuTMatGCMSPlant4[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSPlant4[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSPlant4 <- t.test(stuTMatGCMSPlant4[, 1], stuTMatGCMSPlant4[, 2],
                                var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((GCDataMean[, 4] == 'UV')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '5'))
    tempVec2 <- which((GCDataMean[, 4] == 'CON')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '5'))
    
    stuTMatGCMSPlant5 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                ncol = 2,
                                dimnames = list(c(),
                                                c('UV', 'CON')))
    
    stuTMatGCMSPlant5[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSPlant5[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSPlant5 <- t.test(stuTMatGCMSPlant5[, 1], stuTMatGCMSPlant5[, 2],
                                var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((GCDataMean[, 4] == 'UV')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '6'))
    tempVec2 <- which((GCDataMean[, 4] == 'CON')&
                        !is.na(GCDataMean[, 6]) &
                        (GCDataMean[, 5] == '6'))
    
    stuTMatGCMSPlant6 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                ncol = 2,
                                dimnames = list(c(),
                                                c('UV', 'CON')))
    
    stuTMatGCMSPlant6[1:length(tempVec1), 1] <- GCDataMean[tempVec1, 1]
    stuTMatGCMSPlant6[1:length(tempVec2), 2] <- GCDataMean[tempVec2, 1]
    
    stuTResGCMSPlant6 <- t.test(stuTMatGCMSPlant6[, 1], stuTMatGCMSPlant6[, 2],
                                var.equal = FALSE, alternative = 'greater')
  }
  
  #FTIR PLS
  {
    #FTIRPLS whole
    tempVec1 <- which(pcaMatFTIRGCMalFull[, 7] == 'UV')
    tempVec2 <- which(pcaMatFTIRGCMalFull[, 7] == 'CON')
    
    stuTMatFTIRPLS <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                     ncol = 2,
                                     dimnames = list(c(),
                                                     c('UV', 'CON')))
    
    stuTMatFTIRPLS[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLS[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLS <- t.test(stuTMatFTIRPLS[, 1], stuTMatFTIRPLS[, 2],
                                     var.equal = FALSE, alternative = 'greater')
    
    #FTIRPLS whole subgroup UV
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV') &
                        (pcaMatFTIRGCMalFull[, 5] == 'cul'))
    tempVec2 <- which(pcaMatFTIRGCMalFull[, 7] == 'UV'&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut'))
    
    stuTMatFTIRPLSSubUV <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                          ncol = 2,
                                          dimnames = list(c(),
                                                          c('Cul', 'Cut')))
    
    stuTMatFTIRPLSSubUV[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSSubUV[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSSubUV <- t.test(stuTMatFTIRPLSSubUV[, 1], stuTMatFTIRPLSSubUV[, 2],
                                          var.equal = FALSE, alternative = 'greater')
    
    #FTIRPLS whole subgroup CON
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON') &
                        (pcaMatFTIRGCMalFull[, 5] == 'cul'))
    tempVec2 <- which(pcaMatFTIRGCMalFull[, 7] == 'CON'&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut'))
    
    stuTMatFTIRPLSSubCON <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                           ncol = 2,
                                           dimnames = list(c(),
                                                           c('Cul', 'Cut')))
    
    stuTMatFTIRPLSSubCON[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSSubCON[1:length(tempVec2), 2] <-pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSSubCON <- t.test(stuTMatFTIRPLSSubCON[, 1], stuTMatFTIRPLSSubCON[, 2],
                                           var.equal = FALSE, alternative = 'greater')
    
    #FTIRPLS cul and cut subgroup
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cul'))
    tempVec2 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cul'))
    
    stuTMatFTIRPLSCul <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                        ncol = 2,
                                        dimnames = list(c(),
                                                        c('UV', 'CON')))
    
    stuTMatFTIRPLSCul[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSCul[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSCul <- t.test(stuTMatFTIRPLSCul[, 1], stuTMatFTIRPLSCul[, 2],
                                        var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut'))
    tempVec2 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut'))
    
    stuTMatFTIRPLSCut <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                        ncol = 2,
                                        dimnames = list(c(),
                                                        c('UV', 'CON')))
    
    stuTMatFTIRPLSCut[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSCut[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSCut <- t.test(stuTMatFTIRPLSCut[, 1], stuTMatFTIRPLSCut[, 2],
                                        var.equal = FALSE, alternative = 'greater')
    
    #FTIRPLS Plant subgroup
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '1'))
    tempVec2 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '1'))
    
    stuTMatFTIRPLSPlant1 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                           ncol = 2,
                                           dimnames = list(c(),
                                                           c('UV', 'CON')))
    
    stuTMatFTIRPLSPlant1[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSPlant1[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSPlant1 <- t.test(stuTMatFTIRPLSPlant1[, 1], stuTMatFTIRPLSPlant1[, 2],
                                           var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '2'))
    tempVec2 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '2'))
    
    stuTMatFTIRPLSPlant2 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                           ncol = 2,
                                           dimnames = list(c(),
                                                           c('UV', 'CON')))
    
    stuTMatFTIRPLSPlant2[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSPlant2[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSPlant2 <- t.test(stuTMatFTIRPLSPlant2[, 1], stuTMatFTIRPLSPlant2[, 2],
                                           var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '3'))
    tempVec2 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '3'))
    
    stuTMatFTIRPLSPlant3 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                           ncol = 2,
                                           dimnames = list(c(),
                                                           c('UV', 'CON')))
    
    stuTMatFTIRPLSPlant3[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSPlant3[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSPlant3 <- t.test(stuTMatFTIRPLSPlant3[, 1], stuTMatFTIRPLSPlant3[, 2],
                                           var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '4'))
    tempVec2 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '4'))
    
    stuTMatFTIRPLSPlant4 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                           ncol = 2,
                                           dimnames = list(c(),
                                                           c('UV', 'CON')))
    
    stuTMatFTIRPLSPlant4[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSPlant4[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSPlant4 <- t.test(stuTMatFTIRPLSPlant4[, 1], stuTMatFTIRPLSPlant4[, 2],
                                           var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '5'))
    tempVec2 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '5'))
    
    stuTMatFTIRPLSPlant5 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                           ncol = 2,
                                           dimnames = list(c(),
                                                           c('UV', 'CON')))
    
    stuTMatFTIRPLSPlant5[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSPlant5[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSPlant5 <- t.test(stuTMatFTIRPLSPlant5[, 1], stuTMatFTIRPLSPlant5[, 2],
                                           var.equal = FALSE, alternative = 'greater')
    
    tempVec1 <- which((pcaMatFTIRGCMalFull[, 7] == 'UV')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '6'))
    tempVec2 <- which((pcaMatFTIRGCMalFull[, 7] == 'CON')&
                        (pcaMatFTIRGCMalFull[, 5] == 'cut') &
                        (as.character(as.numeric(pcaMatFTIRGCMalFull[, 8])) == '6'))
    
    stuTMatFTIRPLSPlant6 <- matrix(nrow = max(length(tempVec1), length(tempVec2)),
                                           ncol = 2,
                                           dimnames = list(c(),
                                                           c('UV', 'CON')))
    
    stuTMatFTIRPLSPlant6[1:length(tempVec1), 1] <- pcaMatFTIRGCMalFull[tempVec1, 11]
    stuTMatFTIRPLSPlant6[1:length(tempVec2), 2] <- pcaMatFTIRGCMalFull[tempVec2, 11]
    
    stuTResFTIRPLSPlant6 <- t.test(stuTMatFTIRPLSPlant6[, 1], stuTMatFTIRPLSPlant6[, 2],
                                           var.equal = FALSE, alternative = 'greater')
  }
  
  #Combine all ANOVA results
  {
    #Create matrix for recording all significance test results
    stuTMatAll <- matrix(nrow = 22,
                         ncol = 5,
                         dimnames = list(c(),
                                         c('Group',
                                           'UV_Mean',
                                           'CON_Mean',
                                           'Sig',
                                           'Mark')))
    
    stuTMatAll[, 1] <- c(paste(rep('GCMS', 11), c('', 'SubUV', 'SubCON',
                                                  'Cut', 'Cul',
                                                  paste('Plant', 1:6, sep = ''
                                                  )
    ),
    sep = ''),
    paste(rep('FTIRPLS', 11), c('', 'SubUV', 'SubCON',
                                        'Cut', 'Cul',
                                        paste('Plant', 1:6, sep = ''
                                        )
    ),
    sep = '')
    )
    
    for (loopI in 1:nrow(stuTMatAll)) {
      
      stuTMatAll[loopI, 2:3] <- get(str_replace_all(paste('stuTRes', stuTMatAll[loopI, 1],
                                                          sep = '', collapse = NULL), ' ', ''))[[5]]
      
      stuTMatAll[loopI, 4] <- get(str_replace_all(paste('stuTRes', stuTMatAll[loopI, 1], 
                                                        sep = '', collapse = NULL), ' ', ''))[[3]]
      
      stuTMatAll[loopI, 5] <- sigMarkter(stuTMatAll[loopI, 4], c(0.1, 0.05, 0.01))
    }
  }
  
  #Output results
  {
    write.csv(stuTMatAll, 
              file = file.path('..', 'Output_R', 'stuTMatAll.csv'))
  }
  
  #Remove unused variables
  rm(tempMat,
     loopI)
}
```

### Table on significance test results

```{r significance table, echo = TRUE}
stuTMatAll
```

### Information on package information used in this R project (script): R
version 4.5.3 (2026-03-11) Platform: x86_64-apple-darwin20 Running
under: macOS Tahoe 26.3.1

Matrix products: default BLAS:
/System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
LAPACK:
/Library/Frameworks/R.framework/Versions/4.5-x86_64/Resources/lib/libRlapack.dylib;
LAPACK version 3.12.1

locale: [1]
en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: Europe/Oslo tzcode source: internal

attached base packages: [1] stats graphics grDevices utils datasets
methods base

other attached packages: [1] stringr_1.6.0 reshape2_1.4.5
robCompositions_2.4.2 data.table_1.18.2.1\
[5] pls_2.9-0 easyCODA_0.40.2 compositions_2.0-9 RColorBrewer_1.1-3\
[9] FactoMineR_2.13 factoextra_2.0.0 showtext_0.9-7 showtextdb_3.0\
[13] sysfonts_0.8.9 patchwork_1.3.2 ggplot2_4.0.2 doSNOW_1.0.20\
[17] snow_0.4-4 iterators_1.0.14 foreach_1.5.2 photobiology_0.14.2\
[21] SunCalcMeeus_0.1.3 baseline_1.3-7 prospectr_0.2.8 opusreader_0.4.2

loaded via a namespace (and not attached): [1] vcd_1.4-13 mathjaxr_2.0-0
tensorA_0.36.2.1 magrittr_2.0.4\
[5] rainbow_3.8 estimability_1.5.1 farver_2.1.2 rmarkdown_2.30\
[9] vctrs_0.7.1 RCurl_1.98-1.17 htmltools_0.5.9 truncnorm_1.0-9\
[13] deSolve_1.42 Formula_1.2-5 robustHD_0.8.4 hdrcde_3.5.0\
[17] pracma_2.4.6 parallelly_1.46.1 KernSmooth_2.23-26
htmlwidgets_1.6.4\
[21] mlr3_1.5.0 plyr_1.8.9 zoo_1.8-15 mlr3tuning_1.6.0\
[25] palmerpenguins_0.1.1 emmeans_2.0.2 uuid_1.2-2 lifecycle_1.0.5\
[29] pkgconfig_2.0.3 rsvd_1.0.5 Matrix_1.7-4 R6_2.6.1\
[33] fastmap_1.2.0 future_1.70.0 digest_0.6.39 sparsepca_0.1.2\
[37] colorspace_2.1-2 GGally_2.4.0 mlr3misc_0.21.0 ellipse_0.5.0\
[41] vegan_2.7-3 labeling_0.4.3 abind_1.4-8 mgcv_1.9-4\
[45] compiler_4.5.3 proxy_0.4-29 withr_3.0.2 S7_0.2.1\
[49] backports_1.5.0 carData_3.0-6 ggstats_0.13.0 MASS_7.3-65\
[53] bayesm_3.1-7 scatterplot3d_0.3-45 permute_0.9-10 flashClust_1.1-4\
[57] tools_4.5.3 lmtest_0.9-40 ranger_0.18.0 rrcov_1.7-7\
[61] otel_0.2.0 nnet_7.3-20 glue_1.8.0 lgr_0.5.2\
[65] cvTools_0.3.3 nlme_3.1-168 grid_4.5.3 checkmate_2.3.4\
[69] cluster_2.1.8.2 generics_0.1.4 fda_6.3.0 gtable_0.3.6\
[73] class_7.3-23 ca_0.71.1 tidyr_1.3.2 sp_2.2-1\
[77] car_3.1-5 ggrepel_0.9.8 pillar_1.11.1 VIM_7.0.0\
[81] bbotk_1.9.0 robustbase_0.99-7 splines_4.5.3 dplyr_1.2.0\
[85] lattice_0.22-9 survival_3.8-6 ks_1.15.1 SparseM_1.84-2\
[89] tidyselect_1.2.1 fds_1.8 knitr_1.51 gridExtra_2.3\
[93] zCompositions_1.6.0 stats4_4.5.3 xfun_0.56 DEoptimR_1.1-4\
[97] DT_0.34.0 stringi_1.8.7 yaml_2.3.12 boot_1.3-32\
[101] evaluate_1.0.5 codetools_0.2-20 kernlab_0.9-33 laeken_0.5.3\
[105] tibble_3.3.1 multcompView_0.1-11 cli_3.6.5 mlr3learners_0.14.0
[109] Rcpp_1.1.1 globals_0.19.1 parallel_4.5.3 leaps_3.2\
[113] ggfortify_0.4.19 mclust_6.1.2 bitops_1.0-9 paradox_1.0.1\
[117] listenv_0.10.1 perry_0.3.1 mvtnorm_1.3-6 mlr3pipelines_0.11.0
[121] scales_1.4.0 e1071_1.7-17 pcaPP_2.0-5 purrr_1.2.1\
[125] crayon_1.5.3 rlang_1.1.7
