---
title: "UAC signals in Pine cuttings and cultivars"
format:
  html:
    code-fold: true
    code-tools: true
editor: visual
---

```{r}
#| label: load Packages
#| echo: true
#| output: false
#Load palaeochem package
library(palaeochem)
#Load arial font into R
library(showtext)
font_add(family = "arial", regular = file.path('Packages_R', 'Fonts', 'arial.ttf'))
showtext_auto()

#Packages needed in Vivian's code
source(
  file.path(
    "Code_R", "00_Config_file.r"
  )
)
```

```{r}
#| label: load FTIRS pectra
#| echo: true
#| warning: false

#Read raw FTIR spectra
rawFTIRSpec <- read.csv(file = file.path('Data_R',
                                         'Original_Spectra.csv'),
                        row.names = 1)
colnames(rawFTIRSpec)[-1:-10] <- substring(colnames(rawFTIRSpec)[-1:-10], 2)

rawFTIRSpecSim <- rawFTIRSpec[, c(-1:-2,-4:-10)]
```

## Original FTIR spectra

The original FTIR spectra are between 4000 and 600 cm^-1^, with wavenumber interval of 4 cm^-1^. And the raw spectra looks like:

```{r}
#| label: Figure full raw spectra
#| echo: true
#| warning: false

plot_original_spec <-
rawFTIRSpecSim |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_original_spec)  

```

## Pretreating raw FTIR spectra

```{r}
#| label: prepare for different pretreating methodologies
#| echo: true
#| warning: false

#1 - No treatment - only extract wavenumbers
processing_steps_ftir1 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers

  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(min_wn = 600, max_wn = 3996)
    
  )
)

#2 - MSC
processing_steps_ftir2 <- list(
  functions = c(
    run_msc,
    get_filter_wavenumbers,
    get_baseline_offset
  ),
  parameters = list(
    list(),
    list(min_wn = 600, max_wn = 3996),
    list()
  )
)

#3 - ALS
processing_steps_ftir3 <- list(
  functions = c(
    get_savitzkygolay,
    run_als_baseline_correction,
    get_filter_wavenumbers,
    get_baseline_offset
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(lambda = 5, p = 0.001),
    list(min_wn = 600, max_wn = 3996),
    list()
  )
)

#4 - ALS-MSC
processing_steps_ftir4 <- list(
  functions = c(
    get_savitzkygolay,
    run_als_baseline_correction,
    get_filter_wavenumbers,
    run_msc,
    get_baseline_offset
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(lambda = 5, p = 0.001),
    list(min_wn = 600, max_wn = 3996),
    list(),
    list()
  )
)

#5 - SG-2nd derivative
processing_steps_ftir5 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 2),
    list(min_wn = 600, max_wn = 3996)
  )
)

#6 - signature region - MSC
processing_steps_ftir6 <- list(
  functions = c(
    run_msc,
    get_filter_wavenumbers,
    get_baseline_offset
  ),
  parameters = list(
    list(),
    list(min_wn = 800, max_wn = 1800),
    list()
  )
)

#7 - signature region - ALS
processing_steps_ftir7 <- list(
  functions = c(
    get_savitzkygolay,
    run_als_baseline_correction,
    get_filter_wavenumbers,
    get_baseline_offset
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(lambda = 5, p = 0.001),
    list(min_wn = 800, max_wn = 1800),
    list()
  )
)

#8 - signature region - ALS-MSC
processing_steps_ftir8 <- list(
  functions = c(
    get_savitzkygolay,
    run_als_baseline_correction,
    get_filter_wavenumbers,
    run_msc,
    get_baseline_offset
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(lambda = 5, p = 0.001),
    list(min_wn = 800, max_wn = 1800),
    list(),
    list()
  )
)

#9 - signature region - GC-2nd derivative
processing_steps_ftir9 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 2),
    list(min_wn = 800, max_wn = 1800)
  )
)
```

```{r}
#| label: pretreat the raw FTIR spectra with different methodologies
#| echo: true
#| warning: false

rawFTIRSpecSimMat <- as.matrix(rawFTIRSpecSim[, -1])
row.names(rawFTIRSpecSimMat) <- rawFTIRSpecSim[, 1]
data_preprocessed <- list(
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected1 = run_preprocessing_list(rawFTIRSpecSim, processing_steps_ftir1)
  ),
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected2 = run_preprocessing_list(rawFTIRSpecSimMat, processing_steps_ftir2)
  ),
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected3 = run_preprocessing_list(rawFTIRSpecSim, processing_steps_ftir3)
  ),
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected4 = run_preprocessing_list(rawFTIRSpecSim, processing_steps_ftir4)
  ),
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected5 = run_preprocessing_list(rawFTIRSpecSim, processing_steps_ftir5)
  ),
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected6 = run_preprocessing_list(rawFTIRSpecSimMat, processing_steps_ftir6)
  ),
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected7 = run_preprocessing_list(rawFTIRSpecSim, processing_steps_ftir7)
  ),
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected8 = run_preprocessing_list(rawFTIRSpecSim, processing_steps_ftir8)
  ),
  cbind(
    unique_id = rawFTIRSpecSim[, 1],
    corrected9 = run_preprocessing_list(rawFTIRSpecSim, processing_steps_ftir9)
  )
)
data_preprocessed <- lapply(data_preprocessed, as.data.frame)
for (loopI in 1:length(data_preprocessed)) {
  data_preprocessed[[loopI]][, -1] <-
    apply(data_preprocessed[[loopI]][, -1], c(1,2), as.numeric)
}
```

Several methodologies were applied on the raw FTIR spectra to search for better methodologies, including Multiplicative Scattering Correction (MSC), Asymmetric least squares (ALS) baseline correction and 2nd derivative spectra. Smoothing and baseline offset deduction are also applied on raw spectra. Two types of wavenumber ranges were investigated as well - the whole range (600-4000) and signature area only (800-1000).

### The whole range: MSC

```{r}
#| label: WR MSC
#| echo: true
#| warning: false

plot_msc_spec <-
data_preprocessed[[2]] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_msc_spec)  
```

### The whole range: smoothing + ALS

```{r}
#| label: WR Smoothing ALS
#| echo: true
#| warning: false
plot_als_spec <-
data_preprocessed[[3]] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_als_spec)  
```

### The whole range: smoothing + ALS + MSC

```{r}
#| label: WR Smoothing ALS MSC
#| echo: true
#| warning: false
plot_alsmsc_spec <-
data_preprocessed[[4]] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_alsmsc_spec)
```

### The whole range: smoothing + 2^nd^ derivative

```{r}
#| label: WR Smoothing 2nd derivative
#| echo: true
#| warning: false
plot_2nd_spec <-
data_preprocessed[[5]] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_2nd_spec)
```

### Signature area: MSC

```{r}
#| label: Sg MSC
#| echo: true
#| warning: false

plot_sg_msc_spec <-
data_preprocessed[[6]] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_sg_msc_spec)  
```

### Signature area: smoothing + ALS

```{r}
#| label: Sg Smoothing ALS
#| echo: true
#| warning: false
plot_sg_als_spec <-
data_preprocessed[[7]] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_sg_als_spec)  
```

### Signature area: smoothing + ALS + MSC

```{r}
#| label: Sg Smoothing ALS MSC
#| echo: true
#| warning: false
plot_sg_alsmsc_spec <-
data_preprocessed[[8]] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_sg_alsmsc_spec)
```

### Signature area: smoothing + 2^nd^ derivative

```{r}
#| label: Sg Smoothing 2nd derivative
#| echo: true
#| warning: false
plot_sg_2nd_spec <-
data_preprocessed[[9]] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
)

print(plot_sg_2nd_spec)
```

## Input GCMS signals

```{r}
#| label: read GCMS
#| echo: true
#| warning: false
#GC-MS data on cuttings and pot cultivars
GCDataRawCut <- as.matrix(read.csv(
  file.path('Data_R', '04_11_2025_gc_cutting_data_2025.csv'),
  header = TRUE,
  row.names = 1))
GCDataRawCul <- as.matrix(read.csv(
  file.path('Data_R', '01_07_2025_gc_cultivar_data_2025.csv'),
  header = TRUE,
  row.names = 1))

#Combine these two raw datasets, with batch information
GCData <- rbind(GCDataRawCut[, c(8, 10:14)],
                cbind(GCDataRawCul[, c(8, 10:12)], 
                      sampling_number = NA, 
                      GCDataRawCul[, 13])
)

#Exclude the last batch (u), as an unexpected crash happened before this
#batch, making reseult from this batch abnormal
GCData <- GCData[-which(GCData[, 6] == 'u'),]

#Loop to calculate the mean values for each sample
uniqueBatchInfo <- unique(GCData[, 2:5])

#Create a matrix to fill in the mean GC-MS signals
GCDataMean <- matrix(nrow = nrow(uniqueBatchInfo),
                     ncol = ncol(uniqueBatchInfo) + 2,
                     dimnames = list(c(),
                                     c('calibrated',
                                       'sd',
                                       colnames(uniqueBatchInfo)
                                     )
                     )
)

#Add batch number to the GCMS data matrix
GCDataMean[, 3:ncol(GCDataMean)] <- uniqueBatchInfo

#Loop to calculate the mean signals of GCMS duplicates
for (loopI in 1:nrow(GCDataMean)) {
  
  #Check wheter current sample is cutting or pot cultivar 
  if (is.na(GCDataMean[loopI, 6])) {
    
    #Mean and sd for pot cultivar
    GCDataMean[loopI, 1] <- mean(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) 
    )
    , 1])
    )
    GCDataMean[loopI, 2] <- sd(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5])
    )
    , 1])
    )
    
  }else {
    
    #Mean and sd for cutting
    GCDataMean[loopI, 1] <- mean(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) &
                     (GCData[, 5] == GCDataMean[loopI, 6])
    )
    , 1])
    )
    GCDataMean[loopI, 2] <- sd(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) &
                     (GCData[, 5] == GCDataMean[loopI, 6])
    )
    , 1])
    )
    
  }
}

#Transfer the matrix back to data frame, and transfer values to number if necessary
GCDataMean <- as.data.frame(GCDataMean)
GCDataMean[, 1] <- as.numeric(GCDataMean[, 1])

#Create a vector to stroe the index for batch information matching
tempInd <- vector('numeric', length = nrow(rawFTIRSpec))

#Loop to match the batch information between GC-MS and FTIR datasets
for (loopI in 1:nrow(rawFTIRSpec)) {

#Check whether current sample is pot cultivar
if (rawFTIRSpec[loopI, 5] == 'cul') {
  
  #Check whether current FTIR spectrum has any GCMS signal
  if (any((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
          (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
          (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
          (is.na(GCDataMean[, 6]))
  )
  ) {
    
    #Record the FTIR matrix row index for current GCMS signal
    tempInd[loopI] <- which((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
                              (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
                              (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
                              (is.na(GCDataMean[, 6])))
    
  }else {
    
    #Record nothing if there is no match between current FTIR specrum and
      #any GCMS signal
    tempInd[loopI] <- NA
    
  }
  
#Check whether current sample is cutting
} else if (rawFTIRSpec[loopI, 5] == 'cut') {
  
  #Check whether current FTIR spectrum has any GCMS signal
  if (any((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
          (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
          (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
          (GCDataMean[, 6] == rawFTIRSpec[loopI, 9])
  )
  ) {
    
    #Record the FTIR matrix row index for current GCMS signal
    tempInd[loopI] <- which((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
                              (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
                              (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
                              (GCDataMean[, 6] == rawFTIRSpec[loopI, 9]))
    
  }else {
    
    #Record nothing if there is no match between current FTIR spectrum and
      #any GCMS signal
    tempInd[loopI] <- NA
    
  }
  
} else {
  
  #Record nothing if there is no match between current FTIR spectrum and
    #any GCMS signal
  tempInd[loopI] <- NA
  
}
}

#Attach GCMS signals to  peak statistics (height or area)
pcaMatHeight <- cbind(rawFTIRSpec,
                    GCDataMean[tempInd, c(1,2)])[, -11:-1772]
colnames(pcaMatHeight)[11] <- 'pCA_ng_grain'

rm(loopI,
   tempInd)
```

## PLS analysis between FTIR spectra and GCMS signals

In this section, the GCMS signals were input into PLS model as response variable, and FTIR spectral signals pretreated by different methodologies were input as independent variables. The decision on the number of component is made by function pls::selectNcomp, with the method of randomization.

```{r}
#| label: run PLS
#| echo: true
#| warning: false

#Check whetether there is already a RData file containing the PLS model, skip the lines generating the PLS model and read the archives driectly if true
if (file.exists(file.path('Output_R', 'pls_raw.RData'))) {
  load(file.path('Output_R', 'pls_raw.RData'))
  load(file.path('Output_R', 'pls_msc.RData'))
  load(file.path('Output_R', 'pls_als.RData'))
  load(file.path('Output_R', 'pls_als_msc.RData'))
  load(file.path('Output_R', 'pls_2nd.RData'))
  load(file.path('Output_R', 'pls_sig_msc.RData'))
  load(file.path('Output_R', 'pls_sig_als.RData'))
  load(file.path('Output_R', 'pls_sig_als_msc.RData'))
  load(file.path('Output_R', 'pls_sig_2nd.RData'))
}else {
  pls_raw <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                  data_preprocessed[[1]][, -1],
                  pCA_ng_grain = pcaMatHeight[, 11]),
                 pred_col = 'pCA_ng_grain')

  pls_msc <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[2]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11]),
                   pred_col = 'pCA_ng_grain')
  
  pls_als <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[3]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11]),
                   pred_col = 'pCA_ng_grain')
  
  pls_als_msc <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[4]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11]),
                   pred_col = 'pCA_ng_grain')
  
  pls_2nd <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[5]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11]),
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_msc <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[6]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11]),
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_als <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[7]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11]),
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_als_msc <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[8]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11]),
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_2nd <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[9]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11]),
                   pred_col = 'pCA_ng_grain')
  
  #Save the pls results as .RData indepecdently (so that they do not exceed the 100 MB limitation from GitHub)
  save(pls_raw,file = file.path('Output_R', 'pls_raw.RData'))
  save(pls_msc, file = file.path('Output_R', 'pls_msc.RData'))
  save(pls_als, file = file.path('Output_R', 'pls_als.RData'))
  save(pls_als_msc, file = file.path('Output_R', 'pls_als_msc.RData'))
  save(pls_2nd, file = file.path('Output_R', 'pls_2nd.RData'))
  save(pls_sig_msc, file = file.path('Output_R', 'pls_sig_msc.RData'))
  save(pls_sig_als, file = file.path('Output_R', 'pls_sig_als.RData'))
  save(pls_sig_als_msc, file = file.path('Output_R', 'pls_sig_als_msc.RData'))
  save(pls_sig_2nd, file = file.path('Output_R', 'pls_sig_2nd.RData'))
}


```

## The whole range: raw spectra

### Original spectra

```{r}
#| label: pls - raw spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_raw <- selectNcomp(pls_raw, method = "randomization", plot = FALSE)
numComps_raw <- 6

#Scatter plot
plot_raw_scatter <- plot_pls_scatter(plsModel = pls_raw,
                                     numComp = numComps_raw)

#Coefficients plot
plot_raw_coefficients <- plot_pls_coefficients(pls_raw,
                                               numComp = numComps_raw,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Loadings plot
plot_raw_loadings <- plot_pls_loadings(pls_raw,
                                               numComp = 2,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_raw_variances <- plot_pls_variance(pls_raw,
                                               numComp = numComps_raw)

#Residuals
plot_raw_residuals <- plot_pls_residuals(pls_raw,
                                               numComp = numComps_raw,
                                               wavenumberRange = c(500, 4000, 500)
                                         )
```

```{r}
#| label: pls - raw spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600

 (plot_original_spec | plot_raw_scatter) /
    (plot_raw_variances | plot_raw_coefficients) /
    (plot_raw_loadings | plot_raw_residuals)
```

### The whole range: MSC

```{r}
#| label: pls - msc spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_msc <- selectNcomp(pls_msc, method = "randomization", plot = FALSE)
numComps_msc <- 6

#Scatter plot
plot_msc_scatter <- plot_pls_scatter(plsModel = pls_msc,
                                     numComp = numComps_msc)

#Coefficients plot
plot_msc_coefficients <- plot_pls_coefficients(pls_msc,
                                               numComp = numComps_msc,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[2]])[-1])
                                                          )
                                               )

#Loadings plot
plot_msc_loadings <- plot_pls_loadings(pls_msc,
                                               numComp = 2,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[2]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_msc_variances <- plot_pls_variance(pls_msc,
                                               numComp = numComps_msc)

#Residuals
plot_msc_residuals <- plot_pls_residuals(pls_msc,
                                               numComp = numComps_msc,
                                               wavenumberRange = c(500, 4000, 500)
                                         )
```

```{r}
#| label: pls - msc spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600

 (plot_msc_spec | plot_msc_scatter) /
    (plot_msc_variances | plot_msc_coefficients) /
    (plot_msc_loadings | plot_msc_residuals)
```

### The whole range: smoothing + ALS

```{r}
#| label: pls - als spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_als <- selectNcomp(pls_als, method = "randomization", plot = FALSE)
numComps_als <- 6

#Scatter plot
plot_als_scatter <- plot_pls_scatter(plsModel = pls_als,
                                     numComp = numComps_als)

#Coefficients plot
plot_als_coefficients <- plot_pls_coefficients(pls_als,
                                               numComp = numComps_als,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[3]])[-1])
                                                          )
                                               )

#Loadings plot
plot_als_loadings <- plot_pls_loadings(pls_als,
                                               numComp = 2,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[3]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_als_variances <- plot_pls_variance(pls_als,
                                               numComp = numComps_als)

#Residuals
plot_als_residuals <- plot_pls_residuals(pls_als,
                                               numComp = numComps_als,
                                               wavenumberRange = c(500, 4000, 500)
                                         )
```

```{r}
#| label: pls - als spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600

 (plot_als_spec | plot_als_scatter) /
    (plot_als_variances | plot_als_coefficients) /
    (plot_als_loadings | plot_als_residuals)
```

### The whole range: smoothing + ALS + MSC

```{r}
#| label: pls - als msc spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_als_msc <- selectNcomp(pls_als_msc, method = "randomization", plot = FALSE)
numComps_als_msc <- 6

#Scatter plot
plot_als_msc_scatter <- plot_pls_scatter(plsModel = pls_als_msc,
                                     numComp = numComps_als_msc)

#Coefficients plot
plot_als_msc_coefficients <- plot_pls_coefficients(pls_als_msc,
                                               numComp = numComps_als_msc,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[4]])[-1])
                                                          )
                                               )

#Loadings plot
plot_als_msc_loadings <- plot_pls_loadings(pls_als_msc,
                                               numComp = 2,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[4]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_als_msc_variances <- plot_pls_variance(pls_als_msc,
                                               numComp = numComps_als_msc)

#Residuals
plot_als_msc_residuals <- plot_pls_residuals(pls_als_msc,
                                               numComp = numComps_als_msc,
                                               wavenumberRange = c(500, 4000, 500)
                                         )
```

```{r}
#| label: pls - als msc spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600

 (plot_alsmsc_spec | plot_als_msc_scatter) /
    (plot_als_msc_variances | plot_als_msc_coefficients) /
    (plot_als_msc_loadings | plot_als_msc_residuals)
```

### The whole range: smoothing + 2^nd^ derivative

```{r}
#| label: pls - 2nd derivative spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_2nd <- selectNcomp(pls_2nd, method = "randomization", plot = FALSE)
numComps_2nd <- 6

#Scatter plot
plot_2nd_scatter <- plot_pls_scatter(plsModel = pls_2nd,
                                     numComp = numComps_2nd)

#Coefficients plot
plot_2nd_coefficients <- plot_pls_coefficients(pls_2nd,
                                               numComp = numComps_2nd,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[5]])[-1])
                                                          )
                                               )

#Loadings plot
plot_2nd_loadings <- plot_pls_loadings(pls_2nd,
                                               numComp = 1,
                                               wavenumberRange = c(500, 4000, 500),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[5]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_2nd_variances <- plot_pls_variance(pls_2nd,
                                               numComp = numComps_2nd)

#Residuals
plot_2nd_residuals <- plot_pls_residuals(pls_2nd,
                                               numComp = numComps_2nd,
                                               wavenumberRange = c(500, 4000, 500)
                                         )
```

```{r}
#| label: pls - 2nd derivative spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600

 (plot_2nd_spec | plot_2nd_scatter) /
    (plot_2nd_variances | plot_2nd_coefficients) /
    (plot_2nd_loadings | plot_2nd_residuals)
```

### Signature area: MSC

```{r}
#| label: pls - sig msc spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_sig_msc <- selectNcomp(pls_sig_msc, method = "randomization", plot = FALSE)
numComps_sig_msc <- 6

#Scatter plot
plot_sig_msc_scatter <- plot_pls_scatter(plsModel = pls_sig_msc,
                                     numComp = numComps_sig_msc)

#Coefficients plot
plot_sig_msc_coefficients <- plot_pls_coefficients(pls_sig_msc,
                                               numComp = numComps_sig_msc,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[6]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_msc_loadings <- plot_pls_loadings(pls_sig_msc,
                                               numComp = 2,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[6]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_msc_variances <- plot_pls_variance(pls_sig_msc,
                                               numComp = numComps_sig_msc)

#Residuals
plot_sig_msc_residuals <- plot_pls_residuals(pls_sig_msc,
                                               numComp = numComps_sig_msc,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig msc spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

 (plot_sg_msc_spec | plot_sig_msc_scatter) /
    (plot_sig_msc_variances | plot_sig_msc_coefficients) /
    (plot_sig_msc_loadings | plot_sig_msc_residuals)
```

### Signature area: smoothing + ALS

```{r}
#| label: pls - sig als spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_sig_als <- selectNcomp(pls_sig_als, method = "randomization", plot = FALSE)
numComps_sig_als <- 6

#Scatter plot
plot_sig_als_scatter <- plot_pls_scatter(plsModel = pls_sig_als,
                                     numComp = numComps_sig_als)

#Coefficients plot
plot_sig_als_coefficients <- plot_pls_coefficients(pls_sig_als,
                                               numComp = numComps_sig_als,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[7]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_als_loadings <- plot_pls_loadings(pls_sig_als,
                                               numComp = 2,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[7]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_als_variances <- plot_pls_variance(pls_sig_als,
                                               numComp = numComps_sig_als)

#Residuals
plot_sig_als_residuals <- plot_pls_residuals(pls_sig_als,
                                               numComp = numComps_sig_als,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig als spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

 (plot_sg_als_spec | plot_sig_als_scatter) /
    (plot_sig_als_variances | plot_sig_als_coefficients) /
    (plot_sig_als_loadings | plot_sig_als_residuals)
```

### Signature area: smoothing + ALS + MSC

```{r}
#| label: pls - sig als msc spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_sig_als_msc <- selectNcomp(pls_sig_als_msc, method = "randomization", plot = FALSE)
numComps_sig_als_msc <- 6

#Scatter plot
plot_sig_als_msc_scatter <- plot_pls_scatter(plsModel = pls_sig_als_msc,
                                     numComp = numComps_sig_als_msc)

#Coefficients plot
plot_sig_als_msc_coefficients <- plot_pls_coefficients(pls_sig_als_msc,
                                               numComp = numComps_sig_als_msc,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[8]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_als_msc_loadings <- plot_pls_loadings(pls_sig_als_msc,
                                               numComp = 2,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[8]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_als_msc_variances <- plot_pls_variance(pls_sig_als_msc,
                                               numComp = numComps_sig_als_msc)

#Residuals
plot_sig_als_msc_residuals <- plot_pls_residuals(pls_sig_als_msc,
                                               numComp = numComps_sig_als_msc,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig als msc spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

 (plot_sg_alsmsc_spec | plot_sig_als_msc_scatter) /
    (plot_sig_als_msc_variances | plot_sig_als_msc_coefficients) /
    (plot_sig_als_msc_loadings | plot_sig_als_msc_residuals)
```

### Signature area: smoothing + 2^nd^ derivative

```{r}
#| label: pls - sig 2nd derivative spectra
#| echo: true
#| warning: false
#Setup the number of components to be analysed
numComps_sig_2nd <- selectNcomp(pls_sig_2nd, method = "randomization", plot = FALSE)
numComps_sig_2nd <- 6

#Scatter plot
plot_sig_2nd_scatter <- plot_pls_scatter(plsModel = pls_sig_2nd,
                                     numComp = numComps_sig_2nd)

#Coefficients plot
plot_sig_2nd_coefficients <- plot_pls_coefficients(pls_sig_2nd,
                                               numComp = numComps_sig_2nd,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[9]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_2nd_loadings <- plot_pls_loadings(pls_sig_2nd,
                                               numComp = 2,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[9]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_2nd_variances <- plot_pls_variance(pls_sig_2nd,
                                               numComp = numComps_sig_2nd)

#Residuals
plot_sig_2nd_residuals <- plot_pls_residuals(pls_sig_2nd,
                                               numComp = numComps_sig_2nd,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig 2nd derivative spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

 (plot_sg_2nd_spec | plot_sig_2nd_scatter) /
    (plot_sig_2nd_variances | plot_sig_2nd_coefficients) /
    (plot_sig_2nd_loadings | plot_sig_2nd_residuals)
```
