---
title: "UAC signals in Pine cuttings and cultivars - refined pipeline"
author: "Tianyuan"
format: html
editor: visual
execute:
  warning: false
---

```{r}
#| label: load Packages
#| echo: false
#| output: false
#| warning: false
#| message: false
#Load palaeochem package
library(palaeochem)
#Load arial font into R
library(showtext)
font_add(family = "arial", regular = file.path('Packages_R', 'Fonts', 'arial.ttf'))
showtext_auto()

#Packages needed in Vivian's code
source(
  file.path(
    "Code_R", "00_Config_file.r"
  )
)

rm(package_list)
```

```{r}
#| label: load FTIRS pectra
#| echo: false
#| output: false
#| warning: false
#| message: false

#Read raw FTIR spectra
rawFTIRSpec <- read.csv(file = file.path('Data_R',
                                         'Original_Spectra.csv'),
                        row.names = 1)
colnames(rawFTIRSpec)[-1:-10] <- substring(colnames(rawFTIRSpec)[-1:-10], 2)
```

## Original FTIR spectra

The original FTIR spectra are between 4000 and 600 cm^-1^, with wavenumber interval of 4 cm^-1^. And the raw spectra from potted cultivars and branch cuttings looks like:

```{r}
#| label: Figure full raw spectra
#| echo: false
#| warning: false
#| message: false

plot_original_spec_cul <-
rawFTIRSpec[rawFTIRSpec[, 5] == 'cul', c(-1:-2,-4:-10)] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
  labs(title = 'Potted cultivars',
       x = '')

plot_original_spec_cut <-
rawFTIRSpec[rawFTIRSpec[, 5] == 'cut', c(-1:-2,-4:-10)] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
    labs(title = 'Branch cuttings')

plot_original_spec_cul /
  plot_original_spec_cut

```

## Pretreating raw FTIR spectra - smoothing + 2nd derivatives

The raw FTIR spectra were firstly smoothed using Savitzky-Golay filter (width = 7, poly = 2). And second derivatives were calculated from smoothed spectra (in EMSC::SavitzkyGolay function, deriv = 2). Only signature area (800 - 1800 cm^-1^) was used in following analysis. The pre-treated spectra looked like:

```{r}
#| label: pretreating FTIR spectra (smoothing + 2nd derivatives)
#| echo: false
#| output: false
#| warning: false
#| message: false

#5 - smoothing + 2nd derivative
processing_steps_ftir5 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 2),
    list(min_wn = 800, max_wn = 1800)
  )
)

data_preprocessed <- list(
  cbind(
    unique_id = rawFTIRSpec[, 3],
    corrected5 = run_preprocessing_list(rawFTIRSpec[, c(-1:-2,-4:-10)], processing_steps_ftir5)
  )
)
data_preprocessed <- lapply(data_preprocessed, as.data.frame)
for (loopI in 1:length(data_preprocessed)) {
  data_preprocessed[[loopI]][, -1] <-
    apply(data_preprocessed[[loopI]][, -1], c(1,2), as.numeric)
}

rm(processing_steps_ftir5)
```

```{r}
#| label: signature 2nd derivative
#| echo: true
#| output: true
#| warning: false
#| message: false
plot_2nd_spec_cul <-
data_preprocessed[[1]][rawFTIRSpec[, 5] == 'cul',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
  labs(title = 'Potted cultivars',
       x = '')

plot_2nd_spec_cut <-
data_preprocessed[[1]][rawFTIRSpec[, 5] == 'cut',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
    labs(title = 'Branch cuttings')

plot_2nd_spec_cul /
  plot_2nd_spec_cut
```

## Input GCMS signals

UAC GCMS signals were input and merged with FTIR spectra according to their batch information

```{r}
#| label: read GCMS
#| echo: false
#| output: false
#| warning: false
#| message: false
#GC-MS data on cuttings and pot cultivars
GCDataRawCut <- as.matrix(read.csv(
  file.path('Data_R', '04_11_2025_gc_cutting_data_2025.csv'),
  header = TRUE,
  row.names = 1))
GCDataRawCul <- as.matrix(read.csv(
  file.path('Data_R', '01_07_2025_gc_cultivar_data_2025.csv'),
  header = TRUE,
  row.names = 1))

#Combine these two raw datasets, with batch information
GCData <- rbind(GCDataRawCut[, c(8, 10:14)],
                cbind(GCDataRawCul[, c(8, 10:12)], 
                      sampling_number = NA, 
                      GCDataRawCul[, 13])
)

#Exclude the last batch (u), as an unexpected crash happened before this
#batch, making reseult from this batch abnormal
GCData <- GCData[-which(GCData[, 6] == 'u'),]

#Loop to calculate the mean values for each sample
uniqueBatchInfo <- unique(GCData[, 2:5])

#Create a matrix to fill in the mean GC-MS signals
GCDataMean <- matrix(nrow = nrow(uniqueBatchInfo),
                     ncol = ncol(uniqueBatchInfo) + 2,
                     dimnames = list(c(),
                                     c('calibrated',
                                       'sd',
                                       colnames(uniqueBatchInfo)
                                     )
                     )
)

#Add batch number to the GCMS data matrix
GCDataMean[, 3:ncol(GCDataMean)] <- uniqueBatchInfo

#Loop to calculate the mean signals of GCMS duplicates
for (loopI in 1:nrow(GCDataMean)) {
  
  #Check wheter current sample is cutting or pot cultivar 
  if (is.na(GCDataMean[loopI, 6])) {
    
    #Mean and sd for pot cultivar
    GCDataMean[loopI, 1] <- mean(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) 
    )
    , 1])
    )
    GCDataMean[loopI, 2] <- sd(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5])
    )
    , 1])
    )
    
  }else {
    
    #Mean and sd for cutting
    GCDataMean[loopI, 1] <- mean(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) &
                     (GCData[, 5] == GCDataMean[loopI, 6])
    )
    , 1])
    )
    GCDataMean[loopI, 2] <- sd(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) &
                     (GCData[, 5] == GCDataMean[loopI, 6])
    )
    , 1])
    )
    
  }
}

#Transfer the matrix back to data frame, and transfer values to number if necessary
GCDataMean <- as.data.frame(GCDataMean)
GCDataMean[, 1] <- as.numeric(GCDataMean[, 1])

#Create a vector to stroe the index for batch information matching
tempInd <- vector('numeric', length = nrow(rawFTIRSpec))

#Loop to match the batch information between GC-MS and FTIR datasets
for (loopI in 1:nrow(rawFTIRSpec)) {

#Check whether current sample is pot cultivar
if (rawFTIRSpec[loopI, 5] == 'cul') {
  
  #Check whether current FTIR spectrum has any GCMS signal
  if (any((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
          (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
          (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
          (is.na(GCDataMean[, 6]))
  )
  ) {
    
    #Record the FTIR matrix row index for current GCMS signal
    tempInd[loopI] <- which((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
                              (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
                              (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
                              (is.na(GCDataMean[, 6])))
    
  }else {
    
    #Record nothing if there is no match between current FTIR specrum and
      #any GCMS signal
    tempInd[loopI] <- NA
    
  }
  
#Check whether current sample is cutting
} else if (rawFTIRSpec[loopI, 5] == 'cut') {
  
  #Check whether current FTIR spectrum has any GCMS signal
  if (any((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
          (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
          (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
          (GCDataMean[, 6] == rawFTIRSpec[loopI, 9])
  )
  ) {
    
    #Record the FTIR matrix row index for current GCMS signal
    tempInd[loopI] <- which((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
                              (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
                              (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
                              (GCDataMean[, 6] == rawFTIRSpec[loopI, 9]))
    
  }else {
    
    #Record nothing if there is no match between current FTIR spectrum and
      #any GCMS signal
    tempInd[loopI] <- NA
    
  }
  
} else {
  
  #Record nothing if there is no match between current FTIR spectrum and
    #any GCMS signal
  tempInd[loopI] <- NA
  
}
}

#Attach GCMS signals to  peak statistics (height or area)
pcaMatHeight <- cbind(rawFTIRSpec,
                    GCDataMean[tempInd, c(1,2)])[, -11:-1772]
colnames(pcaMatHeight)[11] <- 'pCA_ng_grain'
pcaMatHeight <- pcaMatHeight[!is.na(pcaMatHeight[, 11]),]

rm(loopI,
   tempInd)
```

## PLS analysis between FTIR spectra and GCMS signals

In this section, the GCMS signals were input into PLS model as response variable, and pretreated FTIR spectral signals were input as independent variables. The decision on the number of component is made according to the accumulated variance explained by components, and the number is selected when \>85% variance is explained.

```{r}
#| label: run PLS
#| echo: false
#| output: false
#| warning: false
#| message: false

#Check whetether there is already a RData file containing the PLS model, skip the lines generating the PLS model and read the archives driectly if true
if (file.exists(file.path('Output_R', 'pls_sig_2nd_cul.RData'))) {
  load(file.path('Output_R', 'pls_sig_2nd_cul.RData'))
  
  load(file.path('Output_R', 'pls_sig_2nd_cut.RData'))
}else {
  pls_sig_2nd_cul <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[1]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cul'),],
                   pred_col = 'pCA_ng_grain')
  
  
  pls_sig_2nd_cut <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[1]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cut'),],
                   pred_col = 'pCA_ng_grain')
  
  #Save the pls results as .RData indepecdently (so that they do not exceed the 100 MB limitation from GitHub)
  save(pls_sig_2nd_cul, file = file.path('Output_R', 'pls_sig_2nd_cul.RData'))
  
  save(pls_sig_2nd_cut, file = file.path('Output_R', 'pls_sig_2nd_cut.RData'))
}
```

```{r}
#| label: pls - sig 2nd derivative spectra
#| echo: false
#| output: false
#| warning: false
#| message: false
#Setup the number of components to be analysed
numComps_sig_2nd_cul <- pick_num_comp(pls_sig_2nd_cul, 0.85)

#Scatter plot
plot_sig_2nd_scatter_cul <- plot_pls_scatter(plsModel = pls_sig_2nd_cul,
                                     numComp = numComps_sig_2nd_cul)

#Coefficients plot
plot_sig_2nd_coefficients_cul <- plot_pls_coefficients(pls_sig_2nd_cul,
                                               numComp = numComps_sig_2nd_cul,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_2nd_loadings_cul <- plot_pls_loadings(pls_sig_2nd_cul,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_2nd_variances_cul <- plot_pls_variance(pls_sig_2nd_cul,
                                               numComp = numComps_sig_2nd_cul)

#Residuals
plot_sig_2nd_residuals_cul <- plot_pls_residuals(pls_sig_2nd_cul,
                                               numComp = numComps_sig_2nd_cul,
                                               wavenumberRange = c(800, 1800, 200)
                                         )

#Setup the number of components to be analysed
numComps_sig_2nd_cut <- pick_num_comp(pls_sig_2nd_cut, 0.85)

#Scatter plot
plot_sig_2nd_scatter_cut <- plot_pls_scatter(plsModel = pls_sig_2nd_cut,
                                     numComp = numComps_sig_2nd_cut)

#Coefficients plot
plot_sig_2nd_coefficients_cut <- plot_pls_coefficients(pls_sig_2nd_cut,
                                               numComp = numComps_sig_2nd_cut,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_2nd_loadings_cut <- plot_pls_loadings(pls_sig_2nd_cut,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_2nd_variances_cut <- plot_pls_variance(pls_sig_2nd_cut,
                                               numComp = numComps_sig_2nd_cut)

#Residuals
plot_sig_2nd_residuals_cut <- plot_pls_residuals(pls_sig_2nd_cut,
                                               numComp = numComps_sig_2nd_cut,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig 2nd derivative spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| echo: false
#| output: true
#| warning: false
#| message: false

 (plot_2nd_spec_cul | plot_sig_2nd_scatter_cul) /
    (plot_sig_2nd_variances_cul | plot_sig_2nd_coefficients_cul) /
    (plot_sig_2nd_loadings_cul | plot_sig_2nd_residuals_cul)

 (plot_2nd_spec_cut | plot_sig_2nd_scatter_cut) /
    (plot_sig_2nd_variances_cut | plot_sig_2nd_coefficients_cut) /
    (plot_sig_2nd_loadings_cut | plot_sig_2nd_residuals_cut)

(plot_2nd_spec_cul | plot_2nd_spec_cut) /
    (plot_sig_2nd_scatter_cul | plot_sig_2nd_scatter_cut) /
    (plot_sig_2nd_coefficients_cul | plot_sig_2nd_coefficients_cut)
```

```{r}
#| label: combine malformation data with PLS and GCMS
#| echo: false
#| output: false
#| warning: false
#| message: false
#Read the cutting branch data, and replace all na with 0
cuttingsRaw <- read.csv(file = 
                          file.path('Data_R', 
                                    'Malformations_Cuttings.csv'),
                        row.names = 1)

#Replace all NAs with 0
cuttingsRaw[is.na(cuttingsRaw)] <- 0

#Extract the batch information
batchInfoTemp <- strsplit(x = cuttingsRaw[,1], split = '_')
cuttingsBatchInfo <- cbind(rep('Cuttings', length(batchInfoTemp)),
                           rep('2025', length(batchInfoTemp)),
                           matrix(unlist(batchInfoTemp), ncol = 4, byrow = TRUE))
colnames(cuttingsBatchInfo) <- c('Sample', 'Year', 'Frame', 'Treatment',
                                 'Plant', 'Duplicate')

#Delete unused variables
rm(batchInfoTemp)

#Read the cultivar branch data, and replace all na with 0
cultivarsRaw <- read.csv(file = 
                           file.path('Data_R', 
                                     'Malformations_Cultivars.csv'),
                         row.names = 1)

#Replace all NAs with 0
cultivarsRaw[is.na(cultivarsRaw)] <- 0

#Extract the batch information
batchInfoTemp <- strsplit(x = cultivarsRaw[,1], split = '_')
cultivarsBatchInfo <- cbind(rep('Cultivars', length(batchInfoTemp)),
                            matrix(unlist(batchInfoTemp), ncol = 4, byrow = TRUE))
colnames(cultivarsBatchInfo) <- c('Sample', 'Year', 'Frame', 'Treatment',
                                  'Plant')

#Delete unused variables
rm(batchInfoTemp)

#Combine counting results in duplicated samples, by summing them up
tempVec <- unique(cultivarsRaw[, 1])

#Matrix for all malformation data
tempMat <- matrix(nrow = length(tempVec),
                  ncol = ncol(cultivarsRaw),
                  dimnames = list(c(),
                                  colnames(cultivarsRaw)))

#Matrix for all batch information on malformation data
tempMat2 <- matrix(nrow = length(tempVec),
                   ncol = ncol(cultivarsBatchInfo),
                   dimnames = list(c(),
                                   colnames(cultivarsBatchInfo)))

#Fill sample label into matrix
tempMat[, 1] <- tempVec

#Loop to add duplicates together
for (loopI in 1:nrow(tempMat)) {
  
  #Check  if there are duplicates in current sample
  if (length(which(cultivarsRaw[, 1] == tempMat[loopI, 1])) > 1) {
    
    #Sum up all counting results among duplicates
    tempMat[loopI, 2:ncol(tempMat)] <- apply(cultivarsRaw[
      which(cultivarsRaw[, 1] == tempMat[loopI, 1]), 
      2:ncol(tempMat)
    ], 2, sum)
    
  }else{
    
    #Transfer couting results into matrix
    tempMat[loopI, 2:ncol(tempMat)] <- cultivarsRaw[
      which(cultivarsRaw[, 1] == tempMat[loopI, 1]), 2:ncol(tempMat)
    ]
    
  }
  
  #Record the corresponding batch information
  tempMat2[loopI,] <- cultivarsBatchInfo[which(cultivarsRaw[, 1] == tempMat[loopI, 1])[1],]
  
}

#Combine batch information with counting results
cultivarsCom <- cbind(tempMat2, tempMat[, -1])

rm(loopI,
   tempMat,
   tempMat2,
   tempVec)

#Combine malformation data with PLS fitted results from different pretreatment methodologies
data_preprocessed_ftir_gcms_malformation <- vector(mode = 'list',
                                                   length = length(data_preprocessed))

#Signature area 2nd derivative spectra
data_preprocessed_ftir_gcms_malformation[[1]] <- matrix_malformation(
  batchInfoMat = pcaMatHeight[,1:10],
  PLSFittedValues = c(pls_sig_2nd_cul$fitted.values[,, numComps_sig_2nd_cul],
                      pls_sig_2nd_cut$fitted.values[,, numComps_sig_2nd_cut]),
  GCMSValuesMat = pcaMatHeight[,11:12],
  cultivarsData = cultivarsCom,
  cuttingsData = cuttingsRaw,
  cuttingsBatch = cuttingsBatchInfo)
```

## Overviews on GC-MS signals and FTIR PLS fitted values

GC-MS signals and PLS fitted values from FTIR spectra were shown in box plots, and were further sub-divided according to UV and Control groups.

```{r}
#| label: preperation on presenting GC-MS and FTIR data
#| echo: false
#| output: false
#| warning: false
#| message: false

#Overall plot on PLS fitted values
plot_PLS_Fitted <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_PLS_Fitted)) {
  plot_PLS_Fitted[[loopI]] <- plot_box(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'PLS')
}

#Plot groupped by potted cultivars and branch cuttings
plot_PLS_Fitted_Materials <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_PLS_Fitted)) {
  plot_PLS_Fitted_Materials[[loopI]] <- plot_box(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'PLS',
                                       groupName = TRUE)
}

#Overall plot on GCMS signals
plot_GCMS <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_GCMS)) {
  plot_GCMS[[loopI]] <- plot_box(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'GCMS')
}

#Plot groupped by potted cultivars and branch cuttings
plot_GCMS_Materials <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_GCMS_Materials)) {
  plot_GCMS_Materials[[loopI]] <- plot_box(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'GCMS',
                                       groupName = TRUE)
}

#Plant-groupped PLS fitted values
plot_PLS_Fitted_Plant <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_PLS_Fitted_Plant)) {
  plot_PLS_Fitted_Plant[[loopI]] <- plot_box_plant(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'PLS')
}

#Plant-groupped GCMS signals
plot_GCMS_Plant <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_GCMS_Plant)) {
  plot_GCMS_Plant[[loopI]] <- plot_box_plant(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[2]],
                                       variableName = 'GCMS')
}

```

```{r}
#| label: overview on GCMS and FTIR - 2nd derivative spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| echo: false
#| output: true
#| warning: false
#| message: false

(plot_GCMS[[1]]|plot_PLS_Fitted[[1]]) /
  (plot_GCMS_Materials[[1]]|plot_PLS_Fitted_Materials[[1]]) /
  (plot_GCMS_Plant[[1]]|plot_PLS_Fitted_Plant[[1]])

#Output as pdf
pdf(file = file.path('Figures_R', 'Boxplot_Comp.pdf'),   # The directory you want to save the file in
    width = 8, # The width of the plot in inches
    height = 12)

(plot_GCMS[[1]]|plot_PLS_Fitted[[1]]) /
  (plot_GCMS_Materials[[1]]|plot_PLS_Fitted_Materials[[1]]) /
  (plot_GCMS_Plant[[1]]|plot_PLS_Fitted_Plant[[1]])
  
dev.off()
```

## Malformation rates of pollen grains

Counting results on malformation rate were input and merged with GC-MS and FTIR signals. An independent figure were plotted here to show the variations of malformation rates among donor plants.

```{r}
#| label: preperation on malformation plotting
#| echo: false
#| output: false
#| warning: false
#| message: false
#Calculate malformation rates in pot cultivars
tempVec <- (cultivarsRaw[, 15] - cultivarsRaw[, 2])/ cultivarsRaw[, 2]
cultivarMean <- vector(length = 4)
cultivarMean[1] <- mean(tempVec[which(cultivarsBatchInfo[, 4] == 'UV')])
cultivarMean[2] <- sd(tempVec[which(cultivarsBatchInfo[, 4] == 'UV')])
cultivarMean[3] <- mean(tempVec[which(cultivarsBatchInfo[, 4] == 'CON')])
cultivarMean[4] <- sd(tempVec[which(cultivarsBatchInfo[, 4] == 'CON')])

#Point plot to compare ratios between con and uv treatment, on different plants
cuttingsPlantsPlot <-
  ggplot() +
  geom_point(data = data_preprocessed_ftir_gcms_malformation[[1]][[3]],
             aes(x = CON_Mal, y = UV_Mal, col = Plant)) +
  geom_point(aes(x = cultivarMean[3] , y = cultivarMean[1], 
                 color = 'Cultivars')) +
  geom_errorbar(data = data_preprocessed_ftir_gcms_malformation[[1]][[3]],
                aes(x = CON_Mal, y = UV_Mal, col = Plant,
                    ymin = UV_Mal - UV_Mal_sd,
                    ymax = UV_Mal + UV_Mal_sd),
                linewidth = 1,
                width = 0.01) +
  geom_errorbar(data = data_preprocessed_ftir_gcms_malformation[[1]][[3]],
                aes(x = CON_Mal, y = UV_Mal, col = Plant,
                    xmin = CON_Mal - CON_Mal_sd,
                    xmax = CON_Mal + CON_Mal_sd),
                linewidth = 1,
                width = 0.01,
                orientation = 'y') +
  geom_errorbar(aes(x = cultivarMean[3] , y = cultivarMean[1], color = 'Cultivars',
                    ymin = cultivarMean[1] - cultivarMean[2],
                    ymax = cultivarMean[1] + cultivarMean[2]),
                linewidth = 1,
                width = 0.005) +
  geom_errorbar(aes(x = cultivarMean[3] , y = cultivarMean[1], color = 'Cultivars',
                    xmin = cultivarMean[3] - cultivarMean[4],
                    xmax = cultivarMean[3] + cultivarMean[4]),
                linewidth = 1,
                width = 0.005,
                orientation = 'y') +
  geom_abline (slope=1, linetype = "dashed", color="grey") +
  scale_x_continuous(limits = c(-0.05, 0.15), name = 'CON malformation percentage',
                     breaks = seq(0, 0.15, by = 0.05)) +
  scale_y_continuous(limits = c(-0.05, 0.25), name = 'UV malformation percentage',
                     breaks = seq(0, 0.25, by = 0.05)) +
  scale_color_manual(name = 'Plant',
                     values = c('1' = brewer.pal(6, 'Set2')[1],
                                '2' = brewer.pal(6, 'Set2')[2],
                                '3' = brewer.pal(6, 'Set2')[3],
                                '4' = brewer.pal(6, 'Set2')[4],
                                '5' = brewer.pal(6, 'Set2')[5],
                                '6' = brewer.pal(6, 'Set2')[6],
                                'Cultivars' = 'black')) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    axis.title.x = element_text(size = 20, family = 'arial'),
    axis.text.x = element_text(size = 16, family = 'arial'),
    
    axis.text.y = element_text(size = 20, family = 'arial'),
    axis.title.y = element_text(size = 16, family = 'arial'),
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.2),
    legend.text = element_text(size = 14, family = 'arial')
  )

rm(tempVec)
```

```{r}
#| label: malformation - figure
#| fig-width: 8
#| fig-height: 8
#| fig-dpi: 600
#| echo: false
#| output: true
#| warning: false
#| message: false
cuttingsPlantsPlot
```

## Student-t analysis on GC-MS signals and PLS fitted values from FTIR spectra

For signals between different subgroups (UV/CONTROL, plant 1-6, cuttings/cultivars), student-t test is applied to describe the difference between two grouping creteria.

```{r}
#| label: Student-t test on GC-MS and FTIR PLS fitted values
#| echo: false
#| output: false
#| warning: false
#| message: false
#Adjust the GCMS data frame to indicate the material types in the last column 
GCDataMeanAdj <- GCData[, -6]
GCDataMeanAdj[which(!is.na(GCData[, 5])), 5] <- 'cut'
GCDataMeanAdj[which(is.na(GCData[, 5])), 5] <- 'cul'

#Remve whitespaces in GCMS and PLS data frame
GCDataMeanAdj <- apply(GCDataMeanAdj, c(1,2), gsub, pattern = "\\s+", replacement = "")
colnames(GCDataMeanAdj)[4:5] <- c('Plant', 'Materials')
GCDataMeanAdj <- as.data.frame(GCDataMeanAdj)
GCDataMeanAdj[, 1:2] <- apply(GCDataMeanAdj[, 1:2], c(1,2), as.numeric)

PLSFTIRAdj <- data_preprocessed_ftir_gcms_malformation[[1]][[1]]
PLSFTIRAdj <- apply(PLSFTIRAdj, c(1,2), gsub, pattern = "\\s+", replacement = "")
colnames(PLSFTIRAdj)[8] <- c('Plant')
PLSFTIRAdj <- as.data.frame(PLSFTIRAdj)
PLSFTIRAdj[, 11:12] <- apply(PLSFTIRAdj[, 11:12], c(1,2), as.numeric)

#Setup the information to be input into student-t analysis
studTDataList <- list(
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    GCDataMeanAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj,
    PLSFTIRAdj
)
studTDataListName <- c(
  'UV and Control - GCMS',
  'UV - Potted cultivars and branch cuttings - GCMS',
  'Control - Potted cultivars and branch cuttings - GCMS',
  'Potted cultivars - UV and Control - GCMS',
  'Branch cuttings - UV and Control - GCMS',
  'Donor plant 1 - UV and Control - GCMS',
  'Donor plant 2 - UV and Control - GCMS',
  'Donor plant 3 - UV and Control - GCMS',
  'Donor plant 4 - UV and Control - GCMS',
  'Donor plant 5 - UV and Control - GCMS',
  'Donor plant 6 - UV and Control - GCMS',
  'UV and Control - FTIR',
  'UV - Potted cultivars and branch cuttings - FTIR',
  'Control - Potted cultivars and branch cuttings - FTIR',
  'Potted cultivars - UV and Control - FTIR',
  'Branch cuttings - UV and Control - FTIR',
  'Donor plant 1 - UV and Control - FTIR',
  'Donor plant 2 - UV and Control - FTIR',
  'Donor plant 3 - UV and Control - FTIR',
  'Donor plant 4 - UV and Control - FTIR',
  'Donor plant 5 - UV and Control - FTIR',
  'Donor plant 6 - UV and Control - FTIR'
  )
studTVarNmae <- c(
  rep('calibrated', 11),
  rep('PLS', 11)
)
studTCreName1 <- c(
  rep('Treatment', 3),
  rep('Materials', 2),
  rep('Plant', 6),
  rep('Treatment', 3),
  rep('Materials', 2),
  rep('Plant', 6)
)
studTCre1 <- list(
  c('UV', 'CON'),
  'UV',
  'CON',
  'cul',
  'cut',
  '1',
  '2',
  '3',
  '4',
  '5',
  '6',
  c('UV', 'CON'),
  'UV',
  'CON',
  'cul',
  'cut',
  '1',
  '2',
  '3',
  '4',
  '5',
  '6'
)
studTCreName2 <- c(
  NA,
  rep('Materials', 2),
  rep('Treatment', 2),
  rep('Materials', 6),
  NA,
  rep('Materials', 2),
  rep('Treatment', 2),
  rep('Materials', 6)
)
studTCre2 <- list(
  NA,
  c('cul', 'cut'),
  c('cul', 'cut'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  'cut',
  'cut',
  'cut',
  'cut',
  'cut',
  'cut',
  NA,
  c('cul', 'cut'),
  c('cul', 'cut'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  'cut',
  'cut',
  'cut',
  'cut',
  'cut',
  'cut'
)
studTCreName3 <- c(
  rep(NA, 5),
  rep('Treatment', 6),
  rep(NA, 5),
  rep('Treatment', 6)
)
studTCre3 <- list(
  NA,
  NA,
  NA,
  NA,
  NA,
  c('UV', 'CON'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  NA,
  NA,
  NA,
  NA,
  NA,
  c('UV', 'CON'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  c('UV', 'CON'),
  c('UV', 'CON')
)

#Student-t analysis
studTAnaResuTab <- sig_test_stu_t(studTDataList,
                                  studTDataListName,
                                  studTVarNmae,
                                  studTCreName1,
                                  studTCre1,
                                  studTCreName2,
                                  studTCre2,
                                  studTCreName3,
                                  studTCre3)
```

```{r}
#| label: show the significance test results
knitr::kable(studTAnaResuTab)

write.csv(studTAnaResuTab, file = file.path('Figures_R', 'studTAnaResuTab.csv'))
```

## Donor plant comparison - mean and variance of malformation rates/GCMS/FTIR signals

The malformation rate indicates plant's stability of mitosis process, including both its mean and sd. The GCMS/FTIR signals indicate the quantification of UAC within pollen grains, positively correlated to screening ability towards UV-B radiation. This section aims at exploring the relationship between malformation rate and GCMS/FTIR signals, by looking at their means and variances, with assumption that accumulation of UAC may contribute to more stable mitosis process.

```{r}
#| label: Mean and sd comparision
#| echo: false
#| output: false
#| warning: false
#| message: false
#Remve whitespaces in GCMS and PLS data frame
data_preprocessed_ftir_gcms_malformation[[1]][[2]] <- apply(data_preprocessed_ftir_gcms_malformation[[1]][[2]], c(1,2), gsub, pattern = "\\s+", replacement = "")
colnames(data_preprocessed_ftir_gcms_malformation[[1]][[2]])[8] <- c('Plant')
data_preprocessed_ftir_gcms_malformation[[1]][[2]] <- as.data.frame(data_preprocessed_ftir_gcms_malformation[[1]][[2]])
data_preprocessed_ftir_gcms_malformation[[1]][[2]][, 14] <- as.numeric(data_preprocessed_ftir_gcms_malformation[[1]][[2]][, 14])

#Setup the information to be input into student-t analysis
meanSDCompDataList <- list(
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    data_preprocessed_ftir_gcms_malformation[[1]][[2]][which(data_preprocessed_ftir_gcms_malformation[[1]][[2]]$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    GCDataMeanAdj[which(GCDataMeanAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),],
    PLSFTIRAdj[which(PLSFTIRAdj$Materials == 'cut'),]
)
meanSDCompDataListName <- c(
  'Malformation - all',
  'Malformation - UV',
  'Malformation - Control',
  'Malformation - UV - Plant 1',
  'Malformation - UV - Plant 2',
  'Malformation - UV - Plant 3',
  'Malformation - UV - Plant 4',
  'Malformation - UV - Plant 5',
  'Malformation - UV - Plant 6',
  'Malformation - CON - Plant 1',
  'Malformation - CON - Plant 2',
  'Malformation - CON - Plant 3',
  'Malformation - CON - Plant 4',
  'Malformation - CON - Plant 5',
  'Malformation - CON - Plant 6',
  'GCMS - all',
  'GCMS - UV',
  'GCMS - Control',
  'GCMS - UV - Plant 1',
  'GCMS - UV - Plant 2',
  'GCMS - UV - Plant 3',
  'GCMS - UV - Plant 4',
  'GCMS - UV - Plant 5',
  'GCMS - UV - Plant 6',
  'GCMS - CON - Plant 1',
  'GCMS - CON - Plant 2',
  'GCMS - CON - Plant 3',
  'GCMS - CON - Plant 4',
  'GCMS - CON - Plant 5',
  'GCMS - CON - Plant 6',
  'FTIR - all',
  'FTIR - UV',
  'FTIR - Control',
  'FTIR - UV - Plant 1',
  'FTIR - UV - Plant 2',
  'FTIR - UV - Plant 3',
  'FTIR - UV - Plant 4',
  'FTIR - UV - Plant 5',
  'FTIR - UV - Plant 6',
  'FTIR - CON - Plant 1',
  'FTIR - CON - Plant 2',
  'FTIR - CON - Plant 3',
  'FTIR - CON - Plant 4',
  'FTIR - CON - Plant 5',
  'FTIR - CON - Plant 6'
  )
meanSDCompVarNmae <- c(
  rep('Mal_Ratio', 15),
  rep('calibrated', 15),
  rep('PLS', 15)
)
meanSDCompCreName1 <- c(
  NA,
  rep('Treatment', 14),
  NA,
  rep('Treatment', 14),
  NA,
  rep('Treatment', 14)
)
meanSDCompCre1 <- c(
  NA,
  c('UV', 'CON'),
  rep('UV', 6),
  rep('CON', 6),
  NA,
  c('UV', 'CON'),
  rep('UV', 6),
  rep('CON', 6),
  NA,
  c('UV', 'CON'),
  rep('UV', 6),
  rep('CON', 6)
)
meanSDCompCreName2 <- c(
  rep(NA, 3),
  rep('Plant', 12),
  rep(NA, 3),
  rep('Plant', 12),
  rep(NA, 3),
  rep('Plant', 12)
)
meanSDCompCre2 <- c(
  rep(NA, 3),
  rep(as.character(1:6), 2),
  rep(NA, 3),
  rep(as.character(1:6), 2),
  rep(NA, 3),
  rep(as.character(1:6), 2)
)

meanSDCompMat <- mean_sd_com(meanSDCompDataList,
                             meanSDCompDataListName,
                             meanSDCompVarNmae,
                             meanSDCompCreName1,
                             meanSDCompCre1,
                             meanSDCompCreName2,
                             meanSDCompCre2)

#Adjust the matrix into wide one
meanSDCompMatAdj <- matrix(nrow = 45,
                           ncol = 5,
                           dimnames = list(c(),
                                           c('Calculation_Method',
                                             'Treatment',
                                             'Plant',
                                             'Mean',
                                             'SD')))
meanSDCompMatAdj[, 1] <- c(
  rep('Malformation', 15),
  rep('GCMS', 15),
  rep('FTIR-PLS', 15)
)
meanSDCompMatAdj[, 2] <- c(
  NA,
  c('UV', 'CON'),
  rep('UV', 6),
  rep('CON', 6),
  NA,
  c('UV', 'CON'),
  rep('UV', 6),
  rep('CON', 6),
  NA,
  c('UV', 'CON'),
  rep('UV', 6),
  rep('CON', 6)
)
meanSDCompMatAdj[, 3] <- c(
  rep(NA, 3),
  rep(as.character(1:6), 2),
  rep(NA, 3),
  rep(as.character(1:6), 2),
  rep(NA, 3),
  rep(as.character(1:6), 2)
)
meanSDCompMatAdj[, 4:5] <- meanSDCompMat[, 2:3]

meanSDCompMatAdjPlant <- cbind(meanSDCompMatAdj[4:15,2:4],
                               Malformation_SD = meanSDCompMatAdj[4:15,5],
                               GCMS = meanSDCompMatAdj[19:30,4],
                               GCMS_SD = meanSDCompMatAdj[19:30,5],
                               FTIR = meanSDCompMatAdj[34:45,4],
                               FTIR_SD = meanSDCompMatAdj[34:45,5])
colnames(meanSDCompMatAdjPlant)[3] <- 'Malformation'
meanSDCompMatAdjPlant <- as.data.frame(meanSDCompMatAdjPlant)
meanSDCompMatAdjPlant[,3:8] <- apply(meanSDCompMatAdjPlant[, 3:8], c(1,2), as.numeric)
meanSDCompMatAdjPlantDiff <- cbind(Plant = meanSDCompMatAdjPlant[1:6, 2],
                                   Malformation = meanSDCompMatAdjPlant[1:6,3] - 
                                     meanSDCompMatAdjPlant[7:12,3],
                                   Malformation_SD = meanSDCompMatAdjPlant[1:6,4] - 
                                     meanSDCompMatAdjPlant[7:12,4],
                                   GCMS = meanSDCompMatAdjPlant[1:6,5] - 
                                     meanSDCompMatAdjPlant[7:12,5],
                                   GCMS_SD = meanSDCompMatAdjPlant[1:6,6] - 
                                     meanSDCompMatAdjPlant[7:12,6],
                                   FTIR = meanSDCompMatAdjPlant[1:6,7] - 
                                     meanSDCompMatAdjPlant[7:12,7],
                                   FTIR_SD = meanSDCompMatAdjPlant[1:6,8] - 
                                     meanSDCompMatAdjPlant[7:12,8])
meanSDCompMatAdjPlantDiff <- as.data.frame(meanSDCompMatAdjPlantDiff)
meanSDCompMatAdjPlantDiff[,2:7] <- apply(meanSDCompMatAdjPlantDiff[2:7], c(1,2), as.numeric)
meanSDCompMatAdjPlantDiffSca <- meanSDCompMatAdjPlantDiff
meanSDCompMatAdjPlantDiffSca[, -1] <- apply(meanSDCompMatAdjPlantDiffSca[,-1], 2, scale)

#Scatter plot on malformation rate and GCMS signal, sub-groupped by donor plant
plot_mean_com_mal_gcms <- ggplot(meanSDCompMatAdjPlant,
                                 aes(x = GCMS,
                                     y = Malformation)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = paste(after_stat(eq.label), 
                                   after_stat(rr.label), 
                                   sep = "*\", \"*")), 
                 formula = y ~ x, parse = TRUE) +
    labs(x = 'GC-MS (ng/150gr)',
         y = 'Malformation rate') + 
    theme_classic() + 
    theme(plot.title = element_blank(), 
          axis.title.y = element_text(size = 14, family = 'arial'),
          axis.text.y = element_text(size = 12, family = 'arial'),
          axis.title.x = element_text(size = 14, family = 'arial'),
          axis.text.x = element_text(size = 12, family = 'arial'),
          panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
    )

plot_mean_com_mal_gcms_diff <-
  ggplot(meanSDCompMatAdjPlantDiff,
                                 aes(x = GCMS,
                                     y = Malformation)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = paste(after_stat(eq.label), 
                                   after_stat(rr.label), 
                                   sep = "*\", \"*")), 
                 formula = y ~ x, parse = TRUE) +
    labs(x = 'GC-MS (ng/150gr) difference between UV and CON',
         y = 'Malformation rate difference between UV and CON') + 
    theme_classic() + 
    theme(plot.title = element_blank(), 
          axis.title.y = element_text(size = 14, family = 'arial'),
          axis.text.y = element_text(size = 12, family = 'arial'),
          axis.title.x = element_text(size = 14, family = 'arial'),
          axis.text.x = element_text(size = 12, family = 'arial'),
          panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
    )

plot_mean_com_mal_ftir <- ggplot(meanSDCompMatAdjPlant,
                                 aes(x = FTIR,
                                     y = Malformation)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = paste(after_stat(eq.label), 
                                   after_stat(rr.label), 
                                   sep = "*\", \"*")), 
                 formula = y ~ x, parse = TRUE) +
    labs(x = 'PLS fitted values',
         y = 'Malformation rate') + 
    theme_classic() + 
    theme(plot.title = element_blank(), 
          axis.title.y = element_text(size = 14, family = 'arial'),
          axis.text.y = element_text(size = 12, family = 'arial'),
          axis.title.x = element_text(size = 14, family = 'arial'),
          axis.text.x = element_text(size = 12, family = 'arial'),
          panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
    )

plot_mean_com_mal_ftir_diff <-
  ggplot(meanSDCompMatAdjPlantDiff,
                                 aes(x = FTIR,
                                     y = Malformation)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = paste(after_stat(eq.label), 
                                   after_stat(rr.label), 
                                   sep = "*\", \"*")), 
                 formula = y ~ x, parse = TRUE) +
    labs(x = 'GC-MS (ng/150gr) difference between UV and CON',
         y = 'Malformation rate difference between UV and CON') + 
    theme_classic() + 
    theme(plot.title = element_blank(), 
          axis.title.y = element_text(size = 14, family = 'arial'),
          axis.text.y = element_text(size = 12, family = 'arial'),
          axis.title.x = element_text(size = 14, family = 'arial'),
          axis.text.x = element_text(size = 12, family = 'arial'),
          panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
    )

plot_scatter_mean_sd_mal <-   
  ggplot(meanSDCompMatAdjPlantDiffSca) +
  geom_point(aes(x = Malformation,
             y = Malformation_SD,
             color = Plant),
             size = 6
             ) +
  geom_hline(yintercept = 0, color = "black") + 
  geom_vline(xintercept = 0, color = "black") +
  scale_x_continuous(limits = c(-2,3)) +
  scale_y_continuous(limits = c(-2,3)) +
  scale_colour_discrete(palette = 'Set2') +
  labs(x = 'Malformation mean rate difference between UV and CON',
       y = 'Malformation rate variance difference between UV and CON') + 
  theme_classic() + 
  theme(plot.title = element_blank(), 
        axis.title.y = element_text(size = 14, family = 'arial'),
        axis.text.y = element_text(size = 12, family = 'arial'),
        axis.title.x = element_text(size = 14, family = 'arial'),
        axis.text.x = element_text(size = 12, family = 'arial'),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
  )

plot_scatter_mean_sd_gcms <-   
  ggplot(meanSDCompMatAdjPlantDiffSca) +
  geom_point(aes(x = GCMS,
             y = GCMS_SD,
             color = Plant),
             size = 6
             ) +
  geom_hline(yintercept = 0, color = "black") + 
  geom_vline(xintercept = 0, color = "black") +
  scale_x_continuous(limits = c(-2,3)) +
  scale_y_continuous(limits = c(-2,3)) +
  scale_colour_discrete(palette = 'Set2') +
  labs(x = 'GCMS signal difference between UV and CON',
       y = 'GCMS signal variance difference between UV and CON') + 
  theme_classic() + 
  theme(plot.title = element_blank(), 
        axis.title.y = element_text(size = 14, family = 'arial'),
        axis.text.y = element_text(size = 12, family = 'arial'),
        axis.title.x = element_text(size = 14, family = 'arial'),
        axis.text.x = element_text(size = 12, family = 'arial'),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
  )

plot_scatter_mean_sd_ftir <-   
  ggplot(meanSDCompMatAdjPlantDiffSca) +
  geom_point(aes(x = FTIR,
             y = FTIR_SD,
             color = Plant),
             size = 6
             ) +
  geom_hline(yintercept = 0, color = "black") + 
  geom_vline(xintercept = 0, color = "black") +
  scale_x_continuous(limits = c(-2,3)) +
  scale_y_continuous(limits = c(-2,3)) +
  scale_colour_discrete(palette = 'Set2') +
  labs(x = 'PLS fitted value difference between UV and CON',
       y = 'PLS fitted value variance difference between UV and CON') + 
  theme_classic() + 
  theme(plot.title = element_blank(), 
        axis.title.y = element_text(size = 14, family = 'arial'),
        axis.text.y = element_text(size = 12, family = 'arial'),
        axis.title.x = element_text(size = 14, family = 'arial'),
        axis.text.x = element_text(size = 12, family = 'arial'),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1)
  )
```

```{r}
#| label: Mean and sd comparision - figure
#| fig-width: 16
#| fig-height: 16
#| fig-dpi: 600
#| echo: false
#| output: true
#| warning: false
#| message: false
(plot_mean_com_mal_gcms | plot_mean_com_mal_ftir) /
  (plot_scatter_mean_sd_mal | plot_scatter_mean_sd_gcms)

#Output as pdf
pdf(file = file.path('Figures_R', 'Signal_Mal_Comp.pdf'),   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 6)

(plot_mean_com_mal_gcms | plot_mean_com_mal_ftir)
  
dev.off()
```
