---
title: "UAC signals in Pine cuttings and cultivars"
format:
  html:
    code-fold: true
    code-tools: true
editor: visual
---

```{r}
#| label: load Packages
#| echo: false
#| output: false
#| warning: false
#| message: false
#Load palaeochem package
library(palaeochem)
#Load arial font into R
library(showtext)
font_add(family = "arial", regular = file.path('Packages_R', 'Fonts', 'arial.ttf'))
showtext_auto()

#Packages needed in Vivian's code
source(
  file.path(
    "Code_R", "00_Config_file.r"
  )
)
```

```{r}
#| label: load FTIRS pectra
#| echo: true
#| warning: false

#Read raw FTIR spectra
rawFTIRSpec <- read.csv(file = file.path('Data_R',
                                         'Original_Spectra.csv'),
                        row.names = 1)
colnames(rawFTIRSpec)[-1:-10] <- substring(colnames(rawFTIRSpec)[-1:-10], 2)
```

## Original FTIR spectra

The original FTIR spectra are between 4000 and 600 cm^-1^, with wavenumber interval of 4 cm^-1^. And the raw spectra from potted cultivars and branch cuttings looks like:

```{r}
#| label: Figure full raw spectra
#| echo: false
#| warning: false

plot_original_spec_cul <-
rawFTIRSpec[rawFTIRSpec[, 5] == 'cul', c(-1:-2,-4:-10)] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
  labs(title = 'Potted cultivars',
       x = '')

plot_original_spec_cut <-
rawFTIRSpec[rawFTIRSpec[, 5] == 'cut', c(-1:-2,-4:-10)] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
    labs(title = 'Branch cuttings')

plot_original_spec_cul /
  plot_original_spec_cut

```

## Pretreating raw FTIR spectra

```{r}
#| label: prepare for different pretreating methodologies
#| echo: true
#| warning: false

#1 - smoothing + normalisation + baseline offset
processing_steps_ftir1 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers,
    get_area_normalisation,
    get_baseline_offset
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(min_wn = 800, max_wn = 1800),
    list(method = "area"),
    list()
  )
)

#2 - smoothing + normalisation + MSC + baseline offset
processing_steps_ftir2 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers,
    get_area_normalisation,
    run_msc,
    get_baseline_offset
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(min_wn = 800, max_wn = 1800),
    list(method = "area"),
    list(),
    list()
  )
)

#3 - smoothing + normalisation + ALS + baseline offset
processing_steps_ftir3 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers,
    get_area_normalisation,
    run_als_baseline_correction,
    get_baseline_offset
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(min_wn = 800, max_wn = 1800),
    list(method = "area"),
    list(lambda = 5, p = 0.001),
    list()
  )
)

#8 - smoothing + nomralisation + ALS + MSC + baseline offset
processing_steps_ftir4 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers,
    get_area_normalisation,
    run_als_baseline_correction,
    run_msc,
    get_baseline_offset
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 0),
    list(min_wn = 800, max_wn = 1800),
    list(method = "area"),
    list(lambda = 5, p = 0.001),
    list(),
    list()
  )
)

#5 - smoothing + 2nd derivative
processing_steps_ftir5 <- list(
  functions = c(
    get_savitzkygolay,
    get_filter_wavenumbers
  ),
  parameters = list(
    list(width = 7, poly = 2, deriv = 2),
    list(min_wn = 800, max_wn = 1800)
  )
)
```

```{r}
#| label: pretreat the raw FTIR spectra with different methodologies
#| echo: true
#| warning: false

rawFTIRSpecMat <- apply(as.matrix(rawFTIRSpec[, -1:-10]), c(1,2), as.numeric)
row.names(rawFTIRSpecMat) <- rawFTIRSpec[, 3]
data_preprocessed <- list(
  cbind(
    unique_id = rawFTIRSpec[, 3],
    corrected1 = run_preprocessing_list(rawFTIRSpec[, c(-1:-2,-4:-10)], processing_steps_ftir1)
  ),
  cbind(
    unique_id = rawFTIRSpec[, 3],
    corrected2 = run_preprocessing_list(rawFTIRSpecMat, processing_steps_ftir2)
  ),
  cbind(
    unique_id = rawFTIRSpec[, 3],
    corrected3 = run_preprocessing_list(rawFTIRSpec[, c(-1:-2,-4:-10)], processing_steps_ftir3)
  ),
  cbind(
    unique_id = rawFTIRSpec[, 3],
    corrected4 = run_preprocessing_list(rawFTIRSpec[, c(-1:-2,-4:-10)], processing_steps_ftir4)
  ),
  cbind(
    unique_id = rawFTIRSpec[, 3],
    corrected5 = run_preprocessing_list(rawFTIRSpec[, c(-1:-2,-4:-10)], processing_steps_ftir5)
  )
)
data_preprocessed <- lapply(data_preprocessed, as.data.frame)
for (loopI in 1:length(data_preprocessed)) {
  data_preprocessed[[loopI]][, -1] <-
    apply(data_preprocessed[[loopI]][, -1], c(1,2), as.numeric)
}

rm(rawFTIRSpecMat)
```

Several methodologies were applied on the raw FTIR spectra to search for better methodologies, including Multiplicative Scattering Correction (MSC), Asymmetric least squares (ALS) baseline correction and 2nd derivative spectra. Smoothing and baseline offset deduction are also applied on raw spectra. Two types of wavenumber ranges were investigated as well - the whole range (600-4000) and signature area only (800-1000).

### Signature range: smoothing + normalised by peak area + baseline offset

```{r}
#| label: signature raw
#| echo: true
#| warning: false

plot_raw_spec_cul <-
data_preprocessed[[1]][rawFTIRSpec[, 5] == 'cul',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
  labs(title = 'Potted cultivars',
       x = '')

plot_raw_spec_cut <-
data_preprocessed[[1]][rawFTIRSpec[, 5] == 'cut',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
    labs(title = 'Branch cuttings')

plot_raw_spec_cul /
  plot_raw_spec_cut
```

### Signature range: smoothing + normalised by peak area + MSC + baseline offset

```{r}
#| label: signature MSC
#| echo: true
#| warning: false

plot_msc_spec_cul <-
data_preprocessed[[2]][rawFTIRSpec[, 5] == 'cul',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
  labs(title = 'Potted cultivars',
       x = '')

plot_msc_spec_cut <-
data_preprocessed[[2]][rawFTIRSpec[, 5] == 'cut',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
    labs(title = 'Branch cuttings')

plot_msc_spec_cul /
  plot_msc_spec_cut
```

### Signature range: smoothing + normalised by peak area + ALS + baseline offset

```{r}
#| label: signature ALS
#| echo: true
#| warning: false
plot_als_spec_cul <-
data_preprocessed[[3]][rawFTIRSpec[, 5] == 'cul',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
  labs(title = 'Potted cultivars',
       x = '')

plot_als_spec_cut <-
data_preprocessed[[3]][rawFTIRSpec[, 5] == 'cut',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
    labs(title = 'Branch cuttings')

plot_als_spec_cul /
  plot_als_spec_cut
```

### Signature range: smoothing + normalised by peak area + ALS + MSC + baseline offset

```{r}
#| label: signature ALS MSC
#| echo: true
#| warning: false
plot_als_msc_spec_cul <-
data_preprocessed[[4]][rawFTIRSpec[, 5] == 'cul',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
  labs(title = 'Potted cultivars',
       x = '')

plot_als_msc_spec_cut <-
data_preprocessed[[4]][rawFTIRSpec[, 5] == 'cut',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
    labs(title = 'Branch cuttings')

plot_als_msc_spec_cul /
  plot_als_msc_spec_cut
```

### Signature range: smoothing + 2^nd^ derivative

```{r}
#| label: signature 2nd derivative
#| echo: true
#| warning: false
plot_2nd_spec_cul <-
data_preprocessed[[5]][rawFTIRSpec[, 5] == 'cul',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
  labs(title = 'Potted cultivars',
       x = '')

plot_2nd_spec_cut <-
data_preprocessed[[5]][rawFTIRSpec[, 5] == 'cut',] |> plot_spectra(
  group_var = 'unique_id',
  col_var = 'unique_id'
) +
    labs(title = 'Branch cuttings')

plot_2nd_spec_cul /
  plot_2nd_spec_cut
```

## Input GCMS signals

```{r}
#| label: read GCMS
#| echo: false
#| warning: false
#GC-MS data on cuttings and pot cultivars
GCDataRawCut <- as.matrix(read.csv(
  file.path('Data_R', '04_11_2025_gc_cutting_data_2025.csv'),
  header = TRUE,
  row.names = 1))
GCDataRawCul <- as.matrix(read.csv(
  file.path('Data_R', '01_07_2025_gc_cultivar_data_2025.csv'),
  header = TRUE,
  row.names = 1))

#Combine these two raw datasets, with batch information
GCData <- rbind(GCDataRawCut[, c(8, 10:14)],
                cbind(GCDataRawCul[, c(8, 10:12)], 
                      sampling_number = NA, 
                      GCDataRawCul[, 13])
)

#Exclude the last batch (u), as an unexpected crash happened before this
#batch, making reseult from this batch abnormal
GCData <- GCData[-which(GCData[, 6] == 'u'),]

#Loop to calculate the mean values for each sample
uniqueBatchInfo <- unique(GCData[, 2:5])

#Create a matrix to fill in the mean GC-MS signals
GCDataMean <- matrix(nrow = nrow(uniqueBatchInfo),
                     ncol = ncol(uniqueBatchInfo) + 2,
                     dimnames = list(c(),
                                     c('calibrated',
                                       'sd',
                                       colnames(uniqueBatchInfo)
                                     )
                     )
)

#Add batch number to the GCMS data matrix
GCDataMean[, 3:ncol(GCDataMean)] <- uniqueBatchInfo

#Loop to calculate the mean signals of GCMS duplicates
for (loopI in 1:nrow(GCDataMean)) {
  
  #Check wheter current sample is cutting or pot cultivar 
  if (is.na(GCDataMean[loopI, 6])) {
    
    #Mean and sd for pot cultivar
    GCDataMean[loopI, 1] <- mean(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) 
    )
    , 1])
    )
    GCDataMean[loopI, 2] <- sd(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5])
    )
    , 1])
    )
    
  }else {
    
    #Mean and sd for cutting
    GCDataMean[loopI, 1] <- mean(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) &
                     (GCData[, 5] == GCDataMean[loopI, 6])
    )
    , 1])
    )
    GCDataMean[loopI, 2] <- sd(as.numeric(
      GCData[which((GCData[, 2] == GCDataMean[loopI, 3]) &
                     (GCData[, 3] == GCDataMean[loopI, 4]) &
                     (GCData[, 4] == GCDataMean[loopI, 5]) &
                     (GCData[, 5] == GCDataMean[loopI, 6])
    )
    , 1])
    )
    
  }
}

#Transfer the matrix back to data frame, and transfer values to number if necessary
GCDataMean <- as.data.frame(GCDataMean)
GCDataMean[, 1] <- as.numeric(GCDataMean[, 1])

#Create a vector to stroe the index for batch information matching
tempInd <- vector('numeric', length = nrow(rawFTIRSpec))

#Loop to match the batch information between GC-MS and FTIR datasets
for (loopI in 1:nrow(rawFTIRSpec)) {

#Check whether current sample is pot cultivar
if (rawFTIRSpec[loopI, 5] == 'cul') {
  
  #Check whether current FTIR spectrum has any GCMS signal
  if (any((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
          (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
          (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
          (is.na(GCDataMean[, 6]))
  )
  ) {
    
    #Record the FTIR matrix row index for current GCMS signal
    tempInd[loopI] <- which((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
                              (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
                              (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
                              (is.na(GCDataMean[, 6])))
    
  }else {
    
    #Record nothing if there is no match between current FTIR specrum and
      #any GCMS signal
    tempInd[loopI] <- NA
    
  }
  
#Check whether current sample is cutting
} else if (rawFTIRSpec[loopI, 5] == 'cut') {
  
  #Check whether current FTIR spectrum has any GCMS signal
  if (any((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
          (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
          (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
          (GCDataMean[, 6] == rawFTIRSpec[loopI, 9])
  )
  ) {
    
    #Record the FTIR matrix row index for current GCMS signal
    tempInd[loopI] <- which((GCDataMean[, 3] == rawFTIRSpec[loopI, 6]) &
                              (GCDataMean[, 4] == rawFTIRSpec[loopI, 7]) &
                              (GCDataMean[, 5] == rawFTIRSpec[loopI, 8]) &
                              (GCDataMean[, 6] == rawFTIRSpec[loopI, 9]))
    
  }else {
    
    #Record nothing if there is no match between current FTIR spectrum and
      #any GCMS signal
    tempInd[loopI] <- NA
    
  }
  
} else {
  
  #Record nothing if there is no match between current FTIR spectrum and
    #any GCMS signal
  tempInd[loopI] <- NA
  
}
}

#Attach GCMS signals to  peak statistics (height or area)
pcaMatHeight <- cbind(rawFTIRSpec,
                    GCDataMean[tempInd, c(1,2)])[, -11:-1772]
colnames(pcaMatHeight)[11] <- 'pCA_ng_grain'
pcaMatHeight <- pcaMatHeight[!is.na(pcaMatHeight[, 11]),]

rm(loopI,
   tempInd)
```

## PLS analysis between FTIR spectra and GCMS signals

In this section, the GCMS signals were input into PLS model as response variable, and FTIR spectral signals pretreated by different methodologies were input as independent variables. The decision on the number of component is made by function pls::selectNcomp, with the method of randomization.

```{r}
#| label: run PLS
#| echo: false
#| warning: false

#Check whetether there is already a RData file containing the PLS model, skip the lines generating the PLS model and read the archives driectly if true
if (file.exists(file.path('Output_R', 'pls_sig_raw_cul.RData'))) {
  load(file.path('Output_R', 'pls_sig_raw_cul.RData'))
  load(file.path('Output_R', 'pls_sig_msc_cul.RData'))
  load(file.path('Output_R', 'pls_sig_als_cul.RData'))
  load(file.path('Output_R', 'pls_sig_als_msc_cul.RData'))
  load(file.path('Output_R', 'pls_sig_2nd_cul.RData'))
  
  load(file.path('Output_R', 'pls_sig_raw_cut.RData'))
  load(file.path('Output_R', 'pls_sig_msc_cut.RData'))
  load(file.path('Output_R', 'pls_sig_als_cut.RData'))
  load(file.path('Output_R', 'pls_sig_als_msc_cut.RData'))
  load(file.path('Output_R', 'pls_sig_2nd_cut.RData'))
}else {
  pls_sig_raw_cul <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                  data_preprocessed[[1]][, -1],
                  pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cul'),],
                 pred_col = 'pCA_ng_grain')

  pls_sig_msc_cul <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[2]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cul'),],
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_als_cul <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[3]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cul'),],
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_als_msc_cul <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[4]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cul'),],
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_2nd_cul <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[5]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cul'),],
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_raw_cut <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                  data_preprocessed[[1]][, -1],
                  pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cut'),],
                 pred_col = 'pCA_ng_grain')

  pls_sig_msc_cut <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[2]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cut'),],
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_als_cut <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[3]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cut'),],
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_als_msc_cut <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[4]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cut'),],
                   pred_col = 'pCA_ng_grain')
  
  pls_sig_2nd_cut <- run_pls_mod(cbind(pcaMatHeight[,1:10],
                    data_preprocessed[[5]][, -1],
                    pCA_ng_grain = pcaMatHeight[, 11])[which(pcaMatHeight[, 5] == 'cut'),],
                   pred_col = 'pCA_ng_grain')
  
  #Save the pls results as .RData indepecdently (so that they do not exceed the 100 MB limitation from GitHub)
  save(pls_sig_raw_cul,file = file.path('Output_R', 'pls_sig_raw_cul.RData'))
  save(pls_sig_msc_cul, file = file.path('Output_R', 'pls_sig_msc_cul.RData'))
  save(pls_sig_als_cul, file = file.path('Output_R', 'pls_sig_als_cul.RData'))
  save(pls_sig_als_msc_cul, file = file.path('Output_R', 'pls_sig_als_msc_cul.RData'))
  save(pls_sig_2nd_cul, file = file.path('Output_R', 'pls_sig_2nd_cul.RData'))
  
  save(pls_sig_raw_cut,file = file.path('Output_R', 'pls_sig_raw_cut.RData'))
  save(pls_sig_msc_cut, file = file.path('Output_R', 'pls_sig_msc_cut.RData'))
  save(pls_sig_als_cut, file = file.path('Output_R', 'pls_sig_als_cut.RData'))
  save(pls_sig_als_msc_cut, file = file.path('Output_R', 'pls_sig_als_msc_cut.RData'))
  save(pls_sig_2nd_cut, file = file.path('Output_R', 'pls_sig_2nd_cut.RData'))
}
```

### Original spectra

```{r}
#| label: pls - raw spectra
#| echo: false
#| warning: false
#Setup the number of components to be analysed
numComps_raw_cul <- pick_num_comp(pls_sig_raw_cul, 0.95)

#Scatter plot
plot_raw_scatter_cul <- plot_pls_scatter(plsModel = pls_sig_raw_cul,
                                     numComp = numComps_raw_cul)

#Coefficients plot
plot_raw_coefficients_cul <- plot_pls_coefficients(pls_sig_raw_cul,
                                               numComp = numComps_raw_cul,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Loadings plot
plot_raw_loadings_cul <- plot_pls_loadings(pls_sig_raw_cul,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_raw_variances_cul <- plot_pls_variance(pls_sig_raw_cul,
                                               numComp = numComps_raw_cul)

#Residuals
plot_raw_residuals_cul <- plot_pls_residuals(pls_sig_raw_cul,
                                               numComp = numComps_raw_cul,
                                               wavenumberRange = c(800, 1800, 200)
                                         )

#Setup the number of components to be analysed
numComps_raw_cut <- pick_num_comp(pls_sig_raw_cut, 0.95)

#Scatter plot
plot_raw_scatter_cut <- plot_pls_scatter(plsModel = pls_sig_raw_cut,
                                     numComp = numComps_raw_cut)

#Coefficients plot
plot_raw_coefficients_cut <- plot_pls_coefficients(pls_sig_raw_cut,
                                               numComp = numComps_raw_cut,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Loadings plot
plot_raw_loadings_cut <- plot_pls_loadings(pls_sig_raw_cut,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[1]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_raw_variances_cut <- plot_pls_variance(pls_sig_raw_cut,
                                               numComp = numComps_raw_cut)

#Residuals
plot_raw_residuals_cut <- plot_pls_residuals(pls_sig_raw_cut,
                                               numComp = numComps_raw_cut,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - raw spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false
#| message: false

 (plot_raw_spec_cul | plot_raw_scatter_cul) /
    (plot_raw_variances_cul | plot_raw_coefficients_cul) /
    (plot_raw_loadings_cul | plot_raw_residuals_cul)

 (plot_raw_spec_cut | plot_raw_scatter_cut) /
    (plot_raw_variances_cut | plot_raw_coefficients_cut) /
    (plot_raw_loadings_cut | plot_raw_residuals_cut)
```

### Signature area: MSC

```{r}
#| label: pls - sig msc spectra
#| echo: false
#| warning: false
#Setup the number of components to be analysed
numComps_sig_msc_cul <- pick_num_comp(pls_sig_msc_cul, 0.95)

#Scatter plot
plot_sig_msc_scatter_cul <- plot_pls_scatter(plsModel = pls_sig_msc_cul,
                                     numComp = numComps_sig_msc_cul)

#Coefficients plot
plot_sig_msc_coefficients_cul <- plot_pls_coefficients(pls_sig_msc_cul,
                                               numComp = numComps_sig_msc_cul,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[2]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_msc_loadings_cul <- plot_pls_loadings(pls_sig_msc_cul,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[2]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_msc_variances_cul <- plot_pls_variance(pls_sig_msc_cul,
                                               numComp = numComps_sig_msc_cul)

#Residuals
plot_sig_msc_residuals_cul <- plot_pls_residuals(pls_sig_msc_cul,
                                               numComp = numComps_sig_msc_cul,
                                               wavenumberRange = c(800, 1800, 200)
                                         )

#Setup the number of components to be analysed
numComps_sig_msc_cut <- pick_num_comp(pls_sig_msc_cut, 0.95)

#Scatter plot
plot_sig_msc_scatter_cut <- plot_pls_scatter(plsModel = pls_sig_msc_cut,
                                     numComp = numComps_sig_msc_cut)

#Coefficients plot
plot_sig_msc_coefficients_cut <- plot_pls_coefficients(pls_sig_msc_cut,
                                               numComp = numComps_sig_msc_cut,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[2]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_msc_loadings_cut <- plot_pls_loadings(pls_sig_msc_cut,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[2]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_msc_variances_cut <- plot_pls_variance(pls_sig_msc_cut,
                                               numComp = numComps_sig_msc_cut)

#Residuals
plot_sig_msc_residuals_cut <- plot_pls_residuals(pls_sig_msc_cut,
                                               numComp = numComps_sig_msc_cut,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig msc spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false
#| message: false

 (plot_msc_spec_cul | plot_sig_msc_scatter_cul) /
    (plot_sig_msc_variances_cul | plot_sig_msc_coefficients_cul) /
    (plot_sig_msc_loadings_cul | plot_sig_msc_residuals_cul)

 (plot_msc_spec_cut | plot_sig_msc_scatter_cut) /
    (plot_sig_msc_variances_cut | plot_sig_msc_coefficients_cut) /
    (plot_sig_msc_loadings_cut | plot_sig_msc_residuals_cut)
```

### Signature area: smoothing + ALS

```{r}
#| label: pls - sig als spectra
#| echo: false
#| warning: false
#Setup the number of components to be analysed
numComps_sig_als_cul <- pick_num_comp(pls_sig_als_cul, 0.95)

#Scatter plot
plot_sig_als_scatter_cul <- plot_pls_scatter(plsModel = pls_sig_als_cul,
                                     numComp = numComps_sig_als_cul)

#Coefficients plot
plot_sig_als_coefficients_cul <- plot_pls_coefficients(pls_sig_als_cul,
                                               numComp = numComps_sig_als_cul,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[3]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_als_loadings_cul <- plot_pls_loadings(pls_sig_als_cul,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[3]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_als_variances_cul <- plot_pls_variance(pls_sig_als_cul,
                                               numComp = numComps_sig_als_cul)

#Residuals
plot_sig_als_residuals_cul <- plot_pls_residuals(pls_sig_als_cul,
                                               numComp = numComps_sig_als_cul,
                                               wavenumberRange = c(800, 1800, 200)
                                         )

#Setup the number of components to be analysed
numComps_sig_als_cut <- pick_num_comp(pls_sig_als_cut, 0.95)

#Scatter plot
plot_sig_als_scatter_cut <- plot_pls_scatter(plsModel = pls_sig_als_cut,
                                     numComp = numComps_sig_als_cut)

#Coefficients plot
plot_sig_als_coefficients_cut <- plot_pls_coefficients(pls_sig_als_cut,
                                               numComp = numComps_sig_als_cut,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[3]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_als_loadings_cut <- plot_pls_loadings(pls_sig_als_cut,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[3]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_als_variances_cut <- plot_pls_variance(pls_sig_als_cut,
                                               numComp = numComps_sig_als_cut)

#Residuals
plot_sig_als_residuals_cut <- plot_pls_residuals(pls_sig_als_cut,
                                               numComp = numComps_sig_als_cut,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig als spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false
#| message: false

 (plot_als_spec_cul | plot_sig_als_scatter_cul) /
    (plot_sig_als_variances_cul | plot_sig_als_coefficients_cul) /
    (plot_sig_als_loadings_cul | plot_sig_als_residuals_cul)

 (plot_als_spec_cut | plot_sig_als_scatter_cut) /
    (plot_sig_als_variances_cut | plot_sig_als_coefficients_cut) /
    (plot_sig_als_loadings_cut | plot_sig_als_residuals_cut)
```

### Signature area: smoothing + ALS + MSC

```{r}
#| label: pls - sig als msc spectra
#| echo: false
#| warning: false
#Setup the number of components to be analysed
numComps_sig_als_msc_cul <- pick_num_comp(pls_sig_als_msc_cul, 0.95)

#Scatter plot
plot_sig_als_msc_scatter_cul <- plot_pls_scatter(plsModel = pls_sig_als_msc_cul,
                                     numComp = numComps_sig_als_msc_cul)

#Coefficients plot
plot_sig_als_msc_coefficients_cul <- plot_pls_coefficients(pls_sig_als_msc_cul,
                                               numComp = numComps_sig_als_msc_cul,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[4]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_als_msc_loadings_cul <- plot_pls_loadings(pls_sig_als_msc_cul,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[4]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_als_msc_variances_cul <- plot_pls_variance(pls_sig_als_msc_cul,
                                               numComp = numComps_sig_als_msc_cul)

#Residuals
plot_sig_als_msc_residuals_cul <- plot_pls_residuals(pls_sig_als_msc_cul,
                                               numComp = numComps_sig_als_msc_cul,
                                               wavenumberRange = c(800, 1800, 200)
                                         )

#Setup the number of components to be analysed
numComps_sig_als_msc_cut <- pick_num_comp(pls_sig_als_msc_cut, 0.95)

#Scatter plot
plot_sig_als_msc_scatter_cut <- plot_pls_scatter(plsModel = pls_sig_als_msc_cut,
                                     numComp = numComps_sig_als_msc_cut)

#Coefficients plot
plot_sig_als_msc_coefficients_cut <- plot_pls_coefficients(pls_sig_als_msc_cut,
                                               numComp = numComps_sig_als_msc_cut,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[4]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_als_msc_loadings_cut <- plot_pls_loadings(pls_sig_als_msc_cut,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[4]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_als_msc_variances_cut <- plot_pls_variance(pls_sig_als_msc_cut,
                                               numComp = numComps_sig_als_msc_cut)

#Residuals
plot_sig_als_msc_residuals_cut <- plot_pls_residuals(pls_sig_als_msc_cut,
                                               numComp = numComps_sig_als_msc_cut,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig als msc spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false
#| message: false

 (plot_als_msc_spec_cul | plot_sig_als_msc_scatter_cul) /
    (plot_sig_als_msc_variances_cul | plot_sig_als_msc_coefficients_cul) /
    (plot_sig_als_msc_loadings_cul | plot_sig_als_msc_residuals_cul)

 (plot_als_msc_spec_cut | plot_sig_als_msc_scatter_cut) /
    (plot_sig_als_msc_variances_cut | plot_sig_als_msc_coefficients_cut) /
    (plot_sig_als_msc_loadings_cut | plot_sig_als_msc_residuals_cut)
```

### Signature area: smoothing + 2^nd^ derivative

```{r}
#| label: pls - sig 2nd derivative spectra
#| echo: false
#| warning: false
#Setup the number of components to be analysed
numComps_sig_2nd_cul <- pick_num_comp(pls_sig_2nd_cul, 0.85)

#Scatter plot
plot_sig_2nd_scatter_cul <- plot_pls_scatter(plsModel = pls_sig_2nd_cul,
                                     numComp = numComps_sig_2nd_cul)

#Coefficients plot
plot_sig_2nd_coefficients_cul <- plot_pls_coefficients(pls_sig_2nd_cul,
                                               numComp = numComps_sig_2nd_cul,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[5]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_2nd_loadings_cul <- plot_pls_loadings(pls_sig_2nd_cul,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[5]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_2nd_variances_cul <- plot_pls_variance(pls_sig_2nd_cul,
                                               numComp = numComps_sig_2nd_cul)

#Residuals
plot_sig_2nd_residuals_cul <- plot_pls_residuals(pls_sig_2nd_cul,
                                               numComp = numComps_sig_2nd_cul,
                                               wavenumberRange = c(800, 1800, 200)
                                         )

#Setup the number of components to be analysed
numComps_sig_2nd_cut <- pick_num_comp(pls_sig_2nd_cut, 0.85)

#Scatter plot
plot_sig_2nd_scatter_cut <- plot_pls_scatter(plsModel = pls_sig_2nd_cut,
                                     numComp = numComps_sig_2nd_cut)

#Coefficients plot
plot_sig_2nd_coefficients_cut <- plot_pls_coefficients(pls_sig_2nd_cut,
                                               numComp = numComps_sig_2nd_cut,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[5]])[-1])
                                                          )
                                               )

#Loadings plot
plot_sig_2nd_loadings_cut <- plot_pls_loadings(pls_sig_2nd_cut,
                                               numComp = 1,
                                               wavenumberRange = c(800, 1800, 200),
                                               peakPos = (as.numeric(colnames(data_preprocessed[[5]])[-1])
                                                          )
                                               )

#Variance explained by different components
plot_sig_2nd_variances_cut <- plot_pls_variance(pls_sig_2nd_cut,
                                               numComp = numComps_sig_2nd_cut)

#Residuals
plot_sig_2nd_residuals_cut <- plot_pls_residuals(pls_sig_2nd_cut,
                                               numComp = numComps_sig_2nd_cut,
                                               wavenumberRange = c(800, 1800, 200)
                                         )
```

```{r}
#| label: pls - sig 2nd derivative spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false
#| message: false

 (plot_2nd_spec_cul | plot_sig_2nd_scatter_cul) /
    (plot_sig_2nd_variances_cul | plot_sig_2nd_coefficients_cul) /
    (plot_sig_2nd_loadings_cul | plot_sig_2nd_residuals_cul)

 (plot_2nd_spec_cut | plot_sig_2nd_scatter_cut) /
    (plot_sig_2nd_variances_cut | plot_sig_2nd_coefficients_cut) /
    (plot_sig_2nd_loadings_cut | plot_sig_2nd_residuals_cut)
```

```{r}
#| label: combine malformation data with PLS and GCMS
#| echo: false
#| warning: false
#Read the cutting branch data, and replace all na with 0
cuttingsRaw <- read.csv(file = 
                          file.path('Data_R', 
                                    'Malformations_Cuttings.csv'),
                        row.names = 1)

#Replace all NAs with 0
cuttingsRaw[is.na(cuttingsRaw)] <- 0

#Extract the batch information
batchInfoTemp <- strsplit(x = cuttingsRaw[,1], split = '_')
cuttingsBatchInfo <- cbind(rep('Cuttings', length(batchInfoTemp)),
                           rep('2025', length(batchInfoTemp)),
                           matrix(unlist(batchInfoTemp), ncol = 4, byrow = TRUE))
colnames(cuttingsBatchInfo) <- c('Sample', 'Year', 'Frame', 'Treatment',
                                 'Plant', 'Duplicate')

#Delete unused variables
rm(batchInfoTemp)

#Read the cultivar branch data, and replace all na with 0
cultivarsRaw <- read.csv(file = 
                           file.path('Data_R', 
                                     'Malformations_Cultivars.csv'),
                         row.names = 1)

#Replace all NAs with 0
cultivarsRaw[is.na(cultivarsRaw)] <- 0

#Extract the batch information
batchInfoTemp <- strsplit(x = cultivarsRaw[,1], split = '_')
cultivarsBatchInfo <- cbind(rep('Cultivars', length(batchInfoTemp)),
                            matrix(unlist(batchInfoTemp), ncol = 4, byrow = TRUE))
colnames(cultivarsBatchInfo) <- c('Sample', 'Year', 'Frame', 'Treatment',
                                  'Plant')

#Delete unused variables
rm(batchInfoTemp)

#Combine counting results in duplicated samples, by summing them up
tempVec <- unique(cultivarsRaw[, 1])

#Matrix for all malformation data
tempMat <- matrix(nrow = length(tempVec),
                  ncol = ncol(cultivarsRaw),
                  dimnames = list(c(),
                                  colnames(cultivarsRaw)))

#Matrix for all batch information on malformation data
tempMat2 <- matrix(nrow = length(tempVec),
                   ncol = ncol(cultivarsBatchInfo),
                   dimnames = list(c(),
                                   colnames(cultivarsBatchInfo)))

#Fill sample label into matrix
tempMat[, 1] <- tempVec

#Loop to add duplicates together
for (loopI in 1:nrow(tempMat)) {
  
  #Check  if there are duplicates in current sample
  if (length(which(cultivarsRaw[, 1] == tempMat[loopI, 1])) > 1) {
    
    #Sum up all counting results among duplicates
    tempMat[loopI, 2:ncol(tempMat)] <- apply(cultivarsRaw[
      which(cultivarsRaw[, 1] == tempMat[loopI, 1]), 
      2:ncol(tempMat)
    ], 2, sum)
    
  }else{
    
    #Transfer couting results into matrix
    tempMat[loopI, 2:ncol(tempMat)] <- cultivarsRaw[
      which(cultivarsRaw[, 1] == tempMat[loopI, 1]), 2:ncol(tempMat)
    ]
    
  }
  
  #Record the corresponding batch information
  tempMat2[loopI,] <- cultivarsBatchInfo[which(cultivarsRaw[, 1] == tempMat[loopI, 1])[1],]
  
}

#Combine batch information with counting results
cultivarsCom <- cbind(tempMat2, tempMat[, -1])

rm(loopI,
   tempMat,
   tempMat2,
   tempVec)

#Combine malformation data with PLS fitted results from different pretreatment methodologies
data_preprocessed_ftir_gcms_malformation <- vector(mode = 'list',
                                                   length = length(data_preprocessed))

#Signature area Raw spectra
data_preprocessed_ftir_gcms_malformation[[1]] <- matrix_malformation(
  batchInfoMat = pcaMatHeight[,1:10],
  PLSFittedValues = c(pls_sig_raw_cul$fitted.values[,, numComps_raw_cul],
                      pls_sig_raw_cut$fitted.values[,, numComps_raw_cut]),
  GCMSValuesMat = pcaMatHeight[,11:12],
  cultivarsData = cultivarsCom,
  cuttingsData = cuttingsRaw,
  cuttingsBatch = cuttingsBatchInfo)

#Signature area MSC spectra
data_preprocessed_ftir_gcms_malformation[[2]] <- matrix_malformation(
  batchInfoMat = pcaMatHeight[,1:10],
  PLSFittedValues = c(pls_sig_msc_cul$fitted.values[,, numComps_sig_msc_cul],
                      pls_sig_msc_cut$fitted.values[,, numComps_sig_msc_cut]),
  GCMSValuesMat = pcaMatHeight[,11:12],
  cultivarsData = cultivarsCom,
  cuttingsData = cuttingsRaw,
  cuttingsBatch = cuttingsBatchInfo)

#Signature area ALS spectra
data_preprocessed_ftir_gcms_malformation[[3]] <- matrix_malformation(
  batchInfoMat = pcaMatHeight[,1:10],
  PLSFittedValues = c(pls_sig_als_cul$fitted.values[,, numComps_sig_als_cul],
                      pls_sig_als_cut$fitted.values[,, numComps_sig_als_cut]),
  GCMSValuesMat = pcaMatHeight[,11:12],
  cultivarsData = cultivarsCom,
  cuttingsData = cuttingsRaw,
  cuttingsBatch = cuttingsBatchInfo)

#Signature area ALS-MSC spectra
data_preprocessed_ftir_gcms_malformation[[4]] <- matrix_malformation(
  batchInfoMat = pcaMatHeight[,1:10],
  PLSFittedValues = c(pls_sig_als_msc_cul$fitted.values[,, numComps_sig_als_msc_cul],
                      pls_sig_als_msc_cut$fitted.values[,, numComps_sig_als_msc_cut]),
  GCMSValuesMat = pcaMatHeight[,11:12],
  cultivarsData = cultivarsCom,
  cuttingsData = cuttingsRaw,
  cuttingsBatch = cuttingsBatchInfo)

#Signature area 2nd derivative spectra
data_preprocessed_ftir_gcms_malformation[[5]] <- matrix_malformation(
  batchInfoMat = pcaMatHeight[,1:10],
  PLSFittedValues = c(pls_sig_2nd_cul$fitted.values[,, numComps_sig_2nd_cul],
                      pls_sig_2nd_cut$fitted.values[,, numComps_sig_2nd_cut]),
  GCMSValuesMat = pcaMatHeight[,11:12],
  cultivarsData = cultivarsCom,
  cuttingsData = cuttingsRaw,
  cuttingsBatch = cuttingsBatchInfo)
```

## Overviews on GC-MS signals and FTIR PLS fitted values

```{r}
#| label: preperation on presenting GC-MS and FTIR data
#| echo: false
#| warning: false

#Overall plot on PLS fitted values
plot_PLS_Fitted <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_PLS_Fitted)) {
  plot_PLS_Fitted[[loopI]] <- plot_box(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'PLS')
}

#Plot groupped by potted cultivars and branch cuttings
plot_PLS_Fitted_Materials <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_PLS_Fitted)) {
  plot_PLS_Fitted_Materials[[loopI]] <- plot_box(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'PLS',
                                       groupName = TRUE)
}

#Overall plot on GCMS signals
plot_GCMS <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_GCMS)) {
  plot_GCMS[[loopI]] <- plot_box(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'GCMS')
}

#Plot groupped by potted cultivars and branch cuttings
plot_GCMS_Materials <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_GCMS_Materials)) {
  plot_GCMS_Materials[[loopI]] <- plot_box(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'GCMS',
                                       groupName = TRUE)
}

#Plant-groupped PLS fitted values
plot_PLS_Fitted_Plant <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_PLS_Fitted_Plant)) {
  plot_PLS_Fitted_Plant[[loopI]] <- plot_box_plant(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[1]],
                                       variableName = 'PLS')
}

#Plant-groupped GCMS signals
plot_GCMS_Plant <- vector(mode = 'list',
                          length = length(data_preprocessed))

for (loopI in 1:length(plot_GCMS_Plant)) {
  plot_GCMS_Plant[[loopI]] <- plot_box_plant(dataMat = data_preprocessed_ftir_gcms_malformation[[loopI]][[2]],
                                       variableName = 'GCMS')
}

```

### Raw spectra

```{r}
#| label: overview on GCMS and FTIR - raw spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

(plot_GCMS[[1]]|plot_PLS_Fitted[[1]]) /
  (plot_GCMS_Materials[[1]]|plot_PLS_Fitted_Materials[[1]]) /
  (plot_GCMS_Plant[[1]]|plot_PLS_Fitted_Plant[[1]])
```

### Signature area: MSC

```{r}
#| label: overview on GCMS and FTIR - msc spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

(plot_GCMS[[2]]|plot_PLS_Fitted[[2]]) /
  (plot_GCMS_Materials[[2]]|plot_PLS_Fitted_Materials[[2]]) /
  (plot_GCMS_Plant[[2]]|plot_PLS_Fitted_Plant[[2]])
```

### Signature area: ALS

```{r}
#| label: overview on GCMS and FTIR - als spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

(plot_GCMS[[3]]|plot_PLS_Fitted[[3]]) /
  (plot_GCMS_Materials[[3]]|plot_PLS_Fitted_Materials[[3]]) /
  (plot_GCMS_Plant[[3]]|plot_PLS_Fitted_Plant[[3]])
```

### Signature area: ALS + MSC

```{r}
#| label: overview on GCMS and FTIR - als msc spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

(plot_GCMS[[4]]|plot_PLS_Fitted[[4]]) /
  (plot_GCMS_Materials[[4]]|plot_PLS_Fitted_Materials[[4]]) /
  (plot_GCMS_Plant[[4]]|plot_PLS_Fitted_Plant[[4]])
```

### Signature area: 2^nd^ derivative

```{r}
#| label: overview on GCMS and FTIR - 2nd derivative spectra - figure
#| fig-width: 8
#| fig-height: 12
#| fig-dpi: 600
#| warning: false

(plot_GCMS[[5]]|plot_PLS_Fitted[[5]]) /
  (plot_GCMS_Materials[[5]]|plot_PLS_Fitted_Materials[[5]]) /
  (plot_GCMS_Plant[[5]]|plot_PLS_Fitted_Plant[[5]])
```

## Malformation rates of pollen grains

```{r}
#| label: preperation on malformation plotting
#| echo: false
#| warning: false
#Calculate malformation rates in pot cultivars
tempVec <- (cultivarsRaw[, 15] - cultivarsRaw[, 2])/ cultivarsRaw[, 2]
cultivarMean <- vector(length = 4)
cultivarMean[1] <- mean(tempVec[which(cultivarsBatchInfo[, 4] == 'UV')])
cultivarMean[2] <- sd(tempVec[which(cultivarsBatchInfo[, 4] == 'UV')])
cultivarMean[3] <- mean(tempVec[which(cultivarsBatchInfo[, 4] == 'CON')])
cultivarMean[4] <- sd(tempVec[which(cultivarsBatchInfo[, 4] == 'CON')])

#Point plot to compare ratios between con and uv treatment, on different plants
cuttingsPlantsPlot <-
  ggplot() +
  geom_point(data = data_preprocessed_ftir_gcms_malformation[[1]][[3]],
             aes(x = CON_Mal, y = UV_Mal, col = Plant)) +
  geom_point(aes(x = cultivarMean[3] , y = cultivarMean[1], 
                 color = 'Cultivars')) +
  geom_errorbar(data = data_preprocessed_ftir_gcms_malformation[[1]][[3]],
                aes(x = CON_Mal, y = UV_Mal, col = Plant,
                    ymin = UV_Mal - UV_Mal_sd,
                    ymax = UV_Mal + UV_Mal_sd),
                linewidth = 1,
                width = 0.01) +
  geom_errorbar(data = data_preprocessed_ftir_gcms_malformation[[1]][[3]],
                aes(x = CON_Mal, y = UV_Mal, col = Plant,
                    xmin = CON_Mal - CON_Mal_sd,
                    xmax = CON_Mal + CON_Mal_sd),
                linewidth = 1,
                width = 0.01,
                orientation = 'y') +
  geom_errorbar(aes(x = cultivarMean[3] , y = cultivarMean[1], color = 'Cultivars',
                    ymin = cultivarMean[1] - cultivarMean[2],
                    ymax = cultivarMean[1] + cultivarMean[2]),
                linewidth = 1,
                width = 0.005) +
  geom_errorbar(aes(x = cultivarMean[3] , y = cultivarMean[1], color = 'Cultivars',
                    xmin = cultivarMean[3] - cultivarMean[4],
                    xmax = cultivarMean[3] + cultivarMean[4]),
                linewidth = 1,
                width = 0.005,
                orientation = 'y') +
  geom_abline (slope=1, linetype = "dashed", color="grey") +
  scale_x_continuous(limits = c(-0.05, 0.15), name = 'CON malformation percentage',
                     breaks = seq(0, 0.15, by = 0.05)) +
  scale_y_continuous(limits = c(-0.05, 0.25), name = 'UV malformation percentage',
                     breaks = seq(0, 0.25, by = 0.05)) +
  scale_color_manual(name = 'Plant',
                     values = c('1' = brewer.pal(6, 'Set2')[1],
                                '2' = brewer.pal(6, 'Set2')[2],
                                '3' = brewer.pal(6, 'Set2')[3],
                                '4' = brewer.pal(6, 'Set2')[4],
                                '5' = brewer.pal(6, 'Set2')[5],
                                '6' = brewer.pal(6, 'Set2')[6],
                                'Cultivars' = 'black')) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    axis.title.x = element_text(size = 20, family = 'arial'),
    axis.text.x = element_text(size = 16, family = 'arial'),
    
    axis.text.y = element_text(size = 20, family = 'arial'),
    axis.title.y = element_text(size = 16, family = 'arial'),
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.2),
    legend.text = element_text(size = 14, family = 'arial')
  )

rm(tempVec)
```

```{r}
#| label: malformation - figure
#| fig-width: 8
#| fig-height: 8
#| fig-dpi: 600
#| warning: false
cuttingsPlantsPlot
```
