---
title: "UAC signals in Pine cuttings and cultivars"
format:
  html:
    code-fold: true
    code-tools: true
editor: visual
---

```{r}
#| label: load Packages
#| echo: true
#| output: false
#Load palaeochem package
library(palaeochem)

#Packages needed in Vivian's code
source(
  file.path(
    "Code_R", "00_Config_file.r"
  )
)
source(
  file.path(
    "~/palaeochem/R/get_data_wide.r"
  )
)
```

```{r}
#| label: load FTIRS pectra
#| echo: true
# This workflow consumes the saved artefact.
ftir_processed <- readRDS(here::here("Data_R", "Palaechem_data_compilation_2026-04-09.rds"))
```

```{r}
#| label: extract GH2025 spectra from general dataset
#| echo: true
#| warning: false

# restructure and nest spectra
ftir_spectra_compilation <- ftir_processed |>
  tidyr::nest(spectra_raw = -c(
    dataset_name,
    project,
    data_collector,
    location,
    sample_year,
    ftir_method,
    taxon_name
  )
  ) |>
  dplyr::mutate(
    df_data_raw = purrr::map(
      .x = spectra_raw,
      .f = ~get_spectra_matrix_wide(.x)
    ),
    matrix_spectra_raw = purrr::map(
      .x = df_data_raw,
      .f = ~get_spectra_matrix(.x)
    ),
    sample_meta_data = purrr::map(
        .x = df_data_raw,
        .f = ~get_sample_meta_data(.x)
    )
  ) |>
  dplyr::select(-spectra_raw)
```

## Original FTIR spectra

The original FTIR spectra are between 4000 and 600 cm^-1^, with wavenumber interval of 4 cm^-1^. The original spectral dataset looks like:

```{r}
#| label: Figure full raw spectra
#| echo: true
#| warning: false

data_long <- data_to_process$df_data_raw |>
pivot_longer(
  dplyr::matches("\\d+"),
  names_to = "wavenumber",
  values_to = "absorbance"
) 

plot_original_spec <-
data_long |> plot_spectra()

print(plot_original_spec)  

```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
