---
title: "UAC signals in Pine cuttings and cultivars"
format:
  html:
    code-fold: true
    code-tools: true
editor: visual
---

```{r}
#| label: loadPackages
#Read OPUS files
library(opusreader)
library(palaeochem)

#Declare a function to automatically split the name of spectral file, 
#to get all necessary
#folderNames - single character string, the original file name to be split
#to extract the batch information
#seperator - single character string, the symbol used in the original file 
#name for separating different types of batch information
#defaultSettings - vector with length of 10, indicating the default batch
#information
namesSplit <- function(folderNames, seperator, defaultSettings) {
  
  splitedNames <- defaultSettings
  
  #Split the names
  splitedNamesRaw <- strsplit( x = folderNames, split = seperator)[[1]]
  
  #Check the information given in the folder name, 
  #and change the corresponding sections in output if necessary
  if(splitedNamesRaw[2] != 'GH'){
    
    splitedNames[c(1:8, 10)] <- splitedNamesRaw[c(1:9)]
    
  } else if(splitedNamesRaw[4] == 'cul'){
    
    splitedNames[c(1,3:8, 10)] <- splitedNamesRaw[c(1:8)]
    
  }else {
    
    splitedNames[c(1,3:10)] <- splitedNamesRaw[c(1:9)]
    
  }
  
  if(splitedNamesRaw[5]  == 'ROOF' | splitedNamesRaw[6] == 'ROOF') {
    
    splitedNames[c(1,3:5, 8,10)]  <- splitedNamesRaw[c(1,5,3,4,6,7)]
    
    splitedNames[c(6,7,9)]  <- NA
    
  }
  
  return(splitedNames)
}

  #Declare a function to calculate the number of all elements in a list given
#tempList - list to be used for calculation
numOfEle <- function(tempList) {
  
  #Vector storing the number
  tempVec <- 0
  
  for (loopI in 1:length(tempList)) {
    tempVec <- tempVec + length(tempList[[loopI]])
  }
  
  return(tempVec)
}

#List all the folder for different measuring days
daysVec <- list.files(path = file.path('Data_R', 'FTIR_Raw_Spectra'))

#Loop to list all .o files in each folder above
#List for storing all names of spectra files
namesList <- list()
for (loopI in 1:length(daysVec)) {
  
  namesList[[length(namesList) + 1]] <- 
    list.files(path = file.path('Data_R', 'FTIR_Raw_Spectra',
                                  daysVec[loopI]), 
               pattern = '\\.0$')
  
}

#Loop to read all spectra
#List for storing all spectra
rawSpecList <- list()

#Matrix for storing all batch information
rawSpecBatchInfo <- matrix(nrow = numOfEle(namesList),
                           ncol = 10,
                           dimnames = list(c(),
                                           c('Exp_Year',
                                             'Sample_Year',
                                             'Location',
                                             'Species',
                                             'Materials',
                                             'Group',
                                             'Treatment',
                                             'Sample_ID',
                                             'Sample_Dup',
                                             'FTIR_Dup')))


#Two recorders on row index for matrix filling
tempInd <- 0

for (loopI in 1:length(daysVec)) {
  
  rawSpecList[[loopI]] <- list()
  
  for (loopJ in 1:length(namesList[[loopI]])) {
    rawSpecList[[loopI]][[loopJ]] <- 
      opus_read(file = file.path('Data_R','FTIR_Raw_Spectra', 
                                 daysVec[loopI], 
                                 namesList[[loopI]][loopJ]
      ),
      simplify = TRUE)
    
    #Extract the sample information from the file names at the same time
    tempVec <- namesSplit(substr(namesList[[loopI]][loopJ],
                                 1,
                                 nchar(namesList[[loopI]][loopJ]) - 2), '_',
                          c(2025,2025,
                            'GH', 'Pmu', 
                            'cul', 'A', 
                            'UV', '001',
                            '1', '1'))
    
    
    tempInd <- tempInd + 1
    
    rawSpecBatchInfo[tempInd,] <- tempVec
    
    rm(tempVec)
  }
}

#Create a matrix to store all raw spectra
rawSpecMat <- matrix(nrow = numOfEle(namesList),
                     ncol = length(rawSpecList[[1]][[1]][['spec']]),
                     dimnames = list(c(),
                                     c(rawSpecList[[1]][[1]][['wavenumbers']])
                     )
)

#Loop to fill the matrix
#Create a vector recording the current index
tempIndRawSpec <- 0

for (loopI in 1:length(rawSpecList)) {
  
  for (loopJ in 1:length(rawSpecList[[loopI]])) {
    
    tempIndRawSpec <- tempIndRawSpec + 1
    
    rawSpecMat[tempIndRawSpec,] <- rawSpecList[[loopI]][[loopJ]][['spec']]
  }
}
```

## Original FTIR spectra

The original FTIR spectra are between 4000 and 600 cm^-1^, with wavenumber interval of 4 cm^-1^.

```{r}         

```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
